### 2020-03-28 paper.tex
first commit.

\documentclass[preprint,12pt]{elsarticle}
\usepackage{etex}

\makeatletter
\providecommand{\doi}[1]{%
	\begingroup
	\let\bibinfo\@secondoftwo
	\urlstyle{rm}%
	\href{http://dx.doi.org/#1}{%
		doi:\discretionary{}{}{}%
		\nolinkurl{#1}%
	}%
	\endgroup
}
\makeatother
%\usepackage{natbib}
% encoding and languages
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[ngerman,english]{babel}
\usepackage{wrapfig}
\usepackage{times}
\usepackage{paralist}
\usepackage{graphicx, color}
%\usepackage{subfigure}
\usepackage{subfig}
\usepackage{subfloat}
\usepackage{booktabs}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{paralist}
\usepackage{amssymb, amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsopn}
\newtheorem{myProb}{Problem}
\newtheorem{mythm}{Theorem}
\newtheorem{myproof}{Proof}
\newtheorem{mysketch}{Proof Sketch}
%\usepackage[ruled]{algorithm2e}
%\usepackage[colorinlistoftodos]{todonotes}
%\usepackage{qtree}
\usepackage{tikz, tikz-qtree}
\usepackage[ruled]{algorithm2e}
%\usepackage{Algorithmic}
\usepackage{times}
\usepackage{microtype}
\usepackage{url}
\usepackage{balance}
\let\proof\relax 
\let\endproof\relax
\usepackage{multirow}
\usepackage{relsize}
\usepackage{caption}
\usepackage{color, colortbl}
\usepackage{array, booktabs, ragged2e}
\usepackage{makecell}
\usepackage{tikz}
\usepackage{textpos}

\usepackage{pdflscape}
\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}

\graphicspath{{./Figures/}}

\newcommand{\mypar}[1]{\smallskip\noindent\textbf{#1.}}

\newtheorem{prob}{Problem}

\newtheorem{mydef}{Definition}
%\newtheorem{remark}{Remark}

\newcommand{\todo}[1]{{\textcolor{red}{\bf {#1}}}}

%\newcommand{\mytitle}{Exploiting Software and Hardware Advances  in Optimization Methods for Improved Wind Farm Design}
\newcommand{\mytitle}{Exploiting Hardware and Software Advances 
 for Quadratic Models of Wind Farm Layout Optimization}
%\newcommand{\mytitlerunning}{\mytitle}
%\newcommand{\myauthor}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}
%\newcommand{\myauthorhyperref}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}

%\hypersetup{
%    bookmarks=true,        % show bookmarks bar?
%    unicode=true,          % non-latin characters in bookmarks?
%    pdffitwindow=false,    % window fit to page when opened
%    pdfstartview={FitH},   % fits width of the page to the window
%    pdftoolbar=true,       % show acrobat toolbar?
%    pdfmenubar=true,       % show acrobat menu?
%    colorlinks=true,      % false: boxed links; true: color links
%    linkcolor={blue!80!black},       % color of internal links
%    urlcolor={blue!30!black},        % color of external links
%    citecolor={green!60!black},       % color of links to bibliography
%    filecolor=black,       % color of file links
%    pdftitle={\mytitle},           % title
%    pdfauthor={\myauthorhyperref}, % author
%    pdfcreator={\myauthorhyperref} % creator of pdf
%}

\begin{document}

\title{\mytitle}
%\subtitle{Extended Abstract}
%\author{}
%\institute{}
%\author{\myauthor}
%\institute{Department of Mechanical and Industrial Engineering, University of Toronto\\
%\email{sariks@mie.utoronto.ca, ecohen@mie.utoronto.ca, jcb@mie.utoronto.ca}}

\author[add1]{Arik Senderovich}

%\ead[url]{https://ischool.utoronto.ca/profile/arik-senderovich/}
\ead{arik.senderovich@utoronto.ca}

\author[add2]{Eldan Cohen}
\ead{ecohen@mie.utoronto.ca}
%\ead[url]{homepageldan}

%\address[1]{}

%\address[2]{40 St. George St., Toronto, Canada}
\author[add2]{Jiachen Zhang}
\ead{jasonzjc@mie.utoronto.ca}

\author[add2]{J. Christopher Beck\corref{cor1}}
\ead{jcb@mie.utoronto.ca}

%\ead[url]{homepageldan}
%\address[3]{}
\address[add1]{Faculty of Information, University of Toronto, 140 St. George St., Toronto, Canada}
\cortext[cor1]{Corresponding author}
\address[add2]{Department of Mechanical and Industrial Engineering, University of Toronto, 40 St. George St., Toronto, Canada}

\begin{abstract}
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem concerns
selecting turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic constraint as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model excels at finding high quality solutions very quickly while the QCOP model is able to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

\maketitle

\section{Introduction}

We revisit 
the wind farm layout optimization (WFLO) problem 
that was first considered 
in~\citet{MOSETTI1994105}. 
In WFLO, we aim to place wind turbines 
in a predefined area to maximize 
the total generated power. 
Turbine location decisions    
influence
the
total power production due to  
interference effects, aka wakes, generated by upstream 
turbines. Furthermore, 
there are proximity constraints between turbines
that prevent from turbines to be placed next to each other.
These inter-turbine effects 
are naturally captured using quadratic constraints and objective
functions. Several exact and heuristic optimization approaches have 
been proposed to solve
the WFLO problem~\cite{turner2014new,Zhang2014}. However, these
approaches represented the quadratic effects and constraints
using linear models that approximate the real dependencies.

In this work, we introduce two 
novel quadratic WFLO models. The first model 
represents WFLO using a 
quadratic constrained optimization problem (QCOP). 
The QCOP can be solved using state-of-the-art solvers that
specialize in such problems, e.g., Gurobi~\cite{guorbi}. 
Our second model is based on a quadratic unconstrained binary optimization (QUBO) representation of the WFLO 
problem.
This enables the use of specialized optimization hardware tailored to solve
QUBOs. In this paper, we use Fujitsu's digital annealer to demonstrate the strength
of the approach. The main contributions of this paper are: 
\begin{enumerate} 
\item We propose two novel quadratic WFLO models that capture wake and proximity
effects directly without going through a linear approximation, and, 
\item We test the two formulations
on a commerical software-based quadratic solver (Gurobi) and on Fujitsu's digital 
annealer. 
\end{enumerate} Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we solve 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new}). 
We show, empirically, that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance, often outperforming
existing solutions when provided sufficient run-time. Furthermore, we
we observe that the digital 
annealer often provides order of magnitude faster feasible solutions
compared to our QCOP based approach.

The remainder is structured as follows \todo{adjust structure}.

\section{Background}
\label{sec:related}

In this part, we first present the
physical model of a wind farm that we use throughout the paper. 
Subsequently, we first provide an overview of 
existing mathematical
optimization methods for WFLO, focusing
on their similarities and differences
compared to our approach.

\subsection{Wind Farm Model}

We consider a square shaped wind farm that we model 
as an $N\times N$ grid
with $N$ being the number of cells in each axis. 
Each cell has the area of $c^2$ (in meters), i.e.,
if $c=100$ meters and $N=20$, the farm area will be $2km\times 2km$.
In this work, we assume that the terrain is flat and that 
we have no constraints related to the terrain.
Furthermore, we assume a sterile environment and do not consider the influence of noise
on the wind farm's surroundings. The wind farm literature treats the more complex
cases of non-flat terrain and environmental considerations in~\cite{} and~\cite{}, respectively.
In our work, we introduce a novel optimization approach that can be extended,
in the future, to include 
such considerations.

In principle, a higher number of turbines placed in the wind farm produces more energy. However,
we must consider two types of interference between the turbines.

First, 
turbines cannot be placed at high proximity to each other. 
Specifically, the turbines must be placed at least five rotor diameters apart from each other. Depending on the cell size, we shall formulate the corresponding constraints as part of the optimization model. 
For example, assuming that a turbine is placed
at the center of the cell (cell size of $100$ meters) and for rotor diameter of $40$ meters, 
we must ensure that two turbines will not placed in adjacent cells.

Second, turbines influence each other through an effect called \emph{wake}. Wakes cause reduction of the effective power produced by a turbine due to upstream turbines changing airflow dynamics.

\todo{Explain the wake model - single wake and linear superposition. Explain that it is less accurate but leads to a quadratic objective. We then assess it using sum of squares. Explain that there are $m$ turbines to be placed. use Zheng and use Mosetti.}

\todo{Wind direction}

\subsection{Optimization Methods for WFLO}

The allocation of turbines in a wind farm was first considered 
as an optimization problem in~\cite{MOSETTI1994105}. Since
then, WFLO has caught the attention of optimization researchers. 
The WFLO problem can either be solved \emph{exactly} using
techniques from operations research (e.g., linear programming, 
mixed-integer programming, stochastic programming, and constraint programming)\cite{Zhang2014,turner2014new}
or solved \emph{approximately} using various meta-heuristics (e.g., genetic algorithms and local search) \cite{MOSETTI1994105}.
The main advantage of the former techniques
is that they provide guarantees on solution quality, 
while the latter methods generate 
quick and sufficiently accurate solutions. In this work, we compare 
exact and approximate methods for WFLO. Both use 
quadratic optimization models... \todo{add more variations and explain that we
	focus on one to demonstrate the usefulness of quadratic models}

\section{Quadratic Models for WFLO}
\label{sec:QUBO4WFLO}

In this part, 
we formulate the WFLO problem
as two quadratic optimization problems: a
quadratic constrained optimization (QCOP) problem and 
a quadratic unconstrained binary optimization (QUBO) problem. 
To this end,
we follow standard WFLO definitions
and notation presented in~\cite{Zhang2014}. First, we present WFLO's inputs and decision variables.
Subsequently, we present the mathematical formulation of the two problems.

\paragraph{WFLO Inputs and Decision Variables} Let $x_i$ be a binary decision variable that represents whether a 
turbine is positioned at location $i \in \mathcal{N}$ with $\mathcal{N}$ being the set 
of $n$ possible locations and 
$m$ being the total number of required turbines. Note that we assume a discrete grid with 
a finite number of possible positions. We 
write $x$ as shorthand for $x = (x_1,\ldots, x_n)$. Furthermore,
we denote by $\mathcal{E} \subseteq \mathcal{N}\times \mathcal{N}$
the set of location pairs that cannot simultaneously host turbines 
due to proximity constraints. 
Let $u_{id}$ and $u_{id, \infty} $ 
be the wind speeds at turbine location $i$ for wind regime $d \in \mathcal{D}$
with and without interference from other turbines due to wake effects, respectively. A wind regime is a combination 
of wind speed and direction. 
The probability of wind regime 
$d$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. For example, 
regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind.

Next, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
the set of downstream 
turbine locations 
given that a turbine is placed at cell $i$ for wind regime $d$ (i.e., turbines placed in these locations 
will be influenced by a turbine placed at $i$). Finally,
we denote by $u_{ijd}$ the wind speed at turbine $j$ due to a single wake from upstream turbine $i$ with $j \in \mathcal{U}_{id}$. 
The following objective function  
represents the total expected energy of a wind farm given placement decisions 
$x$: \begin{equation}
f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \label{ObjFunc}\\
%&s.t.:& \nonumber\\
%&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
%&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
%&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{equation}

\subsection{WFLO as QCOP}

To 
maximize $f(x)$ and satisfy the number of turbines and proximity 
constraints we can 
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{eqnarray}

Solving QCOP with Gurobi. Explain.

\subsection{WFLO as QUBO}

The equivalent QUBO formulation is given by,
\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
with the term multiplied by $\lambda_1$ 
ensures that exactly $m$ turbines are placed, while the term multiplied by $\lambda_2$ penalizes
solutions that place turbines in adjacent locations. The two penalty terms $\lambda_1$ and $\lambda_2$ set large enough to ensure constraint satisfaction (see Section~\ref{sec:theory}). The QUBO in Eq. (\ref{QUBO}) allows us
to solve the WFLO problem using quantum-inspired hardware.

\section{Evaluation}
\label{sec:eval}

\subsection{Experimental Design}

\paragraph{WFLO Instances}

\paragraph{Optimization Benchmarks}

\paragraph{Experimental Setting and Implementation Details}

\subsection{Main Results}

\subsection{Discussion \& Limitations}

\section{Conclusion}
\label{sec:conclusion}

\bibliographystyle{elsarticle-num-names}
%\bibliographystyle{plainnat}
\bibliography{Biblio}

\end{document}

%
%\begin{table}[t]
%	
%
%	\label{tab:results1}
%	\vspace{-0.5em}
%	\scriptsize
%	\centering
%	\begin{tabular}{ c @{\hspace{1em}} c @{\hspace{1em}} c @{\hspace{1em}} c}
%		\toprule
%		Paper & Exact Optimization & Discrete Grid & Flat Terrain  \\
%		\midrule
%		\citet{Zhang2014,turner2014new} & + & + & +  \\
%		\citet{kuo2016wind} & + & + & -  \\
%		 \citet{} & + & - & + \\
%		
%		\bottomrule
%	\end{tabular}
%	\vspace{-1em}
%	\caption{Literature review of Wind Farm Optimization Layout solutions.}
%		
%\end{table}

### 2020-04-06 paper.tex
commit.

@@
 \usepackage{subfig}
-\usepackage{subfloat}
+%\usepackage{subfloat}
 \usepackage{booktabs}
@@
 \newtheorem{mythm}{Theorem}
+
 \newtheorem{myproof}{Proof}
@@
 selecting turbine positions to maximize energy production.
-While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic constraint as, indeed, has been done in some existing optimization models.
+While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models.
 Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
@@
 are naturally captured using quadratic constraints and objective
-functions. Several exact and heuristic optimization approaches have 
+functions. Several exact optimization approaches have 
 been proposed to solve
@@
 approaches represented the quadratic effects and constraints
-using linear models that approximate the real dependencies.
+using linear models that approximate the true dependencies.
 
@@
 
-In this part, we first present the
-physical model of a wind farm that we use throughout the paper. 
-Subsequently, we first provide an overview of 
+In this part, we start
+by presenting the
+physical model 
+of a wind farm that we use throughout the paper.
+Our wind farm model is based on
+the notation and physical description 
+from~\cite{Zhang2014}.  
+Subsequently, we provide a literature review of 
 existing mathematical
@@
 
-\subsection{Wind Farm Model}
+\subsection{The Wind Farm Model}
+\begin{figure}[t]
+	\centering
+	\includegraphics[scale = 0.9]{field_model.pdf}
 
-We consider a square shaped wind farm that we model 
-as an $N\times N$ grid
-with $N$ being the number of cells in each axis. 
-Each cell has the area of $c^2$ (in meters), i.e.,
-if $c=100$ meters and $N=20$, the farm area will be $2km\times 2km$.
+	\caption{Wind farm model: proximity constraints, wakes, wind direction.}\label{fig:field_model}
+\end{figure}
+
+\paragraph{Wind Farm as a Grid} We consider a square shaped wind farm that we model 
+as an $n\times n$ grid
+with $n$ being the number of cells on each axis. 
+The cells have an area of $c^2$ (in squared meters), i.e.,
+if $c=100$ meters and $n=20$, the farm area will be $2km\times 2km$.
 In this work, we assume that the terrain is flat and that 
 we have no constraints related to the terrain.
-Furthermore, we assume a sterile environment and do not consider the influence of noise
+Furthermore, we do not account for the influence of noise
 on the wind farm's surroundings. The wind farm literature treats the more complex
-cases of non-flat terrain and environmental considerations in~\cite{} and~\cite{}, respectively.
-In our work, we introduce a novel optimization approach that can be extended,
+farm models that capture non-flat terrain~\cite{song2015lazy,kuo2016wind}, multi-typed turbines~\cite{feng2017design}, noise effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex objectives (including turbine installation  and maintenance costs)~\cite{lackner2007analytical}.
+In this work, we focus on the novel quadratic approach for WFLO, and hence
+consider the most abstract wind farm model. Yet, our model can be extended,
 in the future, to include 
-such considerations.
+additional wind farm settings. \todo{perhaps discuss in future work}
 
-In principle, a higher number of turbines placed in the wind farm produces more energy. However,
-we must consider two types of interference between the turbines.
+Returning to the grid model, we assume that the number of homogeneous turbines is known 
+in advance; we denote the number of turbines by $m$. To place these $m$ turbines, 
+we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
+In what follows, we explain the two effects using a toy example of placing a single turbine on an 
+$8\times8$ grid (see Figure~\ref{fig:field_model}).
 
-First, 
-turbines cannot be placed at high proximity to each other. 
-Specifically, the turbines must be placed at least five rotor diameters apart from each other.
-Depending on the cell size, we shall formulate the corresponding constraints as part of the optimization model.
-For example, assuming that a turbine is placed
-at the center of the cell (cell size of $100$ meters) and for rotor diameter of $40$ meters, 
-we must ensure that two turbines will not placed in adjacent cells.
+\paragraph{Proximity Constraints} Turbines must be placed at least \emph{five rotor diameters} apart. Depending on the cell size, 
+	we shall formulate the corresponding constraint as part of the optimization model. For example, assuming that a turbine is placed
+	at the center of the cell (cell size of $100$ meters) and given that the rotor diameter is $40$ meters, 
+	we must ensure that two turbines will not placed in adjacent cells. In Figure~\ref{fig:field_model}, 
+	the red area around the turbine represents the cells in which we cannot place an additional turbine.
 
-Second, turbines influence each other through an effect called \emph{wake}. Wakes cause reduction of the effective power produced by a turbine due to upstream turbines changing airflow dynamics.
+\paragraph{Wakes} Turbines influence each other through interference effects that are referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}. 
+	 Wakes cause reduction of the effective power produced by a turbine due to upstream turbines that change
+	 its airflow dynamics. In Figure~\ref{fig:field_model}, the wind is assumed to be horizontal. Therefore, the wake effect due
+	 to placing the turbine expands as shown in the diagram: it will influence only turbines that will be placed in the orange cells.
+	 In other words, if a turbine is placed in the orange area, its effective
+	 power is reduced due to the upstream turbine. Moreover, wakes can occur between multiple turbines, and hence
+	 multiple wakes must be taken into account for each location when modeling the farm. 
+	 The reduction in power due to the wake effect is determined by three parameters: (1) the distance between the upstream
+	 turbine and the cell for which we want to compute the effect (denoted by $x$ in Figure~\ref{fig:field_model}), (2) the radius of the cone's opening 
+	 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) 
+	 the superposition of the wind regimes that correspond to the farm's location.
+	 This reduction effect can be pre-computed for each pair of cells using known
+	 turbine specification and terrain conditions together with probabilistic assumptions on wind regimes~\cite{Zhang2014}.
+Note that turbines can adjust their direction according to the current wind direction.
 
-\todo{Explain the wake model - single wake and linear superposition.
-Explain that it is less accurate but leads to a quadratic objective.
-We then assess it using sum of squares.
-Explain that there are $m$ turbines to be placed.
-use Zheng and use Mosetti.}
+\paragraph{Effective Power Calculations in the Presence of Wakes} 
+Let $\mathcal{D}$ be the set of wind regimes. Every regime $d$ is a combination of wind direction and free speed. 
+The probability of wind regime 
+$d$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
+we provide one example of a probability distribution 
+function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
+a regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind blowing at the velocity $12km/h$ and the
+probability of this regime is given by $0.008$.
+Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.
+Turbines can turn in alignment with the direction of the current wind regime.
 
-\todo{Wind direction}
+Next, we write 
+the expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}): \begin{equation} E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d. \end{equation}
+Given that we must consider wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of current wind regime.
+The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
+\begin{equation}
+u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
+\end{equation} with $u_{id\infty}$ being the free wind speed (second component of regime $d$), the set of turbines $\mathcal{U}_{id}$ that are upstream to $i$ 
+under wind regime $d$, and $u_{ijd}$ being the wake 
+reduced speed at $i$ due to upstream turbine $j$. 
+The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$.
+Furthermore, the set of upstream turbines, $\mathcal{U}_{id}$ is computed
+by rotating the grid w.r.t.
+to the current wind regime $d \in \mathcal{D}$ and computing which turbine $j$ is located upstream to turbine $i$ given $d$.
+The expression in Eq.
+(\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for
+the total expected power in presence of wakes~\cite{Zhang2014}. This is also known to be the most accurate analytical expression account for multiple wakes.
+
+An alternative 
+form for the expected power in the presence of multiple wakes is the linear superposition (LS) model: \begin{equation} \label{eq:ls}
+E_{LS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \Bigg(\frac{1}{3}u_{id\infty}^3 -\sum_{j\in\mathcal{U}_{id}} \frac{1}{3}(u_{id\infty}^3 - u_{ijd}^3)\Bigg).
+\end{equation} Note that $E_{LS}$ has the same parameters as $E_{SS}$; the difference is in the function that ties these parameters together. 
+The LS model is known to be less accurate compared to SS, yet as we will show in Section it 
+results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.
 
@@
 
-The allocation of turbines in a wind farm was first considered 
-as an optimization problem in~\cite{MOSETTI1994105}. Since
-then, WFLO has caught the attention of optimization researchers. 
-The WFLO problem can either be solved \emph{exactly} using
-techniques from operations research (e.g., linear programming, 
-mixed-integer programming, stochastic programming, and constraint programming)\cite{Zhang2014,turner2014new}
-or solved \emph{approximately} using various meta-heuristics (e.g., genetic algorithms and local search) \cite{MOSETTI1994105}.
-The main advantage of the former techniques
-is that they provide guarantees on solution quality, 
-while the latter methods generate 
-quick and sufficiently accurate solutions. In this work, we compare 
-exact and approximate methods for WFLO. Both use 
-quadratic optimization models... \todo{add more variations and explain that we
-	focus on one to demonstrate the usefulness of quadratic models}
+The allocation of turbines in a grid-like wind farm was first considered 
+as an optimization problem in~\cite{MOSETTI1994105}. The proposed method in~\cite{MOSETTI1994105}
+searches for improved solutions using a 
+genetic approach~\cite{davis1991handbook},
+which is an approximate technique 
+that does not aim to provide an optimal solutions,
+but rather attempts to find a quick solution to WFLO. 
+In follow-up papers, 
+the WFLO problem was solved using \emph{exact}  
+optimization methodologies~\cite{turner2014new,Zhang2014}. 
+These methods guarantee that the returned solution is optimal with respect
+to the specified objective (i.e., maximize total expected power). 
+Indeed, the exact methods were shown to yield higher energy values, yet they often suffer from 
+high computational cost and long runtimes. Therefore, one must strike a balance between optimality and computational requirements of the solution.  
+Below, we provide an overview of the literature of the two approaches (exact vs. approximate) and relate our 
+work to both through quadratic optimization models\footnote{We shall review 
+only literature that considers the same wind farm model as the one presented earlier in this section}.
+
+\paragraph{Exact Optimization Methods} 
+\todo{Rephrase and add citations to exact methods}
+The WFLO problem can be solved to optimality 
+to optimality, using
+techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming).
+These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
+the best solution one can obtain for the given problem. For the wind farm model
+presented in the beginning of this section,
+the two dominant works that apply exact optimization 
+are~\cite{Zhang2014}, and~\cite{turner2014new}. 
+The former builds on~\cite{donovan2005wind} and solves a MILP model for the linear superposition (LS) objective
+and use constrained programming to find a solution for the sum of squares (SS) objective.
+The latter, uses a quadratic and linear approximations to the SS objective.
+Since the SS objective is considered to be more accurate, solutions that come from LS-based optimization
+are plugged into the SS objective.
+
+The main advantage of exact techniques
+is that they provide guarantees on solution quality. However, their disadvantage is
+that they may take a very long time to execute before finding an optimal solution. 
+In fact, we shall show that the WFLO problem is computationally hard, a result
+that we have not yet encountered in the WFLO literature. Therefore,
+when multiple designs are considered and a fast solution is of essence,
+one may turn to approximate WFLO solutions.
+
+\paragraph{Approximate Methods} \todo{tighten} In contrast to exact methods, approximation algorithms often provide   
+quick and sufficiently accurate solutions. Several works solved 
+WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. These methods use two principles (cross-over and mutation)
+to change existing solutions and improve them over time.
+Evolutionary methods do not guarantee termination (i.e., they may run forever), and when they do terminate (in practical cases).
+Hence, one must often
+introduce stopping criteria and thus compromise on the quality of the attained solution~\cite{davis1991handbook}.
+Furthermore, there is no natural way of introducing constraints into genetic methods, which
+leads to a feasibility check for every solution.
+
+The first attempt to address these limitations 
+was made in a local search approach~\cite{ozturk2004heuristic}. 
+The solution still does not guarantee optimality, 
+yet it circumvents the termination and feasibility checking issues of the evolutionary
+solutions. The local search attempts to apply either move, remove or add actions given an incumbent 
+feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
+when a stopping criteria is reached.
+The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
+that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution far
+from the optimal one.
+This was experimentally shown in~\cite{rivas2009solving}.
+To overcome this limitation, the authors of~\cite{rivas2009solving} use simulated annealing (SA), which is a neighborhood search method that allows visits to infeasible solutions.
+This creates a well-connected solution space that allows an escape for local optima, and was shown to be superior to~\cite{ozturk2004heuristic} and to the genetic method proposed in~\cite{grady2005placement}.
+A drawback of the SA approach in~\cite{rivas2009solving} is its ad-hoc nature.
+The various model components must be tailored and geared towards WFLO.
+This means that small adjustments to our WFLO setting
+ (e.g., adding noise considerations), would require major changes to the SA solution. 
+Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.
+
+In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
+our methodology is generic since it is based on quadratic programming models.
+These models can be easily adjusted to additional constraints and components of the objective function.
+In addition, formulating the problem using quadratic programming  enables us to solve WFLO using advanced software exact optimization solvers (e.g., Gurobi optimizer) and approximate hardware solutions (e.g., Fujitsu's digital annealer for solving quadratic problems using SA).
+
+\paragraph{Quadratic Models for Exact and Approximate WFLO}
+\todo{send a clear message here. Perhaps move to next section}.
+
+A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
+approximate methods and found that the two perform in a comparable fashion, without large differences between
+the attained energy~\cite{yang2019simulated}. Exact methods tend to perform better
+in smaller WFLO benchmarks, while simulated annealing methods had slightly dominating results
+for larger benchmarks. This leads to the conclusion that the two types of methods should be considered.
+
+In this work, we use a common optimization model to solve WFLO both exactly and approximately, thus enjoying the  best of both worlds. 
+The common optimization model is a quadratic programming representation of WFLO, 
+which attempts to minimize a quadratic energy function with or without considering additional constraints.
+Specifically, we consider two variations of the quadratic model. 
+The first one is referred to as quadratic constrained optimization problems (QCOP). 
+It enables maximizing a quadratic (LS-based) energy expression, while 
+ensuring that WFLO constraints are maintained (turbine proximity and total number of turbines is $m$). 
+The second model is a quadratic unconstrained binary optimization (QUBO), 
+which
+
+On the one hand, exact solvers for quadratic programs (QCOP and QUBO) such as Gurobi~\cite{}, 
+use cutting edge
+operations research methods to enhance performance.
+On the other hand, simulated annealing methods have been recently embedded onto 
+designated processing units. These units use quantum-inspired technology to solve
+complex problems quickly and with minimal degradation in terms of the resulting solution~\cite{}.
+Specifically, in this work, we shall use Fujitsu's digital annealer (DA) to solve WFLO. The DA
+also requires a quadratic representation of the WFLO problem.
+
+Both of these methods use 
+quadratic optimization models, which we shall present in the next section (Section~\ref{sec:QUBO4WFLO}).
 
@@
 we formulate the WFLO problem
-as two quadratic optimization problems: a
+using quadratic programming. Specifically, we consider two
+quadratic optimization problems that represent WFLO: a
 quadratic constrained optimization (QCOP) problem and 
 a quadratic unconstrained binary optimization (QUBO) problem. 
-To this end,
-we follow standard WFLO definitions
-and notation presented in~\cite{Zhang2014}. First, we present WFLO's inputs and decision variables.
-Subsequently, we present the mathematical formulation of the two problems.
+The former represents WFLO characteristics (proximity and number of turbines)
+as hard constraints, while the latter considers these characteristics as soft constraints that are added to its objective function.
 
-\paragraph{WFLO Inputs and Decision Variables} Let $x_i$ be a binary decision variable that represents whether a 
-turbine is positioned at location $i \in \mathcal{N}$ with $\mathcal{N}$ being the set 
-of $n$ possible locations and 
-$m$ being the total number of required turbines. Note that we assume a discrete grid with 
-a finite number of possible positions. We 
-write $x$ as shorthand for $x = (x_1,\ldots, x_n)$. Furthermore,
+First, we present WFLO's inputs and decision variables.
+Subsequently, we present the two quadratic formulations of WFLO
+and discuss the usefulness of each formulation in the context
+of advanced optimization software and hardware.
+
+\subsection{WFLO Inputs and Decision Variables} We consider the wind farm model that we presented in Section~\ref{sec:related}) and follow the notation from~\cite{Zhang2014}.
+Let $x_i$ be a binary decision variable that represents whether a
+turbine is positioned at location 
+$i \in \mathcal{N}$ with $\mathcal{N}$ being the set 
+of $k = n^2$ possible turbine locations (cells in the grid). We 
+write $x$ as shorthand for $x = (x_1,\ldots, x_k)$. Furthermore,
 we denote by $\mathcal{E} \subseteq \mathcal{N}\times \mathcal{N}$
 the set of location pairs that cannot simultaneously host turbines 
-due to proximity constraints. 
-Let $u_{id}$ and $u_{id, \infty} $ 
-be the wind speeds at turbine location $i$ for wind regime $d \in \mathcal{D}$
-with and without interference from other turbines due to wake effects, respectively. A wind regime is a combination 
-of wind speed and direction. 
-The probability of wind regime 
-$d$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. For example, 
-regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind.
+due to proximity constraints. The set $\mathcal{E}$ can be pre-computed 
+based on problem specifications. For example,
+it is common to assume that two turbines must be placed at least five rotor diameters apart.
+
+Let $u_{id}$ and $u_{id\infty}$ 
+be the wind speeds at turbine location $i \in \mathcal{N}$ for wind regime $d \in \mathcal{D}$
+with and without interference from other turbines due to wake effects, respectively. 
+Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.
 
@@
 the set of downstream 
-turbine locations 
-given that a turbine is placed at cell $i$ for wind regime $d$ (i.e., turbines placed in these locations 
-will be influenced by a turbine placed at $i$). Finally,
-we denote by $u_{ijd}$ the wind speed at turbine $j$ due to a single wake from upstream turbine $i$ with $j \in \mathcal{U}_{id}$. 
-The following objective function  
+turbine locations for wind regime $d$
+given that a turbine is placed at cell $i$ (i.e., turbines placed in these locations 
+will be influenced by a turbine placed in $i$). Finally,
+we denote by $u_{ijd}$ the wind speed at location $j$ due to a single wake from upstream turbine at location $i$ with $j \in \mathcal{U}_{id}$. 
+The following function  
 represents the total expected energy of a wind farm given placement decisions 
 $x$: \begin{equation}
-f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \label{ObjFunc}\\
+f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j).
+\label{ObjFunc}\\
 %&s.t.:& \nonumber\\
@@
 %&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
-\end{equation}
+\end{equation} The objective function corresponds to the linear superposition (LS)
+expression (Eq.~\ref{eq:ls}). We are now ready to provide the two quadratic programs to represent WFLO.
 
@@
 To 
-maximize $f(x)$ and satisfy the number of turbines and proximity 
-constraints we can 
+maximize $f(x)$ and satisfy the total number of turbines and proximity 
+constraints we write
 the following quadratic optimization problem (QCOP):
@@
 &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
-\end{eqnarray}
-
-Solving QCOP with Gurobi. Explain.
+\end{eqnarray} The QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{}. 
+So far, in~\cite{Zhang2014} and~\cite{donovan2005wind}, equivalent (linear) problems were formulated and solved using 
+exact solvers. However, these solvers should only be used if the problem is computationally hard.
+Prior to this work, we are unaware of a complexity result that analyzes LS-based WFLO. In what follows, we provide
+a theorem that proves the computational complexity of the QCOP. 
+\begin{mythm}
+	The computational complexity of LS-based WFLO is $\mathcal{NP}$-hard. 
+\end{mythm}
+\begin{myproof}
+We prove by showing that a special case of the WFLO problem (Eq.~(\ref{QCOP}) is
+$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
+are dropped and we get the HSP, which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
+	\end{myproof} To overcome this limitation associated with the computational hardness of WFLO,
+we reformulate the QCOP into a QUBO problem. This allows us to use a specialized optimization hardware that uses
+simulated annealing in order to solve QUBO problems quickly.
 
@@
 
-The equivalent QUBO formulation is given by,
+Quadratic unconstrained binary optimization formulations, as their name implies, contain only an objective function that 
+one wishes to maximize (or minimize). Furthermore, the decision variables must be binary. Let $\lambda = (\lambda_1,\lambda_2)$ 
+be a vector of constraints parameters with $\lambda_i >0$. The equivalent QUBO formulation of QCOP is given by,
 \begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
-with the term multiplied by $\lambda_1$ 
-ensures that exactly $m$ turbines are placed, while the term multiplied by $\lambda_2$ penalizes
-solutions that place turbines in adjacent locations.
-The two penalty terms $\lambda_1$ and $\lambda_2$ set large enough to ensure constraint satisfaction (see Section~\ref{sec:theory}).
-The QUBO in Eq.
-(\ref{QUBO}) allows us
-to solve the WFLO problem using quantum-inspired hardware.
+with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
+penalizing solutions with more or less than $m$ turbines, while the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizes
+solutions that place turbines in high proximity. The two penalty terms $\lambda_1$ and $\lambda_2$ 
+must be large enough to ensure that the solutions are indeed feasible. These types of constraints
+are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
+for an arbitrary selection of $\lambda$.
 
@@
 \paragraph{WFLO Instances}
+\begin{figure}[t]
+	\centering
+	\includegraphics[scale = 0.3]{prob_wind_pdf.pdf}
+
+	\caption{Wind regimes: x-axis is the angle (times 10 degrees), y-axis is the probability for wind regime, color corresponds to free wind speed.}\label{fig:prob_wind}
+\end{figure}
 
@@
 \subsection{Main Results}
+
+\begin{table}[ht]
+
+	\begin{tabular}{| c | c | c | c | c | c | c | c |}
+		\toprule
+		Wind Directions  & n        & m        & ILP-LS       & QP-LS       & QUBO-LS       & QP-SS       & DA-LS     \\
+		\toprule
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
+			& & 30   & 15739.57 & 15738.45 & \textbf{15742.93} & 15735.1  & 15740.69         \\
+		& & 40 & \textbf{19265.21} & \textbf{19265.21} & 18952.97 & \textbf{19265.21} & \textbf{19265.21}                \\
+				\cline{2-8}
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}           \\
+		&&30   & \textbf{16770.2} & 16758.45 & 16715.19 & 16755.75 & 16570.96                   \\
+		&&40   & \textbf{21977.17} & 21963.8 & 21643.56 & \textbf{21977.17} & 21874.04                     \\
+		\hline
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & 19022.77 & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & \textbf{19221.44} \\
+		&& 30   & 27247.5  & 27442.9  & 27418.74 & 27387.06 & \textbf{27450.99}                  \\
+		&&40   & 34671.62 & \textbf{34938.65} & \textbf{34938.65} & 34817.27 & 34888.81                  \\
+		\cline{2-8}
+		&  \multirow{3}{*}{20}   & 20       & 18448.64 & 19336.33 & 19172.56 & 19354.23 & \textbf{19441.17}          \\
+		&&30   & 26367.2  & 27752 & 27785 & 27492.11 & \textbf{27975.55}                     \\
+		&&40   & 33451.94 & \textbf{35665.6}  & 35127.32 & 35065.85 & 35604.45  \\
+		\bottomrule                   
+	\end{tabular}
+
+\vspace{0.5em}
+\caption{Total Expected Power (kilo Watt/year) after 3600 seconds as function of benchmark and method}\label{tab:results1}
+\end{table}
+
+\begin{figure}[t!]%
+
+	\subfloat[MRE after 50 seconds]{{\includegraphics[width=0.5\textwidth]{energy_50.png} }}%
+	\quad
+	\subfloat[MRE after 1 hour]{{\includegraphics[width=0.5\textwidth]{energy_3600.png} }}\\%\\
+	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime.}%
+
+	\label{fig:main_effects}%
+\end{figure}
 

### 2020-04-08 paper.tex
commit.

@@
 The QCOP can be solved using state-of-the-art solvers that
-specialize in such problems, e.g., Gurobi~\cite{guorbi}. 
+specialize in such problems, e.g., Gurobi~\cite{gurobi}. 
 Our second model is based on a quadratic unconstrained binary optimization (QUBO) representation of the WFLO 
@@
 This enables the use of specialized optimization hardware tailored to solve
-QUBOs. In this paper, we use Fujitsu's digital annealer to demonstrate the strength
-of the approach. The main contributions of this paper are: 
+QUBOs. In this paper, we use Fujitsu's digital annealer, a simulated annealing based QUBO solver, to demonstrate the strength
+of our approach. The main contributions of this paper are: 
 \begin{enumerate} 
-\item We propose two novel quadratic WFLO models that capture wake and proximity
-effects directly without going through a linear approximation, and, 
-\item We test the two formulations
+\item We propose a novel quadratic modeling framework for WFLO that captures wake 
+effects directly through their linear superposition 
+without the need for approximations and ad-hoc solutions,
+\item We provide the computational complexity of WFLO and show that it is $\mathcal{NP}$-hard, and, 
+\item We test our quadratic formulations
 on a commerical software-based quadratic solver (Gurobi) and on Fujitsu's digital 
@@
 existing optimization methods. To this end,
-we solve 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new}). 
-We show, empirically, that solving quadratic WFLO models
+we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
+and show that solving quadratic WFLO models
 using Gurobi and the DA
@@
 we observe that the digital 
-annealer often provides order of magnitude faster feasible solutions
-compared to our QCOP based approach.
+annealer often provides order of magnitude faster solutions
+compared to software-based approaches.
 
-The remainder is structured as follows \todo{adjust structure}.
+The remainder is structured as follows \todo{adjust structure according to final structure of paper}.
 
@@
 
-In this part, we start
+We start this section
 by presenting the
 physical model 
-of a wind farm that we use throughout the paper.
+of a wind farm, which we use throughout the paper.
 Our wind farm model is based on
-the notation and physical description 
-from~\cite{Zhang2014}.  
+the common notation and description 
+presented in~\cite{Zhang2014}.  
 Subsequently, we provide a literature review of 
 existing mathematical
-optimization methods for WFLO, focusing
+optimization methods for solving WFLO, focusing
 on their similarities and differences
@@
 
-Returning to the grid model, we assume that the number of homogeneous turbines is known 
+Returning to the grid model, we assume that the number of homogeneous turbines to be placed is known 
 in advance; we denote the number of turbines by $m$. To place these $m$ turbines, 
@@
 \paragraph{Proximity Constraints} Turbines must be placed at least \emph{five rotor diameters} apart. Depending on the cell size, 
-	we shall formulate the corresponding constraint as part of the optimization model. For example, assuming that a turbine is placed
-	at the center of the cell (cell size of $100$ meters) and given that the rotor diameter is $40$ meters, 
-	we must ensure that two turbines will not placed in adjacent cells. In Figure~\ref{fig:field_model}, 
-	the red area around the turbine represents the cells in which we cannot place an additional turbine.
+	we shall formulate the corresponding constraint as part of the optimization model. For example, assume that a turbine is placed
+	in the center of a cell. Then, given a cell size of $c = 100$ meters and a rotor diameter of $40$ meters, 
+	we must ensure that two turbines will not placed inside two adjacent cells. In Figure~\ref{fig:field_model}, 
+	the red area around the turbine represents the forbidden cells.
 
@@
 	 In other words, if a turbine is placed in the orange area, its effective
-	 power is reduced due to the upstream turbine. Moreover, wakes can occur between multiple turbines, and hence
-	 multiple wakes must be taken into account for each location when modeling the farm. 
-	 The reduction in power due to the wake effect is determined by three parameters: (1) the distance between the upstream
+	 power is reduced due to the upstream turbine that we placed in the $8\times8$ farm. 
+	 Moreover, wakes can occur between multiple turbines, and hence
+	 multiple wakes must be taken into account for each location. 
+	 The reduction of power due to the wake effect is determined by three parameters: (1) the distance between the upstream
 	 turbine and the cell for which we want to compute the effect (denoted by $x$ in Figure~\ref{fig:field_model}), (2) the radius of the cone's opening 
 	 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) 
-	 the superposition of the wind regimes that correspond to the farm's location.
-	 This reduction effect can be pre-computed for each pair of cells using known
-	 turbine specification and terrain conditions together with probabilistic assumptions on wind regimes~\cite{Zhang2014}.
-Note that turbines can adjust their direction according to the current wind direction.
+	 the superposition of several wind regimes (a wind regime is a combination of free speed and directions).
+	 The wake effect can be pre-computed for each pair of cells 
+	 since we know: the distance ($x$), the turbine specifications and terrain conditions (that jointly dictate $r$), and  
+	 probabilistic behavior of wind regimes~\cite{Zhang2014}.
 
 \paragraph{Effective Power Calculations in the Presence of Wakes} 
-Let $\mathcal{D}$ be the set of wind regimes. Every regime $d$ is a combination of wind direction and free speed. 
+Let $\mathcal{D}$ be the set of wind regimes. Every regime $d\in \mathcal{D}$ is a unique combination of wind direction and free speed. 
 The probability of wind regime 
-$d$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
+$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
 we provide one example of a probability distribution 
 function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
-a regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind blowing at the velocity $12km/h$ and the
-probability of this regime is given by $0.008$.
-Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.
-Turbines can turn in alignment with the direction of the current wind regime.
+a regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind blowing at a free velocity of $12km/h$. The
+probability of this regime is $0.008$. Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
+Note that turbines can adjust their direction according to the current wind 
+regime.
 
-Next, we write 
-the expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}): \begin{equation} E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d. \end{equation}
-Given that we must consider wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of current wind regime.
+Given that we must consider wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
 The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
@@
 (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for
-the total expected power in presence of wakes~\cite{Zhang2014}. This is also known to be the most accurate analytical expression account for multiple wakes.
+the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
+the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}): \begin{equation} E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.
+\end{equation} This is also known to be the most accurate analytical total power expression that accounts for multiple wakes.
+However, the expression is not amenable for optimization, and hence, a simplified representation for the total expected energy that considers wakes is required.
 
 An alternative 
-form for the expected power in the presence of multiple wakes is the linear superposition (LS) model: \begin{equation} \label{eq:ls}
+form of the total expected power in the presence of multiple wakes is the linear superposition (LS) expression: \begin{equation} \label{eq:ls}
 E_{LS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \Bigg(\frac{1}{3}u_{id\infty}^3 -\sum_{j\in\mathcal{U}_{id}} \frac{1}{3}(u_{id\infty}^3 - u_{ijd}^3)\Bigg).
 \end{equation} Note that $E_{LS}$ has the same parameters as $E_{SS}$; the difference is in the function that ties these parameters together. 
-The LS model is known to be less accurate compared to SS, yet as we will show in Section it 
+The LS model is known to be less accurate compared to SS, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
 results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.
@@
 which is an approximate technique 
-that does not aim to provide an optimal solutions,
-but rather attempts to find a quick solution to WFLO. 
-In follow-up papers, 
+that does not aim to provide an optimal solutions;
+instead, it attempts to find a quick turbine placement.
+
+In follow-up works, 
 the WFLO problem was solved using \emph{exact}  
@@
 high computational cost and long runtimes. Therefore, one must strike a balance between optimality and computational requirements of the solution. 
-Below, we provide an overview of the literature of the two approaches (exact vs. approximate) and relate our 
-work to both through quadratic optimization models\footnote{We shall review 
-only literature that considers the same wind farm model as the one presented earlier in this section}.
+Fast solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various locations),
+while optimal solutions yield better placements in terms of the total expected energy.
+
+In what follows, we provide a literature review that covers both approaches, namely exact vs. approximate WFLO.
+We subsequently relate our 
+work to both exact and approximate methods via our proposed quadratic optimization paradigm for WFLO\footnote{We shall review 
+only literature that considers the same wind farm model as the one presented at the beginning of this section}.
 
@@
 \todo{Rephrase and add citations to exact methods}
-The WFLO problem can be solved to optimality 
+The WFLO problem can be solved  
 to optimality, using
-techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming).
+techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{}.
 These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
-the best solution one can obtain for the given problem. For the wind farm model
-presented in the beginning of this section,
-the two dominant works that apply exact optimization 
-are~\cite{Zhang2014}, and~\cite{turner2014new}. 
-The former builds on~\cite{donovan2005wind} and solves a MILP model for the linear superposition (LS) objective
-and use constrained programming to find a solution for the sum of squares (SS) objective.
-The latter, uses a quadratic and linear approximations to the SS objective.
-Since the SS objective is considered to be more accurate, solutions that come from LS-based optimization
-are plugged into the SS objective.
+the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
+MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
+approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
+compared to existing approximate and exact methods~\cite{turner2014new}.
 
-The main advantage of exact techniques
+The most recent attempt to solve the wind farm variation that we presented 
+in this section is given in~\cite{Zhang2014}. The authors compared
+previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
+is an optimization technique that does not make assumptions on the functional form of the objective.
+This turned out to be computationally intractable for the larger standard WFLO benchmarks.
+
+In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
+model WFLO using a more natural quadratic representation. The difference between our approach and the quadratic model
+proposed in~\cite{turner2014new} is that our model is not an approximation, but rather an exact representation of the LS objective.
+
+As we mentioned before, the main advantage of exact techniques
 is that they provide guarantees on solution quality. However, their disadvantage is
 that they may take a very long time to execute before finding an optimal solution. 
-In fact, we shall show that the WFLO problem is computationally hard, a result
+In fact, Section~\ref{sec:QUBO4WFLO} shows that WFLO is computationally hard, a result
 that we have not yet encountered in the WFLO literature. Therefore,
-when multiple designs are considered and a fast solution is of essence,
-one may turn to approximate WFLO solutions.
+when one considers solving multiple wind farm design problems, a fast solution is of essence,
+and hence one may turn to approximate WFLO.
 
@@
 quick and sufficiently accurate solutions. Several works solved 
-WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. These methods use two principles (cross-over and mutation)
-to change existing solutions and improve them over time.
-Evolutionary methods do not guarantee termination (i.e., they may run forever), and when they do terminate (in practical cases).
-Hence, one must often
-introduce stopping criteria and thus compromise on the quality of the attained solution~\cite{davis1991handbook}.
+WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
+These methods use two principles (cross-over and mutation)
+to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
+(i.e., they may run forever without achieving a quality criteria), and when they do terminate (in practical cases) they do
+not guarantee optimal solutions. Hence, one must often
+introduce non-quality related stopping criteria, 
+thus compromising on quality of the attained solution~\cite{davis1991handbook}.
 Furthermore, there is no natural way of introducing constraints into genetic methods, which
-leads to a feasibility check for every solution.
+leads to a potentially costly feasibility check for every solution.
 
 The first attempt to address these limitations 
-was made in a local search approach~\cite{ozturk2004heuristic}. 
-The solution still does not guarantee optimality, 
+was made in a local search approach for WFLO~\cite{ozturk2004heuristic}. 
+The solution does not guarantee optimality, 
 yet it circumvents the termination and feasibility checking issues of the evolutionary
-solutions. The local search attempts to apply either move, remove or add actions given an incumbent 
+methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
 feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
@@
 The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
-that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution far
-from the optimal one.
-This was experimentally shown in~\cite{rivas2009solving}.
-To overcome this limitation, the authors of~\cite{rivas2009solving} use simulated annealing (SA), which is a neighborhood search method that allows visits to infeasible solutions.
-This creates a well-connected solution space that allows an escape for local optima, and was shown to be superior to~\cite{ozturk2004heuristic} and to the genetic method proposed in~\cite{grady2005placement}.
-A drawback of the SA approach in~\cite{rivas2009solving} is its ad-hoc nature.
-The various model components must be tailored and geared towards WFLO.
-This means that small adjustments to our WFLO setting
+that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
+that can potentially by far
+from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.
+
+To overcome this limitation, the authors of~\cite{rivas2009solving} used simulated annealing (SA), 
+a neighborhood search method that allows visits to infeasible solutions. 
+This creates a well-connected solution space that leads the algorithm to escape 
+local optima. The SA method was shown to be superior to~\cite{ozturk2004heuristic} 
+and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach 
+is its ad-hoc nature. The various components of the algorithm
+ must be tailored and geared towards WFLO. This means that small adjustments to our WFLO setting
  (e.g., adding noise considerations), would require major changes to the SA solution. 
@@
 In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
-our methodology is generic since it is based on quadratic programming models.
-These models can be easily adjusted to additional constraints and components of the objective function.
-In addition, formulating the problem using quadratic programming  enables us to solve WFLO using advanced software exact optimization solvers (e.g., Gurobi optimizer) and approximate hardware solutions (e.g., Fujitsu's digital annealer for solving quadratic problems using SA).
+our methodology is generic since it is based on our quadratic paradigm and
+on the ability of specialized hardware to solve quadratic (non-ad-hoc) models.
+Quadratic WFLO models can be easily adjusted to include additional problem features, e.g., wind farm characteristics.
 
-\paragraph{Quadratic Models for Exact and Approximate WFLO}
+\paragraph{Quadratic Programming as a Unified Framework for Exact and Approximate WFLO}
 \todo{send a clear message here. Perhaps move to next section}.
 
+In this work, we argue that 
+formulating WFLO using quadratic programming enables us to employ 
+advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
+(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
 A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
 approximate methods and found that the two perform in a comparable fashion, without large differences between
-the attained energy~\cite{yang2019simulated}. Exact methods tend to perform better
-in smaller WFLO benchmarks, while simulated annealing methods had slightly dominating results
-for larger benchmarks. This leads to the conclusion that the two types of methods should be considered.
+the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
+in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
+This leads to the conclusion that both types of methods (exact and approximate) should be considered. 
+Therefore, in this work, we use a unified paradigm that would enable both exact and approximate solutions to WFLO. Thus, 
+our approach `enjoys' the best of both worlds.
 
-In this work, we use a common optimization model to solve WFLO both exactly and approximately, thus enjoying the  best of both worlds. 
-The common optimization model is a quadratic programming representation of WFLO, 
-which attempts to minimize a quadratic energy function with or without considering additional constraints.
-Specifically, we consider two variations of the quadratic model. 
-The first one is referred to as quadratic constrained optimization problems (QCOP). 
-It enables maximizing a quadratic (LS-based) energy expression, while 
-ensuring that WFLO constraints are maintained (turbine proximity and total number of turbines is $m$). 
-The second model is a quadratic unconstrained binary optimization (QUBO), 
-which
-
-On the one hand, exact solvers for quadratic programs (QCOP and QUBO) such as Gurobi~\cite{}, 
-use cutting edge
-operations research methods to enhance performance.
-On the other hand, simulated annealing methods have been recently embedded onto 
-designated processing units. These units use quantum-inspired technology to solve
-complex problems quickly and with minimal degradation in terms of the resulting solution~\cite{}.
-Specifically, in this work, we shall use Fujitsu's digital annealer (DA) to solve WFLO. The DA
-also requires a quadratic representation of the WFLO problem.
-
-Both of these methods use 
-quadratic optimization models, which we shall present in the next section (Section~\ref{sec:QUBO4WFLO}).
+%
+%
+%Specifically, we consider two variations of the quadratic programming paradigm.  
+%The first one is referred to as quadratic constrained optimization problems (QCOP). 
+%It enables maximizing a quadratic (LS-based) energy expression, while 
+%ensuring that WFLO constraints are maintained (turbine proximity and total number of turbines is $m$). 
+%The second model is a quadratic unconstrained binary optimization (QUBO), 
+%which 
+%
+%
+%On the one hand, exact solvers for quadratic programs (QCOP and QUBO) such as Gurobi~\cite{}, 
+%use cutting edge
+%operations research methods to enhance performance.
+%On the other hand, simulated annealing methods have been recently embedded onto 
+%designated processing units. These units use quantum-inspired technology to solve
+%complex problems quickly and with minimal degradation in terms of the resulting solution~\cite{}.
+%Specifically, in this work, we shall use Fujitsu's digital annealer (DA) to solve WFLO. The DA
+%also requires a quadratic representation of the WFLO problem. 
+%
+%Both of these methods use 
+%quadratic optimization models, which we shall present in the next section (Section~\ref{sec:QUBO4WFLO}).
+%
 
@@
 
-First, we present WFLO's inputs and decision variables.
-Subsequently, we present the two quadratic formulations of WFLO
-and discuss the usefulness of each formulation in the context
-of advanced optimization software and hardware.
+We start by presenting the inputs and decision variables into our optimization approaches.
+Subsequently, we introduce the two quadratic formulations of WFLO and 
+prove its computational complexity. Lastly, we discuss the details of existing
+software and hardware solutions that can be used to solve the two formulations.
 
-\subsection{WFLO Inputs and Decision Variables} We consider the wind farm model that we presented in Section~\ref{sec:related}) and follow the notation from~\cite{Zhang2014}.
+\subsection{WFLO Inputs and Decision Variables} We consider the wind farm model that we presented in Section~\ref{sec:related}. 
 Let $x_i$ be a binary decision variable that represents whether a 
@@
 Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.
-
-Next, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
-the set of downstream 
+Further, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
+the set of upstream 
 turbine locations for wind regime $d$
@@
 &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
-\end{eqnarray} The QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{}. 
+\end{eqnarray} Constraints~\ref{Cardinality} ensure that exactly $m$ turbines are placed on the grid, while~\ref{Grid}
+enforce proximity constraints between the turbines.
+
+The QCOP presented as Problem~\ref{QCOP} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
 So far, in~\cite{Zhang2014} and~\cite{donovan2005wind}, equivalent (linear) problems were formulated and solved using 
 exact solvers. However, these solvers should only be used if the problem is computationally hard.
-Prior to this work, we are unaware of a complexity result that analyzes LS-based WFLO. In what follows, we provide
-a theorem that proves the computational complexity of the QCOP. 
+We are unaware of an existing complexity result for the LS variation in Eq.~\ref{QCOP}. In what follows, we provide
+a theorem that proves the computational complexity of WFLO. 
 \begin{mythm}
-	The computational complexity of LS-based WFLO is $\mathcal{NP}$-hard. 
+	The computational complexity of WFLO is $\mathcal{NP}$-hard. 
 \end{mythm}
@@
 $\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
-are dropped and we get the HSP, which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
-	\end{myproof} To overcome this limitation associated with the computational hardness of WFLO,
-we reformulate the QCOP into a QUBO problem. This allows us to use a specialized optimization hardware that uses
-simulated annealing in order to solve QUBO problems quickly.
+are dropped and we get the heaviest k-subgraph problem (HSP), which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
+Since a special case of WFLO is  $\mathcal{NP}$-hard it implies that WFLO is at least $\mathcal{NP}$-hard. On the other hand,
+WFLO is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFLO is $\mathcal{NP}$-hard.
+	\end{myproof} To overcome the challenge associated with the computational hardness of WFLO,
+we reformulate the QCOP into a QUBO problem.
 
@@
 
-Quadratic unconstrained binary optimization formulations, as their name implies, contain only an objective function that 
-one wishes to maximize (or minimize). Furthermore, the decision variables must be binary. Let $\lambda = (\lambda_1,\lambda_2)$ 
-be a vector of constraints parameters with $\lambda_i >0$. The equivalent QUBO formulation of QCOP is given by,
+Quadratic unconstrained binary optimization problems, as their name implies, cannot have constraints. Alternatively, they only 
+have an objective function that 
+one can to maximize (or minimize). Furthermore, the decision variables must be binary. Let $\lambda = (\lambda_1,\lambda_2)$ 
+be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq.(\ref{QCOP}) is given by,
 \begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
 with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
-penalizing solutions with more or less than $m$ turbines, while the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizes
+penalizing solutions that violate the exactly $m$ turbines constraint, while the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizes
 solutions that place turbines in high proximity. The two penalty terms $\lambda_1$ and $\lambda_2$ 
@@
 for an arbitrary selection of $\lambda$.
+
+\subsection{Solving QCOP and QUBO using Advanced Optimization Methods}
 

### 2020-04-08 paper.tex
commit.
removed log and gz

@@
 &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
-\end{eqnarray} Constraints~\ref{Cardinality} ensure that exactly $m$ turbines are placed on the grid, while~\ref{Grid}
-enforce proximity constraints between the turbines.
+\end{eqnarray} Constraint~\ref{Cardinality} and \ref{Grid} ensure that exactly $m$ turbines are placed on the grid
+and enforce proximity constraints between the turbines, respectively.
 
@@
 \subsection{Solving QCOP and QUBO using Advanced Optimization Methods}
+
+\subsubsection{QCOP solving methods}
+
+\todo{Jason's part}
+
+\subsubsection{Specialized Hardware for QUBO}
+
+\todo{Eldan's part}
 

### 2020-04-09 paper.tex
minor in evaluation.

@@
 
+In this section, we 
+present an empirical evaluation of our quadratic models based on 
+standard WFLO instances.
+Our main results are: \begin{itemize}
+	\item QUBO based WFLO using the digital annealer provides state-of-the-art results in seconds.
+	\item Given enough time LS quadratic models are superior to both linear models and to an SS approximations. 
+\end{itemize}
+We start by outlining the experimental design of our evaluation, followed by a presentation of our main results
+and a discussion of the results together with the limitations of the approach.
+
 \subsection{Experimental Design}
@@
 
-\paragraph{Experimental Setting and Implementation Details}
+\paragraph{Experimental Setting}
+
+\paragraph{Implementation Details}
 

### 2020-04-14 paper.tex
Edited Abstract and Introduction.

@@
 %role as a source of renewable energy and 
-The wind farm layout optimization (WFLO) problem concerns
-selecting turbine positions to maximize energy production.
+The wind farm layout optimization (WFLO) problem requires the
+selection of wind turbine positions to maximize energy production.
 While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models.
-Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
+Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
 %It is a key problem 
@@
 improve over existing WFLO solutions.
-In particular, the QUBO model excels at finding high quality solutions very quickly while the QCOP model is able to find better solutions and provide quality guarantees over a longer run-time.
+In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
 \end{abstract}
@@
 \section{Introduction}
+% Todo
+The goal in the wind farm layout optimization (WFLO)
+problem~\cite{MOSETTI1994105} is to locate wind turbines within a
+predefined area to maximize the total power harvested from the wind
+stream.  While physical considerations based on the diameter of the
+blades prevent turbines being placed too close to each other, turbine
+location decisions influence the total power production due to
+interference effects (i.e., wakes) generated by upstream turbines.
+These inter-turbine effects are naturally captured using quadratic
+constraints and objective functions. However, with a few exceptions (e.g., \cite{turner2014new}), exact optimization
+approaches that have been proposed for the WFLO
+problem~\cite{Zhang2014,kuo2016wind} \todo{[Add more references here]}
+represent the quadratic relations using linear models
+that approximate the true dependencies.
 
-We revisit 
-the wind farm layout optimization (WFLO) problem 
-that was first considered 
-in~\citet{MOSETTI1994105}. 
-In WFLO, we aim to place wind turbines 
-in a predefined area to maximize 
-the total generated power. 
-Turbine location decisions    
-influence
-the
-total power production due to  
-interference effects, aka wakes, generated by upstream 
-turbines. Furthermore, 
-there are proximity constraints between turbines
-that prevent from turbines to be placed next to each other.
-These inter-turbine effects 
-are naturally captured using quadratic constraints and objective
-functions. Several exact optimization approaches have 
-been proposed to solve
-the WFLO problem~\cite{turner2014new,Zhang2014}. However, these
-approaches represented the quadratic effects and constraints
-using linear models that approximate the true dependencies.
-
-In this work, we introduce two 
-novel quadratic WFLO models. The first model 
-represents WFLO using a 
-quadratic constrained optimization problem (QCOP). 
-The QCOP can be solved using state-of-the-art solvers that
-specialize in such problems, e.g., Gurobi~\cite{gurobi}. 
-Our second model is based on a quadratic unconstrained binary optimization (QUBO) representation of the WFLO 
-problem.
-This enables the use of specialized optimization hardware tailored to solve
-QUBOs. In this paper, we use Fujitsu's digital annealer, a simulated annealing based QUBO solver, to demonstrate the strength
-of our approach. The main contributions of this paper are: 
+In this work, we introduce two novel quadratic WFLO models. The first
+model represents WFLO with a quadratic constrained optimization
+problem (QCOP).  The QCOP can be solved using state-of-the-art solvers
+that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
+second model is based on a quadratic unconstrained binary optimization
+(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
+specialized optimization hardware tailored to solve QUBOs. In this
+paper, we use Fujitsu's Digital Annealer (DA), a hardware-based
+QUBO solver, to demonstrate the strength of our approach. The main
+contributions of this paper are:
 \begin{enumerate} 
 \item We propose a novel quadratic modeling framework for WFLO that captures wake 
-effects directly through their linear superposition 
-without the need for approximations and ad-hoc solutions,
-\item We provide the computational complexity of WFLO and show that it is $\mathcal{NP}$-hard, and, 
-\item We test our quadratic formulations
-on a commerical software-based quadratic solver (Gurobi) and on Fujitsu's digital 
-annealer. 
-\end{enumerate} Our experiments are focused on the performance of the two approaches
+effects directly through their linear superposition. 
+%without the need for approximations and ad-hoc solutions,
+\item We prove that WFLO is $\mathcal{NP}$-hard.
+\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
+\end{enumerate}
+
+Our experiments are focused on the performance of the two approaches
 compared to 
@@
 existing solutions when provided sufficient run-time. Furthermore, we
-we observe that the digital 
-annealer often provides order of magnitude faster solutions
-compared to software-based approaches.
+we observe that the DA 
+often provides solutions of similar quality an order of magnitude faster
+than software-based approaches.
 

### 2020-04-14 paper.tex
Edited to end of Sect 2.1

@@
 approaches that have been proposed for the WFLO
-problem~\cite{Zhang2014,kuo2016wind} \todo{[Add more references here]}
+problem~\cite{Zhang2014,kuo2016wind} \todo{[JCB: Add more references here]}
 represent the quadratic relations using linear models
@@
 
-We start this section
-by presenting the
-physical model 
-of a wind farm, which we use throughout the paper.
-Our wind farm model is based on
-the common notation and description 
-presented in~\cite{Zhang2014}.  
-Subsequently, we provide a literature review of 
-existing mathematical
-optimization methods for solving WFLO, focusing
-on their similarities and differences
-compared to our approach.
+We start this section by presenting the physical model of a wind farm
+used throughout the paper.  Our wind farm model is based on the common
+notation and description presented in~\cite{Zhang2014}.  Subsequently,
+we provide a literature review of existing mathematical optimization
+methods for solving WFLO, focusing on their similarities and
+differences with our approach.
 
 \subsection{The Wind Farm Model}
+
+%\paragraph{Wind Farm as a Grid}
+We consider a square-shaped wind farm that we model as an $n\times n$
+grid with $n$ being the number of cells on each axis.  The cells have
+an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
+$n=20$, the farm area will be $2km\times 2km$.  In this work, we
+assume that the terrain is flat and that we have no constraints
+related to the terrain.  Furthermore, we do not account for the
+influence of noise on the wind farm's surroundings. The wind farm
+literature treats the more complex farm models that capture non-flat
+terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
+turbines~\cite{feng2017design}, noise
+effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
+objectives (including turbine installation and maintenance
+costs)~\cite{lackner2007analytical}.  In this work, we focus on the
+novel quadratic approach for WFLO, and hence consider the most
+abstract wind farm model. Our model can be extended, in the
+future, to include additional wind farm settings. \todo{perhaps
+  discuss in future work}
+
+We assume that the number of identical turbines to be placed is known 
+in advance and denote the number of turbines by $m$. To place these $m$ turbines, 
+we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
+In what follows, we explain the two effects using a toy example of placing a single turbine on an 
+$8\times8$ grid (see Figure~\ref{fig:field_model}).
+
 \begin{figure}[t]
@@
 
-\paragraph{Wind Farm as a Grid} We consider a square shaped wind farm that we model 
-as an $n\times n$ grid
-with $n$ being the number of cells on each axis. 
-The cells have an area of $c^2$ (in squared meters), i.e.,
-if $c=100$ meters and $n=20$, the farm area will be $2km\times 2km$.
-In this work, we assume that the terrain is flat and that 
-we have no constraints related to the terrain.
-Furthermore, we do not account for the influence of noise
-on the wind farm's surroundings. The wind farm literature treats the more complex
-farm models that capture non-flat terrain~\cite{song2015lazy,kuo2016wind}, multi-typed turbines~\cite{feng2017design}, noise effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex objectives (including turbine installation  and maintenance costs)~\cite{lackner2007analytical}.
-In this work, we focus on the novel quadratic approach for WFLO, and hence
-consider the most abstract wind farm model. Yet, our model can be extended,
-in the future, to include 
-additional wind farm settings. \todo{perhaps discuss in future work}
+\paragraph{Proximity Constraints}
+Turbines must be placed at least \emph{five rotor diameters}
+apart. Depending on the cell size, we shall formulate the
+corresponding constraint as part of the optimization model. For
+example, assume that a turbine is placed in the center of a
+cell. Then, given a cell size of $c = 100$ meters and a rotor diameter
+of $40$ meters, we must ensure that two turbines are not placed in
+two adjacent cells. In Figure~\ref{fig:field_model}, the red area
+around the turbine represents the forbidden cells.
 
-Returning to the grid model, we assume that the number of homogeneous turbines to be placed is known 
-in advance; we denote the number of turbines by $m$. To place these $m$ turbines, 
-we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
-In what follows, we explain the two effects using a toy example of placing a single turbine on an 
-$8\times8$ grid (see Figure~\ref{fig:field_model}).
-
-\paragraph{Proximity Constraints} Turbines must be placed at least \emph{five rotor diameters} apart. Depending on the cell size, 
-	we shall formulate the corresponding constraint as part of the optimization model. For example, assume that a turbine is placed
-	in the center of a cell. Then, given a cell size of $c = 100$ meters and a rotor diameter of $40$ meters, 
-	we must ensure that two turbines will not placed inside two adjacent cells. In Figure~\ref{fig:field_model}, 
-	the red area around the turbine represents the forbidden cells.
-
-\paragraph{Wakes} Turbines influence each other through interference effects that are referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}. 
-	 Wakes cause reduction of the effective power produced by a turbine due to upstream turbines that change
-	 its airflow dynamics. In Figure~\ref{fig:field_model}, the wind is assumed to be horizontal. Therefore, the wake effect due
-	 to placing the turbine expands as shown in the diagram: it will influence only turbines that will be placed in the orange cells.
-	 In other words, if a turbine is placed in the orange area, its effective
-	 power is reduced due to the upstream turbine that we placed in the $8\times8$ farm. 
-	 Moreover, wakes can occur between multiple turbines, and hence
-	 multiple wakes must be taken into account for each location. 
-	 The reduction of power due to the wake effect is determined by three parameters: (1) the distance between the upstream
-	 turbine and the cell for which we want to compute the effect (denoted by $x$ in Figure~\ref{fig:field_model}), (2) the radius of the cone's opening 
-	 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) 
-	 the superposition of several wind regimes (a wind regime is a combination of free speed and directions).
-	 The wake effect can be pre-computed for each pair of cells 
-	 since we know: the distance ($x$), the turbine specifications and terrain conditions (that jointly dictate $r$), and  
+\paragraph{Wakes}
+Turbines influence each other through interference effects that are
+referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
+Wakes cause reduction of the effective power produced by a turbine due
+to upstream turbines that change the airflow dynamics. In
+Figure~\ref{fig:field_model}, the wind is assumed to be
+blowing from left to right. The wake effect due to placing the turbine
+expands as shown in the diagram.
+%: it will influence only turbines that
+%are placed in the orange cells.  In other words,
+If a turbine is placed in the orange area, its effective power is
+reduced due to the upstream turbine.
+% that we placed in the $8\times8$ farm.
+Moreover, a given turbine may be placed downstream from multiple turbines and hence
+multiple wakes must be taken into account for each location.  The
+reduction of power due to the wake effect is determined by three
+parameters: (1) the distance between the upstream turbine and the cell
+for which we want to compute the effect (denoted by $x$ in
+Figure~\ref{fig:field_model}), (2) the radius of the cone's opening
+(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
+superposition of several wind regimes (a wind regime is a combination
+of free speed and directions).  The wake effect can be pre-computed
+for each pair of cells since we know: the distance ($x$), the turbine
+specifications and terrain conditions (that jointly dictate $r$), and
 probabilistic behavior of wind regimes~\cite{Zhang2014}.
@@
 \paragraph{Effective Power Calculations in the Presence of Wakes} 
-Let $\mathcal{D}$ be the set of wind regimes. Every regime $d\in \mathcal{D}$ is a unique combination of wind direction and free speed. 
+Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a unique \todo{[JCB: Mathematically, I don't think it matters if the regime is unique.
+So delete the word?]} combination of wind direction and free speed.
 The probability of wind regime 
@@
 function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
-a regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind blowing at a free velocity of $12km/h$. The
-probability of this regime is $0.008$. Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
+a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
+probability of this regime is $0.008$.
+
+Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
 Note that turbines can adjust their direction according to the current wind 
@@
 
-Given that we must consider wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
+Given the interference of the wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
 The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
@@
 u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
-\end{equation} with $u_{id\infty}$ being the free wind speed (second component of regime $d$), the set of turbines $\mathcal{U}_{id}$ that are upstream to $i$ 
-under wind regime $d$, and $u_{ijd}$ being the wake 
-reduced speed at $i$ due to upstream turbine $j$. 
+\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
+under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
 The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$.
-Furthermore, the set of upstream turbines, $\mathcal{U}_{id}$ is computed
-by rotating the grid w.r.t.
-to the current wind regime $d \in \mathcal{D}$ and computing which turbine $j$ is located upstream to turbine $i$ given $d$.
+The set of upstream turbines, $\mathcal{U}_{id}$ is computed
+by rotating the grid w.r.t.\ to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$.
 The expression in Eq.
@@
 the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
-the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}): \begin{equation} E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.
-\end{equation} This is also known to be the most accurate analytical total power expression that accounts for multiple wakes.
-However, the expression is not amenable for optimization, and hence, a simplified representation for the total expected energy that considers wakes is required.
+the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}):
+\begin{equation}
+  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
+\end{equation}
+
+  Eq.
+(\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes.\todo{[JCB: REF?]} However, the expression is not amenable to optimization \todo{[JCB: because?]}.
+Hence, a simplified representation for the total expected energy that considers wakes is required.
 
 An alternative 
-form of the total expected power in the presence of multiple wakes is the linear superposition (LS) expression: \begin{equation} \label{eq:ls}
+form of the total expected power in the presence of multiple wakes is the linear superposition (LS) expression:
+
+\begin{equation} \label{eq:ls}
 E_{LS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \Bigg(\frac{1}{3}u_{id\infty}^3 -\sum_{j\in\mathcal{U}_{id}} \frac{1}{3}(u_{id\infty}^3 - u_{ijd}^3)\Bigg).
-\end{equation} Note that $E_{LS}$ has the same parameters as $E_{SS}$; the difference is in the function that ties these parameters together. 
-The LS model is known to be less accurate compared to SS, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
+\end{equation}
+
+Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
+The LS model is less accurate compared to SS,\todo{[JCB: REF?]} yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
 results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.
+\todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic.
+Sounds like it is not natural and we choose it due to its optimization properties.]}
 

### 2020-04-14 paper.tex
Edited up to the first 3 paragraphs in Sect 2.2

@@
 
-The allocation of turbines in a grid-like wind farm was first considered 
-as an optimization problem in~\cite{MOSETTI1994105}. The proposed method in~\cite{MOSETTI1994105}
-searches for improved solutions using a 
-genetic approach~\cite{davis1991handbook},
-which is an approximate technique 
-that does not aim to provide an optimal solutions;
-instead, it attempts to find a quick turbine placement.
+The allocation of turbines in a grid-like wind farm was first
+considered as an optimization problem by~\citet{MOSETTI1994105}. Their
+method employs a genetic
+algorithm~\cite{davis1991handbook}, an approximate technique
+with no guarantees of solution quality but with the goal of quickly finding a high-quality turbine placement.
 
-In follow-up works, 
+In subsequent work~\cite{turner2014new,Zhang2014}, 
 the WFLO problem was solved using \emph{exact}  
-optimization methodologies~\cite{turner2014new,Zhang2014}. 
-These methods guarantee that the returned solution is optimal with respect
-to the specified objective (i.e., maximize total expected power). 
-Indeed, the exact methods were shown to yield higher energy values, yet they often suffer from 
-high computational cost and long runtimes. Therefore, one must strike a balance between optimality and computational requirements of the solution. 
-Fast solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various locations),
-while optimal solutions yield better placements in terms of the total expected energy.
+optimization methodologies that guarantee that the returned solution is optimal with respect
+to the specified objective or is provably within a given bound of optimal.
+%(i.e., maximize total expected power). 
+The exact methods were shown to yield higher energy values, yet they often suffered from 
+high computational cost and long runtimes.
+In practice, it is often desireable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.
 
-In what follows, we provide a literature review that covers both approaches, namely exact vs. approximate WFLO.
-We subsequently relate our 
-work to both exact and approximate methods via our proposed quadratic optimization paradigm for WFLO\footnote{We shall review 
-only literature that considers the same wind farm model as the one presented at the beginning of this section}.
+In what follows, we provide a literature review that covers both exact and approximate methods for WFLO. We subsequently relate our 
+work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
+only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see \todo{[REF]}.}
 

### 2020-04-18 paper.tex
pass over abstract, intro, 2,3.

@@
 selection of wind turbine positions to maximize energy production.
-While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models.
+While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is common to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models.
 Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
@@
 interference effects (i.e., wakes) generated by upstream turbines.
-These inter-turbine effects are naturally captured using quadratic
-constraints and objective functions. However, with a few exceptions (e.g., \cite{turner2014new}), exact optimization
-approaches that have been proposed for the WFLO
-problem~\cite{Zhang2014,kuo2016wind} \todo{[JCB: Add more references here]}
+These inter-turbine effects 
+is best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
+using a sum-of-squares expression~\cite{Zhang2014}. However,
+this wake modeling yields intractable optimization problems
+that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
+An alternative formulation of WFLO relies on a less accurate wakes model, which uses linear superpositioning
+to capture multiple wakes~\cite{donovan2005wind}.
+The latter can be captured using quadratic
+constraints and objective functions. However, exact optimization
+approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
 represent the quadratic relations using linear models
-that approximate the true dependencies.
+that approximate wake dependencies, since until recent years solving linear programming problems
+was computationally faster.
 
@@
 model represents WFLO with a quadratic constrained optimization
-problem (QCOP).  The QCOP can be solved using state-of-the-art solvers
+problem (QCOP).  The QCOP can be solved quickly using state-of-the-art solvers
 that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
@@
 
-The remainder is structured as follows \todo{adjust structure according to final structure of paper}.
+The remainder is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
+optimization solutions to the layout problem. The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO}.
+Section~\ref{sec:eval} demonstrates the usefulness of our approach via a thorough empirical evaluation of the proposed methods. 
+Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.
 
@@
 objectives (including turbine installation and maintenance
-costs)~\cite{lackner2007analytical}.  In this work, we focus on the
+costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we focus on the
 novel quadratic approach for WFLO, and hence consider the most
 abstract wind farm model. Our model can be extended, in the
-future, to include additional wind farm settings. \todo{perhaps
-  discuss in future work}
+future, to include additional wind farm settings.
 
@@
 \paragraph{Effective Power Calculations in the Presence of Wakes} 
-Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a unique \todo{[JCB: Mathematically, I don't think it matters if the regime is unique.
-So delete the word?]} combination of wind direction and free speed.
+Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
 The probability of wind regime 
@@
 
-  Eq.
-(\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes.\todo{[JCB: REF?]} However, the expression is not amenable to optimization \todo{[JCB: because?]}.
+  Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
+  However, the expression is not amenable to optimization since it is a squared root function of a cubic expression. 
   Hence, a simplified representation for the total expected energy that considers wakes is required.
@@
 Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
-The LS model is less accurate compared to SS,\todo{[JCB: REF?]} yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
+The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
 results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.
 \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic.
-Sounds like it is not natural and we choose it due to its optimization properties.]}
+Sounds like it is not natural and we choose it due to its optimization properties.
+Arik: I changed the arguments - please see the abstract and the beginning of the introduction.]}
 
@@
 work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
-only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see \todo{[REF]}.}
+only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see~\cite{Zhang2014}.}
 
 \paragraph{Exact Optimization Methods} 
-\todo{Rephrase and add citations to exact methods}
 The WFLO problem can be solved  
 to optimality, using
-techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{}.
+techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
 These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
@@
 approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
-compared to existing approximate and exact methods~\cite{turner2014new}.
+compared to existing approximate and earlier exact solutions.
 
@@
 is an optimization technique that does not make assumptions on the functional form of the objective.
-This turned out to be computationally intractable for the larger standard WFLO benchmarks.
+However, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.
 
 In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
-model WFLO using a more natural quadratic representation. The difference between our approach and the quadratic model
-proposed in~\cite{turner2014new} is that our model is not an approximation, but rather an exact representation of the LS objective.
+model WFLO using a quadratic representation. The difference between our approach and the quadratic model
+proposed in~\cite{turner2014new} is that our model is not an approximation of the SS objective, 
+but an exact representation of the LS objective.
 
@@
 
-\paragraph{Approximate Methods} \todo{tighten} In contrast to exact methods, approximation algorithms often provide   
+\paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
 quick and sufficiently accurate solutions. Several works solved 
@@
 introduce non-quality related stopping criteria, 
-thus compromising on quality of the attained solution~\cite{davis1991handbook}.
+thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
 Furthermore, there is no natural way of introducing constraints into genetic methods, which
-leads to a potentially costly feasibility check for every solution.
+leads to a potentially costly feasibility check for every solution found.
 
 The first attempt to address these limitations 
-was made in a local search approach for WFLO~\cite{ozturk2004heuristic}. 
+was made using a local search approach for WFLO~\cite{ozturk2004heuristic}. 
 The solution does not guarantee optimality, 
@@
 that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
-that can potentially by far
+that can potentially be far
 from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.
@@
 is its ad-hoc nature. The various components of the algorithm
- must be tailored and geared towards WFLO. This means that small adjustments to our WFLO setting
+ must be tailored and geared towards WFLO. This means that small adjustments of the WFLO setting
  (e.g., adding noise considerations), would require major changes to the SA solution. 
@@
 In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
-our methodology is generic since it is based on our quadratic paradigm and
-on the ability of specialized hardware to solve quadratic (non-ad-hoc) models.
-Quadratic WFLO models can be easily adjusted to include additional problem features, e.g., wind farm characteristics.
+our methodology is generic since it is based on a model and
+on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.
 
-\paragraph{Quadratic Programming as a Unified Framework for Exact and Approximate WFLO}
-\todo{send a clear message here. Perhaps move to next section}.
-
-In this work, we argue that 
-formulating WFLO using quadratic programming enables us to employ 
-advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
-(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
-A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
-approximate methods and found that the two perform in a comparable fashion, without large differences between
-the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
-in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
-This leads to the conclusion that both types of methods (exact and approximate) should be considered. 
-Therefore, in this work, we use a unified paradigm that would enable both exact and approximate solutions to WFLO. Thus, 
-our approach `enjoys' the best of both worlds.
+%\paragraph{Quadratic Programming as a Unified Framework for Exact and Approximate WFLO}
+%\todo{send a clear message here. Perhaps move to next section}.
 
@@
 
-In this part, 
+In this work, we argue that 
+formulating WFLO using quadratic programming enables us to employ 
+advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
+(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
+A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
+approximate methods and found that the two perform in a comparable fashion, without large differences between
+the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
+in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
+This leads to the conclusion that both types of methods (exact and approximate) should be considered. 
+Therefore, in this work, we introduce a unified paradigm, namely quadratic programming, that enables
+both exact and approximate solutions to WFLO. Thus, 
+our approach `enjoys' the best of both worlds (the exact and the approximate).
+
+In this section,
 we formulate the WFLO problem
@@
 \end{equation} The objective function corresponds to the linear superposition (LS)
-expression (Eq.~\ref{eq:ls}). We are now ready to provide the two quadratic programs to represent WFLO.
+expression (c.f., Eq.~\ref{eq:ls} in Section~\ref{sec:related}). We are now ready to provide the two quadratic programs to represent WFLO.
 
@@
 &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
-\end{eqnarray} Constraint~\ref{Cardinality} and \ref{Grid} ensure that exactly $m$ turbines are placed on the grid
+\end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
 and enforce proximity constraints between the turbines, respectively.
@@
 exact solvers. However, these solvers should only be used if the problem is computationally hard.
-We are unaware of an existing complexity result for the LS variation in Eq.~\ref{QCOP}. In what follows, we provide
-a theorem that proves the computational complexity of WFLO. 
+We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP). In what follows, we provide
+a theorem that proves the computational complexity of the LS-based WFLO QCOP. 
 \begin{mythm}
-	The computational complexity of WFLO is $\mathcal{NP}$-hard. 
+	The computational complexity of QCOP is $\mathcal{NP}$-hard. 
 \end{mythm}
@@
 	\end{myproof} To overcome the challenge associated with the computational hardness of WFLO,
-we reformulate the QCOP into a QUBO problem.
+we reformulate the QCOP into a QUBO problem, which allows us to solve WFLO using specialized optimization hardware.
 
@@
 
-Quadratic unconstrained binary optimization problems, as their name implies, cannot have constraints. Alternatively, they only 
+Quadratic unconstrained binary optimization problems, as their name implies, prohibit hard constraints. Instead, 
+they only 
 have an objective function that 
-one can to maximize (or minimize). Furthermore, the decision variables must be binary. Let $\lambda = (\lambda_1,\lambda_2)$ 
+one can to maximize (or minimize) and constraints must be inserted into the objective function
+using penalty terms.
+
+Let $\lambda = (\lambda_1,\lambda_2)$ 
 be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq.(\ref{QCOP}) is given by,
@@
 	\vspace{0.5em}
-\caption{Total Expected Power (kilo Watt/year) after 3600 seconds as function of benchmark and method}\label{tab:results1}
+	\caption{Total Expected Power (kW/year) after 100 seconds as function of benchmark and method \todo{Change to 100 seconds and re-arrange our benchmarks first.} }\label{tab:results1}
+\end{table}
+
+\begin{table}[ht]
+
+	\begin{tabular}{| c | c | c | c | c | c | c | c |}
+		\toprule
+		Wind Directions  & n        & m        & ILP-LS       & QP-LS       & QUBO-LS       & QP-SS       & DA-LS     \\
+		\toprule
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
+			& & 30   & 15739.57 & 15738.45 & \textbf{15742.93} & 15735.1  & 15740.69         \\
+		& & 40 & \textbf{19265.21} & \textbf{19265.21} & 18952.97 & \textbf{19265.21} & \textbf{19265.21}                \\
+				\cline{2-8}
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}           \\
+		&&30   & \textbf{16770.2} & 16758.45 & 16715.19 & 16755.75 & 16570.96                   \\
+		&&40   & \textbf{21977.17} & 21963.8 & 21643.56 & \textbf{21977.17} & 21874.04                     \\
+		\hline
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & 19022.77 & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & \textbf{19221.44} \\
+		&& 30   & 27247.5  & 27442.9  & 27418.74 & 27387.06 & \textbf{27450.99}                  \\
+		&&40   & 34671.62 & \textbf{34938.65} & \textbf{34938.65} & 34817.27 & 34888.81                  \\
+		\cline{2-8}
+		&  \multirow{3}{*}{20}   & 20       & 18448.64 & 19336.33 & 19172.56 & 19354.23 & \textbf{19441.17}          \\
+		&&30   & 26367.2  & 27752 & 27785 & 27492.11 & \textbf{27975.55}                     \\
+		&&40   & 33451.94 & \textbf{35665.6}  & 35127.32 & 35065.85 & 35604.45  \\
+		\bottomrule                   
+	\end{tabular}
+
+\vspace{0.5em}
+\caption{Total Expected Power (kW/year) after 3600 seconds as function of benchmark and method}\label{tab:results2}
 \end{table}
@@
 \label{sec:conclusion}
+
+Future work: additional wind farm settings (noise, terrain...). Convert into qubo.
 

### 2020-04-19 paper.tex
evaluation.

@@
 In this section, we 
-present an empirical evaluation of our quadratic models based on 
+present an empirical evaluation of the two quadratic models based on 
 standard WFLO instances.
 Our main results are: \begin{itemize}
-	\item QUBO based WFLO using the digital annealer provides state-of-the-art results in seconds.
-	\item Given enough time LS quadratic models are superior to both linear models and to an SS approximations. 
+	\item QUBO based WFLO using the digital annealer is competitive with exact methods. 
+	\item Given sufficient runtime the QCOP model provides superior results compared to state-of-the-art methods.
+	\item The runtime for solving the QUBO model is two orders of magnitude lower than QCOP. 
 \end{itemize}
@@
 
-\paragraph{WFLO Instances}
+\paragraph{WFLO Setting and Measurements}
+
+To test and benchmark our models, we used 12 standard benchmark WFLO instances
+from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement}. 
+The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
+2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
+and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$).
+The wind regime distribution presented in Figure~\ref{fig:prob_wind} was 
+elicited from~\cite{Zhang2014}.
+Additional parameters that were used to calculate 
+the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
+ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.
+We have measured solutions using the SS energy expression.
+To to this end, we converted solutions that came from LS based methods into the corresponding SS expression.
+Lastly, the total mean energy is presented in $kW/year$.
+
 \begin{figure}[t]
@@
 
-\paragraph{Experimental Setting}
+We compared our two models, namely QCOP and QUBO, against two state of the art exact optimization methods:
+the integer linear programming (ILP-LS) approach from~\cite{donovan2005wind}, and the 
+model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. The CP approach presented in~\cite{Zhang2014} was
+reported to perform worse than the rest and was left out of the experiments.
+We used state of the art software, namely Gurobi~\cite{gurobi}, to solve QCOP, ILP-SS and QP-SS, and 
+applied both Fujitsu's digital annealer and GUrobi to solve QUBO.
+
+\paragraph{Experimental Procedure}
+
+\todo{Explain how we solved WFLO using the various instances}
 

### 2020-04-20 paper.tex
Modify the order of columns on both tables.
eval.

@@
 present an empirical evaluation of the two quadratic models based on 
-standard WFLO instances.
+12 standard WFLO instances.
 Our main results are: \begin{itemize}
-	\item QUBO based WFLO using the digital annealer is competitive with exact methods. 
-	\item Given sufficient runtime the QCOP model provides superior results compared to state-of-the-art methods.
-	\item The runtime for solving the QUBO model is two orders of magnitude lower than QCOP. 
+	\item QUBO based WFLO using the digital annealer (DA) achieves a similar performance to state-of-the-art exact methods.
+		\item The runtime for solving the QUBO model on the DA is two orders of magnitude lower than QCOP.  
+	\item Given sufficient runtime, solving the QCOP model yields superior results on the larger instances.
 \end{itemize}
-We start by outlining the experimental design of our evaluation, followed by a presentation of our main results
-and a discussion of the results together with the limitations of the approach.
+We start by outlining the experimental design of our evaluation, followed by an overview of our main results,
+and conclude the section with a discussion of the results together with the limitations of our approach.
 
@@
 
-\paragraph{WFLO Setting and Measurements}
+\paragraph{WFLO Instances and Energy Measurement}
 
@@
 We have measured solutions using the SS energy expression.
-To to this end, we converted solutions that came from LS based methods into the corresponding SS expression.
-Lastly, the total mean energy is presented in $kW/year$.
+To to this end, we converted solutions that came from LS based methods into the corresponding SS expression; 
+the total mean energy is calculated in $kW/year$.
 
@@
 reported to perform worse than the rest and was left out of the experiments.
-We used state of the art software, namely Gurobi~\cite{gurobi}, to solve QCOP, ILP-SS and QP-SS, and 
-applied both Fujitsu's digital annealer and GUrobi to solve QUBO.
+We used state of the art software, namely Gurobi~\cite{gurobi}, to solve QCOP, ILP-SS and QP-SS. 
+We shall refer to the QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
+In addition, we 
+applied both Fujitsu's digital annealer and Gurobi to solve the QUBO variant of WFLO. We shall refer to the former as DA-LS, and to the latter
+as QUBO-LS.
 
@@
 
-\todo{Explain how we solved WFLO using the various instances}
+For exact methods (QP-SS, ILP-LS, QP-lS and QUBO-LS),
+we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
+until that point in time. The best solution in terms of energy is the solution that attained the highest objective
+values (SS in QP-SS and LS in the rest of the methods).
+For the digital annealer, we used a single configuration that was found to work best across the different instances \todo{Jason, please insert details about the DA config.
+here} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime
+of the procedure was recorded.
+\todo{@Chris: do we want to share the code that ran the experiments in batches?
+perhaps it should be separated from the problem construction and the two need to be shared separately}
 
-\paragraph{Implementation Details}
+The exact methods were all solved using Gurobi's quadratic programming solver on \todo{Jason, please add machine details here}.
+Furthermore, QUBO-LS was solved on DA version 1 \todo{Eldan: any additional info that we need to add here?}.
 
@@
 		\toprule
-		Wind Directions  & n        & m        & ILP-LS       & QP-LS       & QUBO-LS       & QP-SS       & DA-LS     \\
+		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
 		\toprule
-		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
-		& & 30   & 15739.57 & 15738.45 & \textbf{15742.93} & 15735.1  & 15740.69         \\
-		& & 40 & \textbf{19265.21} & \textbf{19265.21} & 18952.97 & \textbf{19265.21} & \textbf{19265.21}                \\
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
+		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
+		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}           \\
-		&&30   & \textbf{16770.2} & 16758.45 & 16715.19 & 16755.75 & 16570.96                   \\
-		&&40   & \textbf{21977.17} & 21963.8 & 21643.56 & \textbf{21977.17} & 21874.04                     \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}           \\
+		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
+		&&40   & 21874.04 & 21643.56  & 21905.5 & \textbf{21977.17} & \textbf{21977.17}        \\
 		\hline
-		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & 19022.77 & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & \textbf{19221.44} \\
-		&& 30   & 27247.5  & 27442.9  & 27418.74 & 27387.06 & \textbf{27450.99}                  \\
-		&&40   & 34671.62 & \textbf{34938.65} & \textbf{34938.65} & 34817.27 & 34888.81                  \\
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
+		&& 30  & \textbf{27450.99} & 27418.74 & 27442.9 & 27387.06 & 27247.5                     \\
+		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
 		\cline{2-8}
-		&  \multirow{3}{*}{20}   & 20       & 18448.64 & 19336.33 & 19172.56 & 19354.23 & \textbf{19441.17}          \\
-		&&30   & 26367.2  & 27752 & 27785 & 27492.11 & \textbf{27975.55}                     \\
-		&&40   & 33451.94 & \textbf{35665.6}  & 35127.32 & 35065.85 & 35604.45  \\
+		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
+		&&30   & \textbf{27975.55} & 27443.8  & 27752 & 27492.11  & 26367.2                      \\
+		&&40   & 35604.45 & 35127.32 & \textbf{35665.6}   & 35065.85 & 33451.94 \\
 		\bottomrule                   
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW/year) after 100 seconds as function of benchmark and method \todo{Change to 100 seconds and re-arrange our benchmarks first.} }\label{tab:results1}
+	\caption{Total Expected Power (kW/year) after 100 seconds as function of benchmark and method.
+* stands for optimal solution.
+\todo{@Jason: could you please check if my marking of * for optimal solutions is correct?
+I think that optimality was proven only for $m=20$ and single wind direction, right?} }\label{tab:results1}
 \end{table}
@@
 		\toprule
-		Wind Directions  & n        & m        & ILP-LS       & QP-LS       & QUBO-LS       & QP-SS       & DA-LS     \\
+		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
 		\toprule
-		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
-			& & 30   & 15739.57 & 15738.45 & \textbf{15742.93} & 15735.1  & 15740.69         \\
-		& & 40 & \textbf{19265.21} & \textbf{19265.21} & 18952.97 & \textbf{19265.21} & \textbf{19265.21}                \\
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
+		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
+		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}  & \textbf{11404.8}           \\
-		&&30   & \textbf{16770.2} & 16758.45 & 16715.19 & 16755.75 & 16570.96                   \\
-		&&40   & \textbf{21977.17} & 21963.8 & 21643.56 & \textbf{21977.17} & 21874.04                     \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8} *          \\
+		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.2 }                 \\
+		&&40   & 21874.04 & 21832.5  & 21963.81 & \textbf{21977.17} & \textbf{21977.17}        \\
 		\hline
-		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & 19022.77 & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & \textbf{19221.44} \\
-		&& 30   & 27247.5  & 27442.9  & 27418.74 & 27387.06 & \textbf{27450.99}                  \\
-		&&40   & 34671.62 & \textbf{34938.65} & \textbf{34938.65} & 34817.27 & 34888.81                  \\
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19213.96 \\
+		&& 30  & \textbf{27450.99} & 27442.9 & 27442.9 & 27387.06 & 27389.21                     \\
+		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34856.45          \\
 		\cline{2-8}
-		&  \multirow{3}{*}{20}   & 20       & 18448.64 & 19336.33 & 19172.56 & 19354.23 & \textbf{19441.17}          \\
-		&&30   & 26367.2  & 27752 & 27785 & 27492.11 & \textbf{27975.55}                     \\
-		&&40   & 33451.94 & \textbf{35665.6}  & 35127.32 & 35065.85 & 35604.45  \\
+		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19394.58 & 19336.33 & 19354.23  & 19368.86            \\
+		&&30   & \textbf{27975.55} & 27785.74  & 27752 & 27492.11  & 27633.2                      \\
+		&&40   & 35604.45 & 35480.9 & \textbf{35665.6}   & 35065.85 & 35210.78 \\
 		\bottomrule                   
@@
 \vspace{0.5em}
-\caption{Total Expected Power (kW/year) after 3600 seconds as function of benchmark and method}\label{tab:results2}
+\caption{Total Expected Power (kW/year) after 3600 seconds as function of benchmark and method. * stands for optimal solution.}\label{tab:results2}
 \end{table}

### 2020-04-20 paper.tex
evaluation: finished experimental setting, moved to main results.

@@
 
-\paragraph{WFLO Instances and Energy Measurement}
+\paragraph{Standard WFLO Instances}
 
@@
 2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
-and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$).
+and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). We denote by $\mathcal{W}$ the set of 12 instances. 
 The wind regime distribution presented in Figure~\ref{fig:prob_wind} was 
@@
 ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.
-We have measured solutions using the SS energy expression.
-To to this end, we converted solutions that came from LS based methods into the corresponding SS expression; 
-the total mean energy is calculated in $kW/year$.
 
@@
 
-For exact methods (QP-SS, ILP-LS, QP-lS and QUBO-LS),
+For exact methods (QP-SS, ILP-LS, QP-LS and QUBO-LS),
 we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
@@
 
+\paragraph{Performance Measurement}
+
+We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in $kW/year$.}.
+The only method that optimizes on SS is QP-SS. Hence, for the other methods, we converted solutions
+LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in~\cite{Zhang2014}.
+Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
+Then, the SS based energy expression can be written as, 
+$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$
+
+Next, we tested the performance of our approach as function of solution time. To this end, we 
+compared the average performance of all LS-based methods on all 12 WFLO instances \footnote{The QP-SS method was not considered for performance over time, since it optimizes a different objective function.}.
+Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specs) we used the \emph{mean relative error} (MRE) measure.
+
+The MRE is defined with respect to the best solution obtained per instance and can be computed over time. Then, the
+time-varying MRE can be averaged across the different scenarios and the average MRE (aMRE) can be used
+for the desired comparison. We shall demonstrate the construction below.
+
+Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
+Moreover, we let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
+across the different LS-based methods.
+For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742 kW/year$.
+Therefore, we let $b_w = 15742$.
+Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
+Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS\}$.
+Further, denote by $b_{w,t,a}$ denote the best solution by time $t$ of approach $a$ over instance $w$.
+Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by $$MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.$$ We can now average over the different instances at time $t$.
+The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as $$MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).$$
+We shall present this quantity ($MRE(t,a)$) for each of the approaches over time to demonstrate the speed of convergence of the proposed LS-based methods.
+
 \subsection{Main Results}
 
-\begin{table}[ht]
+In this part, we present the main findings of our empirical evaluation. 
+We shall start by showing the best attained solutions by all methods.
+Table~\ref{tab:results1} contains the best energy (in kW/hour) attained by the methods
+for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. The runtime for DA-LS was always less than $10$ seconds.
+We observe that...
+
+Table~\ref{tab:results2} shows the same type of results after $3600$ seconds.
+
+\begin{table}[t]
 
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 
@@
 
-\begin{figure}[t!]%
+\begin{figure}[t]%
 

### 2020-04-22 paper.tex
evaluation.

@@
 2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
-and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). We denote by $\mathcal{W}$ the set of 12 instances. 
-The wind regime distribution presented in Figure~\ref{fig:prob_wind} was 
-elicited from~\cite{Zhang2014}.
+and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). 
+For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}.
+We denote $\mathcal{W}$ the set of 12 instances.
+Note that instances with $36$ wind regimes are considered computationally harder~\cite{Zhang2014}.
+
 Additional parameters that were used to calculate 
@@
 
-	\caption{Wind regimes: x-axis is the angle (times 10 degrees), y-axis is the probability for wind regime, color corresponds to free wind speed.}\label{fig:prob_wind}
+	\caption{Wind regimes: x-axis is the angle (times 10 degrees), y-axis is the probability for wind regime, and color corresponds to free wind speed for the different wind directions.}\label{fig:prob_wind}
 \end{figure}
@@
 
-We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in $kW/year$.}.
+We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in kW.}.
 The only method that optimizes on SS is QP-SS. Hence, for the other methods, we converted solutions
@@
 across the different LS-based methods.
-For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742 kW/year$.
-Therefore, we let $b_w = 15742$.
+For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
 Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
 Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS\}$.
-Further, denote by $b_{w,t,a}$ denote the best solution by time $t$ of approach $a$ over instance $w$.
-Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by $$MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.$$ We can now average over the different instances at time $t$.
-The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as $$MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).$$
-We shall present this quantity ($MRE(t,a)$) for each of the approaches over time to demonstrate the speed of convergence of the proposed LS-based methods.
+Further, denote by $b_{w,t,a}$ denote the best solution attained before time $t$ of approach $a$ for instance $w$.
+Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.\end{equation} We can now average over the different instances at time $t$.
+The average performance of approach $a$ at time $t$, $aMRE(t,a)$ can be computed as, \begin{equation}aMRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
+We shall present $aMRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
+\todo{Jason: please check that this is consistent with how you calculated MRE in the experiment.}
 
@@
 We shall start by showing the best attained solutions by all methods.
-Table~\ref{tab:results1} contains the best energy (in kW/hour) attained by the methods
-for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. The runtime for DA-LS was always less than $10$ seconds.
-We observe that...
+Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
+for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
+The runtime for DA-LS was always less than $10$ seconds.
 
-Table~\ref{tab:results2} shows the same type of results after $3600$ seconds.
-
-\begin{table}[t]
+\begin{table}[t!]
 
@@
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}           \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*           \\
 		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW/year) after 100 seconds as function of benchmark and method.
+	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method.
 * stands for optimal solution.
@@
 
-\begin{table}[t]
+\begin{table}[t!]
 
@@
 	\vspace{0.5em}
-\caption{Total Expected Power (kW/year) after 3600 seconds as function of benchmark and method. * stands for optimal solution.}\label{tab:results2}
+	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method. * stands for optimal solution.}\label{tab:results2}
 \end{table}
+
+We observe that the DA-based solution provides
+comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
+the three quadratic LS-based approaches (DA-LS, QUBO-LS, and QP-LS) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
+the linear approach (ILP-LS) finds better solutions \todo{@Chris, Jason: do we have a good explanation for this?}.
+Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds.
+We see that the DA is still competitive with the other approaches (recall, its solution was found after $10$ seconds),
+while QP-LS finds the best solutions in 10 out of the 12 scenarios. For the 6 hard scenarios (12 wind directions), our quadratic solutions
+dominate all 6.
+
+We now turn to present the $aMRE(a,t)$ over time: lower values indicate better performance. The different lines 
+correspond to the 5 approaches. The horizontal axis represents time in seconds. Figure~\ref{fig:results1} 
+presents the first $50$ seconds focusing on speed of performance, while Figure~\ref{fig:results2} corresponds
+to the full run ($3600$ seconds).
 
@@
 
-	\subfloat[MRE after 50 seconds]{{\includegraphics[width=0.5\textwidth]{energy_50.png} }}%
-	\quad
-	\subfloat[MRE after 1 hour]{{\includegraphics[width=0.5\textwidth]{energy_3600.png} }}\\%\\
-	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime.}%
+	\includegraphics[width=\textwidth]{energy_50.png}  
+	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (first 50 seconds).}%
 
-	\label{fig:main_effects}%
+	\label{fig:results1}%
+\end{figure}
+
+\begin{figure}[t]%
+
+	\includegraphics[width=\textwidth]{energy_3600.png}   \\%\\
+	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (3600 seconds).}%
+
+	\label{fig:results2}%
 \end{figure}

### 2020-04-22 paper.tex
evaluation and conclusion

@@
 Next, we tested the performance of our approach as function of solution time. To this end, we 
-compared the average performance of all LS-based methods on all 12 WFLO instances \footnote{The QP-SS method was not considered for performance over time, since it optimizes a different objective function.}.
+compared the average performance of all WFLO methods on all 12 WFLO instances.
 Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specs) we used the \emph{mean relative error} (MRE) measure.
-
-The MRE is defined with respect to the best solution obtained per instance and can be computed over time. Then, the
-time-varying MRE can be averaged across the different scenarios and the average MRE (aMRE) can be used
-for the desired comparison. We shall demonstrate the construction below.
+The MRE is defined with respect to the best solution obtained per instance and can be computed over time.
+Then, the
+time-varying MRE can be averaged across the 12 scenarios and presented over time. In what follows, we demonstrate the construction of the MRE measure.
 
@@
 Moreover, we let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
-across the different LS-based methods.
+across the different methods.
 For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
 Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
-Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS\}$.
+Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$.
 Further, denote by $b_{w,t,a}$ denote the best solution attained before time $t$ of approach $a$ for instance $w$.
 Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.\end{equation} We can now average over the different instances at time $t$.
-The average performance of approach $a$ at time $t$, $aMRE(t,a)$ can be computed as, \begin{equation}aMRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
-We shall present $aMRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
+The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
+We shall present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
 \todo{Jason: please check that this is consistent with how you calculated MRE in the experiment.}
@@
 	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method.
-* stands for optimal solution.
+* denotes optimal solution.
 \todo{@Jason: could you please check if my marking of * for optimal solutions is correct?
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method. * stands for optimal solution.}\label{tab:results2}
+	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method. * denotes optimal solution.}\label{tab:results2}
 \end{table}
@@
 while QP-LS finds the best solutions in 10 out of the 12 scenarios. For the 6 hard scenarios (12 wind directions), our quadratic solutions
-dominate all 6.
+dominate all 6. \todo{@Chris: I am not sure that the first table contributes anything beyond the second table}
 
-We now turn to present the $aMRE(a,t)$ over time: lower values indicate better performance. The different lines 
-correspond to the 5 approaches. The horizontal axis represents time in seconds. Figure~\ref{fig:results1} 
-presents the first $50$ seconds focusing on speed of performance, while Figure~\ref{fig:results2} corresponds
-to the full run ($3600$ seconds).
+We now turn to present the mean relative error over time per approach, i.e., we are presenting $MRE(a,t)$). Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
+presents the MRE after $50$ seconds of running the solutions focusing runtime performance, while Figure~\ref{fig:results2} corresponds
+to the full run (after $3600$ seconds). The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
+correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.
+
+The results show that on average (across the 12 instances),
+the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS (after around 1250 seconds).
+
+The two methods that are based on advanced software (QP-LS) and hardware (DA-LS) solutions
+outperform the rest both in terms of runtime (best results in short time), and in terms of overall performance.
+Specifically, for most instances, QP-LS gives the best solution at 3600 seconds, while the DA-based approach
+yields competitive solutions after less than 10 seconds.
+Furthermore, QP-LS converges to the best solution faster than all other exact methods.
+In fact, it is inferior only to the approximate DA-LS method, which is geared mainly toward runtime.
+This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
 
@@
 	\includegraphics[width=\textwidth]{energy_50.png}  
-	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (first 50 seconds).}%
+	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (after 50 seconds).}%
 
@@
 
-\begin{figure}[t]%
+uwo\begin{figure}[t]%
 
 	\includegraphics[width=\textwidth]{energy_3600.png}   \\%\\
-	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (3600 seconds).}%
+	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (after 3600 seconds).}%
 
@@
 
-\subsection{Discussion \& Limitations}
+\subsection{Discussion}
+
+Our result show empirically that the quadratic programming models that we
+introduced in Section~\ref{sec:QUBO4WFLO} (QCOP and QUBO) are useful in two ways.
+First, QCOP and QUBO are the basis for the QP-LS and QUBO-LS methods that couple
+quadratic formulations of WFLO with Gurobi optimizer. This turns out
+to be a powerful method as it gives the best performance after 3600 seconds for most instances (7 out of 12),
+and especially,
+for the hard instances with 36 wind directions. In addition, the runtime performance of QP-LS is second
+only to the digital annealer.  
+Second, the QUBO model can be mounted onto a quantum-inspired optimization hardware (e.g., Fujitsu's digital annealer) demonstrated by the DA-LS approach in our evaluation.
+DA-LS yields the best results in 7 out of 12
+instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime.
+
+The ability of DA-LS to quickly obtain competitive solutions on large WFLO instances 
+can be employed to solve
+multiple wind farm design problems.
+This allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that came
+(so far) with re-solving the problem. In addition, the performance loss due to the use of an approximate method 
+that uses advanced technology is negligible. This balance (instead of trade-off) between
+runtime and total energy, points at the usefulness of the proposed quadratic formulations of WFLO.
+\todo{maybe we should say some of these things in the intro and bring citations to the fact that demand for wind farms increases.
+I found some papers about surge in China and the UK.}
 
@@
 
-Future work: additional wind farm settings (noise, terrain...). Convert into qubo.
+In this work, we presented a novel quadratic programming framework
+for solving 
+wind farm layout optimization. 
+The framework is based on two models: a constrained model (QCOP) and an unconstrained binary
+model (QUBO). The former can be used to solve WFLO with advanced optimization software,
+while the latter can be mounted onto quantum-inspired dedicated hardware that can quickly (yet approximately) solve WFLO. We used the constrained quadratic formulation to prove 
+that the computational complexity of WFLO is $\mathcal{NP}$-hard, a pessimistic result that 
+was also noticed experimentally in~\cite{Zhang2014}. However,
+our empirical evaluation shows that when advanced exact solvers like Gurobi~\cite{gurobi} are
+coupled with our QCOP and QUBO models they can solve small WFLO instances to optimality in less than 3600 seconds.
+We also show that when an advanced hardware is used with the QUBO representation of WFLO, even large instances can be solved quickly with no significant decrease in the total energy compared to state-of-the-art exact methods.
+This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our proposed quadratic framework useful for designing new wind farms, while multiple input parameters.
+
+In future work, we aim at extending our quadratic models to include additional 
+WFLO aspects, such as noise, non-flat terrain and multiple objectives. 
+Moreover, we wish to explore real-world wind farm design problems 
+that may be larger and more complex than the 12 standard instances, and solve them 
+using our quadratic programming approach.
 

### 2020-04-22 paper.tex
literature remark.
literature.
Revised abstract

@@
 \title{\mytitle}
+
 %\subtitle{Extended Abstract}
@@
 %role as a source of renewable energy and 
-The wind farm layout optimization (WFLO) problem requires the
-selection of wind turbine positions to maximize energy production.
-While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is common to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models.
-Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
+  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference among turbines.
+Given the pairwise wake interactions, it is natural
+%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
+to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization models.
+Recent advances in both optimization hardware and software have targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
 %It is a key problem 
@@
 A thorough empirical 
-evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
-Our results show that both approaches yield fast and high-quality solutions 
-and
-improve over existing WFLO solutions.
+evaluation using a commercial solver and specialized QUBO hardware
+%demonstrates the strengths of the two models. 
+%Our results
+show that both approaches yield fast, high-quality solutions 
+that
+improve the state of the art.
 In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
@@
 compared to existing approximate and earlier exact solutions.
+\todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}
 

### 2020-04-22 paper.tex
Edited abstract and Introduction

@@
 to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization models.
-Recent advances in both optimization hardware and software have targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
+Recent advances in both optimization hardware and software have targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
 %It is a key problem 
@@
 interference effects (i.e., wakes) generated by upstream turbines.
-These inter-turbine effects 
-is best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
+These inter-turbine effects are best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
 using a sum-of-squares expression~\cite{Zhang2014}. However,
@@
 that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
-An alternative formulation of WFLO relies on a less accurate wakes model, which uses linear superpositioning
+An alternative formulation of WFLO relies on a less accurate wake model, which uses linear superpositioning
 to capture multiple wakes~\cite{donovan2005wind}.
@@
 approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
-represent the quadratic relations using linear models
-that approximate wake dependencies, since until recent years solving linear programming problems
-was computationally faster.
+represent the quadratic relations using approximate linear models,
+%that approximate wake dependencies,
+as, until recent years, solving quadratic programming problems
+was computationally challenging compared to linear models.
 
 In this work, we introduce two novel quadratic WFLO models. The first
-model represents WFLO with a quadratic constrained optimization
-problem (QCOP).  The QCOP can be solved quickly using state-of-the-art solvers
-that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
-second model is based on a quadratic unconstrained binary optimization
+model represents WFLO as a quadratic constrained optimization
+problem (QCOP).  The QCOP can be solved quickly using state-of-the-art mathematical solvers
+that have been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
+second model is a quadratic unconstrained binary optimization
 (QUBO) representation of the WFLO problem.  Such a formulation enables the use of
-specialized optimization hardware tailored to solve QUBOs. In this
-paper, we use Fujitsu's Digital Annealer (DA), a hardware-based
-QUBO solver, to demonstrate the strength of our approach. The main
+nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics} here.
+%, a hardware-based
+%QUBO solver, to demonstrate the strength of our approach.
+
+The main
 contributions of this paper are:
 \begin{enumerate} 
-\item We propose a novel quadratic modeling framework for WFLO that captures wake 
-effects directly through their linear superposition. 
+\item We propose a novel quadratic modeling framework for WFLO that encompasses both constrained and unconstrained quadratic optimization models.
+ %that captures wake 
+%effects directly through their linear superposition. 
 %without the need for approximations and ad-hoc solutions,
@@
 using Gurobi and the DA
-achieves state-of-the-art performance, often outperforming
-existing solutions when provided sufficient run-time. Furthermore, we
-we observe that the DA 
-often provides solutions of similar quality an order of magnitude faster
+achieves state-of-the-art performance.
+%, often outperforming existing solutions when provided sufficient run-time.
+Furthermore, we observe that the DA 
+often provides solutions of similar quality two orders of magnitude faster
 than software-based approaches.
 
-The remainder is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
-optimization solutions to the layout problem. The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO}.
-Section~\ref{sec:eval} demonstrates the usefulness of our approach via a thorough empirical evaluation of the proposed methods. 
+The remainder of this paper is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
+optimization solutions to the layout problem.
+The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO} and Section~\ref{sec:eval} demonstrates the value of our approach via a thorough empirical evaluation.
+%of the proposed methods. 
 Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.

### 2020-04-22 paper.tex
Edited to end of Sect 2.1

@@
 We start this section by presenting the physical model of a wind farm
-used throughout the paper.  Our wind farm model is based on the common
-notation and description presented in~\cite{Zhang2014}.  Subsequently,
+%used throughout the paper.
+%Our wind farm model is
+based on the common
+notation and description presented in Zhang et al.\ \cite{Zhang2014}.  Subsequently,
 we provide a literature review of existing mathematical optimization
@@
 grid with $n$ being the number of cells on each axis.  The cells have
-an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
-$n=20$, the farm area will be $2km\times 2km$.  In this work, we
+an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
+$n=20$, the farm area will be $2km\times 2km$.  We
 assume that the terrain is flat and that we have no constraints
@@
 abstract wind farm model. Our model can be extended, in the
-future, to include additional wind farm settings.
+future, to include additional problem characteristics.
 
@@
 corresponding constraint as part of the optimization model. For
-example, assume that a turbine is placed in the center of a
-cell. Then, given a cell size of $c = 100$ meters and a rotor diameter
-of $40$ meters, we must ensure that two turbines are not placed in
-two adjacent cells. In Figure~\ref{fig:field_model}, the red area
+example, if a turbine is placed in the center of a
+cell, a cell size of $c = 100$ meters and a rotor diameter
+of $40$ meters, requires that two turbines are not placed in adjacent cells. In Figure~\ref{fig:field_model}, the red area
 around the turbine represents the forbidden cells.
@@
 \paragraph{Wakes}
-Turbines influence each other through interference effects that are
-referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
+Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
+%that are
+%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
 Wakes cause reduction of the effective power produced by a turbine due
@@
 for which we want to compute the effect (denoted by $x$ in
-Figure~\ref{fig:field_model}), (2) the radius of the cone's opening
+Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
-superposition of several wind regimes (a wind regime is a combination
-of free speed and directions).  The wake effect can be pre-computed
+combination of several wind regimes where a wind regime consists of 
+ther wind's free speed and direction.  The wake effect can be pre-computed
 for each pair of cells since we know: the distance ($x$), the turbine
@@
 \paragraph{Effective Power Calculations in the Presence of Wakes} 
-Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
+Let $\mathcal{D}$ be the set of wind regimes.
+%where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
 The probability of wind regime 
@@
 function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
-a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
+a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
 probability of this regime is $0.008$.
 
-Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
-Note that turbines can adjust their direction according to the current wind 
-regime.
-
-Given the interference of the wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
-The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
+Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
+regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
+Rather, $u_{id}$, can be computed using the following equation~\cite{Zhang2014}:
 \begin{equation}
@@
 the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
-the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}):
+the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
+%~(\cite{Zhang2014}):
 \begin{equation}
@@
   Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
-  However, the expression is not amenable to optimization since it is a squared root function of a cubic expression. 
-  Hence, a simplified representation for the total expected energy that considers wakes is required.
+  However, the expression is challenging to exact mathematical optimization techniques due to its complexity (i.e., the cube of a square root).
+  %since it is a squared root function of a cubic expression. 
+  %Hence, a simplified representation for the total expected energy that considers wakes is required.
 
@@
 The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
-results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.
+naturally results in a quadratic objective function in our WFLO formulations. Such functions have been the subject of recent research advances \todo{[REFs]}, motivating the contributions here.
+%, which is a desired property from an optimization standpoint.
 \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic.

### 2020-04-22 paper.tex
Spell check

@@
 combination of several wind regimes where a wind regime consists of 
-ther wind's free speed and direction.  The wake effect can be pre-computed
+the wind's free speed and direction.  The wake effect can be pre-computed
 for each pair of cells since we know: the distance ($x$), the turbine

### 2020-04-22 paper.tex
Finished edit of Section 2

@@
 The exact methods were shown to yield higher energy values, yet they often suffered from 
-high computational cost and long runtimes.
-In practice, it is often desireable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.
+high computational cost and long run-times.
+In practice, it is often desirable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.
 
@@
 work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
-only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see~\cite{Zhang2014}.}
+only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see Zhang et al.\ \cite{Zhang2014}.}
 
@@
 the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
-MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
-approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
-compared to existing approximate and earlier exact solutions.
+MILP approach based on the LS energy expression~\cite{donovan2005wind}.
+Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation.
+The two resulting methods performed well
+compared to existing approximate and exact techniques.
 \todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}
@@
 The most recent attempt to solve the wind farm variation that we presented 
-in this section is given in~\cite{Zhang2014}. The authors compared
+in this section is given in Zhang et al.\ \cite{Zhang2014}. The authors compared
 previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
@@
 
-In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
-model WFLO using a quadratic representation. The difference between our approach and the quadratic model
-proposed in~\cite{turner2014new} is that our model is not an approximation of the SS objective, 
+In our work, we build upon the approach in Donovan~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
+model WFLO using a quadratic representation. %The difference between our approach and the quadratic model
+%proposed in
+Unlike Turner et al.\ \cite{turner2014new} our model is not an approximation of the SS objective, 
 but an exact representation of the LS objective.
 
-As we mentioned before, the main advantage of exact techniques
-is that they provide guarantees on solution quality. However, their disadvantage is
-that they may take a very long time to execute before finding an optimal solution. 
+While the main advantage of exact techniques
+is that they provide guarantees on solution quality, they may take a very long time to  find an optimal solution. 
 In fact, Section~\ref{sec:QUBO4WFLO} shows that WFLO is computationally hard, a result
@@
 \paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
-quick and sufficiently accurate solutions. Several works solved 
+quick and, hopefully, sufficiently accurate solutions. Several works solved 
 WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
@@
 to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
-(i.e., they may run forever without achieving a quality criteria), and when they do terminate (in practical cases) they do
+(i.e., they may run forever without achieving a quality criteria), and even when they do terminate (in practical cases) they do
 not guarantee optimal solutions. Hence, one must often
@@
 when a stopping criteria is reached.
-The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
-that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
-that can potentially be far
-from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.
+The main limitation of this approach 
+%~\cite{ozturk2004heuristic}
+is that it cannot perform an action 
+that leads to an infeasible state and, therefore, quickly converges to a sub-optimal solution (local optimum) 
+that can potentially be much lower quality than the globally optimal solution~\cite{rivas2009solving}.
 
-To overcome this limitation, the authors of~\cite{rivas2009solving} used simulated annealing (SA), 
+To overcome this limitation, Rivas et al.\ \cite{rivas2009solving} used simulated annealing (SA), 
 a neighborhood search method that allows visits to infeasible solutions. 
 This creates a well-connected solution space that leads the algorithm to escape 
-local optima. The SA method was shown to be superior to~\cite{ozturk2004heuristic} 
-and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach 
+local optima. The SA method was shown to be superior to Ozturk \& Norman~\cite{ozturk2004heuristic} 
+and to the genetic method proposed in Grady et al.\ \cite{grady2005placement}. A drawback of the SA approach 
 is its ad-hoc nature. The various components of the algorithm
- must be tailored and geared towards WFLO. This means that small adjustments of the WFLO setting
- (e.g., adding noise considerations), would require major changes to the SA solution. 
-Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.
+ must be tailored to the problem at hand and, as a result, small adjustments of the WFLO setting
+ (e.g., adding noise considerations), would require major changes to the SA implementation. 
+Additional methods and experimental comparisons between various approximate algorithms are available in Samorani~\cite{samorani2013wind}.
 
-In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
-our methodology is generic since it is based on a model and
+In our work, we also use an SA approach (similarly to Rivas et al.\ \cite{rivas2009solving}). However,
+our methodology is generic since it is based on a declarative problem model and
 on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.

### 2020-04-22 paper.tex
Small edit at the beginning of Sect 3

@@
 formulating WFLO using quadratic programming enables us to employ 
-advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
-(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
+recent advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
+(e.g., Fujitsu's Digital Annealer).
+%that solves quadratic problems using principles from simulated annealing). 
 A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 

### 2020-04-22 paper.tex
Working on Sect 3

@@
 
-In this work, we argue that 
-formulating WFLO using quadratic programming enables us to employ 
-recent advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
-(e.g., Fujitsu's Digital Annealer).
-%that solves quadratic problems using principles from simulated annealing). 
-A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
+A benchmark study comparing state-of-the-art exact approaches \cite{Zhang2014} to 
 approximate methods and found that the two perform in a comparable fashion, without large differences between
@@
 in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
-This leads to the conclusion that both types of methods (exact and approximate) should be considered. 
-Therefore, in this work, we introduce a unified paradigm, namely quadratic programming, that enables
-both exact and approximate solutions to WFLO. Thus, 
-our approach `enjoys' the best of both worlds (the exact and the approximate).
+These results lead to the conclusion that both types of methods (exact and approximate) should be considered and, therefore, we introduce a unified paradigm, namely quadratic programming, that enables
+both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
+recent advanced exact software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
+(e.g., Fujitsu's Digital Annealer).
+%Thus, 
+%our approach `enjoys' the best of both worlds (the exact and the approximate).
 
-In this section,
-we formulate the WFLO problem
-using quadratic programming. Specifically, we consider two
+%In this work, we argue that
+
+%that solves quadratic problems using principles from simulated annealing).
+
+%% In this section,
+%% we formulate the WFLO problem
+%% using quadratic programming. Specifically,
+We consider two
 quadratic optimization problems that represent WFLO: a
-quadratic constrained optimization (QCOP) problem and 
+quadratic constrained optimization problem (QCOP) and 
 a quadratic unconstrained binary optimization (QUBO) problem. 
 The former represents WFLO characteristics (proximity and number of turbines)
-as hard constraints, while the latter considers these characteristics as soft constraints that are added to its objective function.
+as hard constraints, while the latter considers these characteristics as soft constraints that are penalized as part of the objective function.
 
@@
 
-\subsection{WFLO Inputs and Decision Variables} We consider the wind farm model that we presented in Section~\ref{sec:related}. 
-Let $x_i$ be a binary decision variable that represents whether a 
+\subsection{WFLO Inputs and Decision Variables}
+Given the wind farm problem presented in Section~\ref{sec:related},
+let $x_i$ be a binary decision variable that represents whether a 
 turbine is positioned at location 
@@
 of $k = n^2$ possible turbine locations (cells in the grid). We 
-write $x$ as shorthand for $x = (x_1,\ldots, x_k)$. Furthermore,
-we denote by $\mathcal{E} \subseteq \mathcal{N}\times \mathcal{N}$
+write $x$ as shorthand for $x = (x_1,\ldots, x_k)$. 
+We denote by $\mathcal{E} \subseteq \mathcal{N}\times \mathcal{N}$
 the set of location pairs that cannot simultaneously host turbines 
 due to proximity constraints. The set $\mathcal{E}$ can be pre-computed 
-based on problem specifications. For example,
-it is common to assume that two turbines must be placed at least five rotor diameters apart.
+based on problem specifications.
+%For example,
+%it is common to assume that two turbines must be placed at least five rotor diameters apart.
 
@@
 given that a turbine is placed at cell $i$ (i.e., turbines placed in these locations 
-will be influenced by a turbine placed in $i$). Finally,
+will influence a turbine placed in $i$). Finally,
 we denote by $u_{ijd}$ the wind speed at location $j$ due to a single wake from upstream turbine at location $i$ with $j \in \mathcal{U}_{id}$. 
@@
 
-\subsection{WFLO as QCOP}
+\subsection{WFLO as a Quadratic Constrained Optimization Problem}
 
@@
 The QCOP presented as Problem~\ref{QCOP} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
-So far, in~\cite{Zhang2014} and~\cite{donovan2005wind}, equivalent (linear) problems were formulated and solved using 
-exact solvers. However, these solvers should only be used if the problem is computationally hard.
+Similar, linear formulations have been previously proposed and solved using exaxt solvers~\cite{Zhang2014,donovan2005wind}.
+\todo{Are these linear formulations ``equivalent'' as you claimed or approximations?
+If they are mathematically equivalent, why do we think a quadratuc model will be any better computationally?}
+
+\subsection{WFLO as a Quadratic Unconstrained Binary Optimization Problem}
+
+Quadratic unconstrained binary optimization problems, as their name implies, do not allow the direct representation of hard constraints. Instead, 
+%they only 
+%have an objective function that 
+%one can to maximize (or minimize) and
+problem constraints must be represented as part of the objective function
+using penalty terms.
+
+Let $\lambda = (\lambda_1,\lambda_2)$ 
+be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
+\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
+with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
+penalizing solutions that violate the exactly $m$ turbines constrain and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
+solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
+must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints.
+%These types of constraints
+%are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
+%for an arbitrary selection of $\lambda$.
+
+\subsection{The Computational Complexity of WFLO}
+
+%(linear) problems were formulated and solved using  exact solvers.
+Despite the existence of
+
+However, these solvers should only be used if the problem is computationally hard.
 We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP). In what follows, we provide
@@
 we reformulate the QCOP into a QUBO problem, which allows us to solve WFLO using specialized optimization hardware.
-
-\subsection{WFLO as QUBO}
-
-Quadratic unconstrained binary optimization problems, as their name implies, prohibit hard constraints. Instead, 
-they only 
-have an objective function that 
-one can to maximize (or minimize) and constraints must be inserted into the objective function
-using penalty terms.
-
-Let $\lambda = (\lambda_1,\lambda_2)$ 
-be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq.(\ref{QCOP}) is given by,
-\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
-with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
-penalizing solutions that violate the exactly $m$ turbines constraint, while the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizes
-solutions that place turbines in high proximity. The two penalty terms $\lambda_1$ and $\lambda_2$ 
-must be large enough to ensure that the solutions are indeed feasible. These types of constraints
-are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
-for an arbitrary selection of $\lambda$.
 

### 2020-04-22 paper.tex
Finished edit up to end of Sect 3.4

@@
 &\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
-&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
+&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars}
 \end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
@@
 The QCOP presented as Problem~\ref{QCOP} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
-Similar, linear formulations have been previously proposed and solved using exaxt solvers~\cite{Zhang2014,donovan2005wind}.
+Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
 \todo{Are these linear formulations ``equivalent'' as you claimed or approximations?
-If they are mathematically equivalent, why do we think a quadratuc model will be any better computationally?}
+If they are mathematically equivalent, why do we think a quadratic model will be any better computationally?}
 
@@
 %(linear) problems were formulated and solved using  exact solvers.
-Despite the existence of
+Despite the existence of mathematical models for WFLO in the literature~\cite{Zhang2014,donovan2005wind} that are expressive enough to represent $\mathcal{NP}$-hard problems, we are unaware of an explicit treatment of the computational complexity of WFLO.
+%However, these solvers should only be used if the problem is computationally hard.
+%We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP).
+We therefore provide a straightforward proof of the computational complexity of the QCOP defined above.
 
-However, these solvers should only be used if the problem is computationally hard.
-We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP). In what follows, we provide
-a theorem that proves the computational complexity of the LS-based WFLO QCOP. 
 \begin{mythm}
-	The computational complexity of QCOP is $\mathcal{NP}$-hard. 
+	The computational complexity of the QCOP defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
 \end{mythm}
 \begin{myproof}
-We prove by showing that a special case of the WFLO problem (Eq.~(\ref{QCOP}) is
+We prove by showing that a special case of the QCOP \eqref{QCOP}-\eqref{Vars}  is
 $\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
-are dropped and we get the heaviest k-subgraph problem (HSP), which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
+are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
 Since a special case of WFLO is  $\mathcal{NP}$-hard it implies that WFLO is at least $\mathcal{NP}$-hard. On the other hand,
 WFLO is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFLO is $\mathcal{NP}$-hard.
-	\end{myproof} To overcome the challenge associated with the computational hardness of WFLO,
-we reformulate the QCOP into a QUBO problem, which allows us to solve WFLO using specialized optimization hardware.
+\end{myproof}
+
+\todo{The proof is a bit awkward as it goes back-and-forth between QCOP and WFLO.
+Technically, the problem as defined in \eqref{QCOP}-\eqref{Vars} but, given the different (not mathematically equivalent) models of WFLO, this is different from claiming that WFLO is NPC.
+On the other hand, we so not want to prove that QCOP in general is NPC.
+So we need to get the language precise.}
+
+%To overcome the challenge associated with the computational hardness of WFLO,
+%we reformulate the QCOP into a QUBO problem, which allows us to solve WFLO using specialized optimization hardware.
 

### 2020-04-22 paper.tex
minor.

--- 2020-04-22 paper.tex
@@
 \title{\mytitle}
-
 %\subtitle{Extended Abstract}
@@
 %role as a source of renewable energy and 
-  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference among turbines.
-Given the pairwise wake interactions, it is natural
-%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
-to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization models.
-Recent advances in both optimization hardware and software have targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
+The wind farm layout optimization (WFLO) problem requires the
+selection of wind turbine positions to maximize energy production.
+While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is common to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models.
+Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems.
 %It is a key problem 
@@
 A thorough empirical 
-evaluation using a commercial solver and specialized QUBO hardware
-%demonstrates the strengths of the two models. 
-%Our results
-show that both approaches yield fast, high-quality solutions 
-that
-improve the state of the art.
+evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
+Our results show that both approaches yield fast and high-quality solutions 
+and
+improve over existing WFLO solutions.
 In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
@@
 interference effects (i.e., wakes) generated by upstream turbines.
-These inter-turbine effects are best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
+These inter-turbine effects 
+is best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
 using a sum-of-squares expression~\cite{Zhang2014}. However,
@@
 that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
-An alternative formulation of WFLO relies on a less accurate wake model, which uses linear superpositioning
+An alternative formulation of WFLO relies on a less accurate wakes model, which uses linear superpositioning
 to capture multiple wakes~\cite{donovan2005wind}.
@@
 approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
-represent the quadratic relations using approximate linear models,
-%that approximate wake dependencies,
-as, until recent years, solving quadratic programming problems
-was computationally challenging compared to linear models.
+represent the quadratic relations using linear models
+that approximate wake dependencies, since until recent years solving linear programming problems
+was computationally faster.
 
 In this work, we introduce two novel quadratic WFLO models. The first
-model represents WFLO as a quadratic constrained optimization
-problem (QCOP).  The QCOP can be solved quickly using state-of-the-art mathematical solvers
-that have been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
-second model is a quadratic unconstrained binary optimization
+model represents WFLO with a quadratic constrained optimization
+problem (QCOP).  The QCOP can be solved quickly using state-of-the-art solvers
+that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
+second model is based on a quadratic unconstrained binary optimization
 (QUBO) representation of the WFLO problem.  Such a formulation enables the use of
-nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics} here.
-%, a hardware-based
-%QUBO solver, to demonstrate the strength of our approach.
-
-The main
+specialized optimization hardware tailored to solve QUBOs. In this
+paper, we use Fujitsu's Digital Annealer (DA), a hardware-based
+QUBO solver, to demonstrate the strength of our approach. The main
 contributions of this paper are:
 \begin{enumerate} 
-\item We propose a novel quadratic modeling framework for WFLO that encompasses both constrained and unconstrained quadratic optimization models.
- %that captures wake 
-%effects directly through their linear superposition. 
+\item We propose a novel quadratic modeling framework for WFLO that captures wake 
+effects directly through their linear superposition. 
 %without the need for approximations and ad-hoc solutions,
@@
 using Gurobi and the DA
-achieves state-of-the-art performance.
-%, often outperforming existing solutions when provided sufficient run-time.
-Furthermore, we observe that the DA 
-often provides solutions of similar quality two orders of magnitude faster
+achieves state-of-the-art performance, often outperforming
+existing solutions when provided sufficient run-time. Furthermore, we
+we observe that the DA 
+often provides solutions of similar quality an order of magnitude faster
 than software-based approaches.
 
-The remainder of this paper is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
-optimization solutions to the layout problem.
-The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO} and Section~\ref{sec:eval} demonstrates the value of our approach via a thorough empirical evaluation.
-%of the proposed methods. 
+The remainder is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
+optimization solutions to the layout problem. The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO}.
+Section~\ref{sec:eval} demonstrates the usefulness of our approach via a thorough empirical evaluation of the proposed methods. 
 Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.
@@
 We start this section by presenting the physical model of a wind farm
-%used throughout the paper.
-%Our wind farm model is
-based on the common
-notation and description presented in Zhang et al.\ \cite{Zhang2014}.  Subsequently,
+used throughout the paper.  Our wind farm model is based on the common
+notation and description presented in~\cite{Zhang2014}.  Subsequently,
 we provide a literature review of existing mathematical optimization
@@
 grid with $n$ being the number of cells on each axis.  The cells have
-an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
-$n=20$, the farm area will be $2km\times 2km$.  We
+an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
+$n=20$, the farm area will be $2km\times 2km$.  In this work, we
 assume that the terrain is flat and that we have no constraints
@@
 abstract wind farm model. Our model can be extended, in the
-future, to include additional problem characteristics.
+future, to include additional wind farm settings.
 
@@
 corresponding constraint as part of the optimization model. For
-example, if a turbine is placed in the center of a
-cell, a cell size of $c = 100$ meters and a rotor diameter
-of $40$ meters, requires that two turbines are not placed in adjacent cells. In Figure~\ref{fig:field_model}, the red area
+example, assume that a turbine is placed in the center of a
+cell. Then, given a cell size of $c = 100$ meters and a rotor diameter
+of $40$ meters, we must ensure that two turbines are not placed in
+two adjacent cells. In Figure~\ref{fig:field_model}, the red area
 around the turbine represents the forbidden cells.
@@
 \paragraph{Wakes}
-Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
-%that are
-%referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
+Turbines influence each other through interference effects that are
+referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
 Wakes cause reduction of the effective power produced by a turbine due
@@
 for which we want to compute the effect (denoted by $x$ in
-Figure~\ref{fig:field_model}), (2) the radius of the cone's opening at distance $x$ 
+Figure~\ref{fig:field_model}), (2) the radius of the cone's opening
 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
-combination of several wind regimes where a wind regime consists of 
-the wind's free speed and direction.  The wake effect can be pre-computed
+superposition of several wind regimes (a wind regime is a combination
+of free speed and directions).  The wake effect can be pre-computed
 for each pair of cells since we know: the distance ($x$), the turbine
@@
 \paragraph{Effective Power Calculations in the Presence of Wakes} 
-Let $\mathcal{D}$ be the set of wind regimes.
-%where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
+Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
 The probability of wind regime 
@@
 function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
-a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
+a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
 probability of this regime is $0.008$.
 
-Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
-regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
-Rather, $u_{id}$, can be computed using the following equation~\cite{Zhang2014}:
+Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
+Note that turbines can adjust their direction according to the current wind 
+regime.
+
+Given the interference of the wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
+The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
 \begin{equation}
@@
 the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
-the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
-%~(\cite{Zhang2014}):
+the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}):
 \begin{equation}
@@
   Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
-  However, the expression is challenging to exact mathematical optimization techniques due to its complexity (i.e., the cube of a square root).
-  %since it is a squared root function of a cubic expression. 
-  %Hence, a simplified representation for the total expected energy that considers wakes is required.
+  However, the expression is not amenable to optimization since it is a squared root function of a cubic expression. 
+  Hence, a simplified representation for the total expected energy that considers wakes is required.
 
@@
 The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
-naturally results in a quadratic objective function in our WFLO formulations. Such functions have been the subject of recent research advances \todo{[REFs]}, motivating the contributions here.
-%, which is a desired property from an optimization standpoint.
+results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.
 \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic.
@@
 The exact methods were shown to yield higher energy values, yet they often suffered from 
-high computational cost and long run-times.
-In practice, it is often desirable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.
+high computational cost and long runtimes.
+In practice, it is often desireable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.
 
@@
 work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
-only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see Zhang et al.\ \cite{Zhang2014}.}
+only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see~\cite{Zhang2014}.}
 
@@
 the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
-MILP approach based on the LS energy expression~\cite{donovan2005wind}.
-Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation.
-The two resulting methods performed well
-compared to existing approximate and exact techniques.
+MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
+approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
+compared to existing approximate and earlier exact solutions.
 \todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}
@@
 The most recent attempt to solve the wind farm variation that we presented 
-in this section is given in Zhang et al.\ \cite{Zhang2014}. The authors compared
+in this section is given in~\cite{Zhang2014}. The authors compared
 previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
@@
 
-In our work, we build upon the approach in Donovan~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
-model WFLO using a quadratic representation. %The difference between our approach and the quadratic model
-%proposed in
-Unlike Turner et al.\ \cite{turner2014new} our model is not an approximation of the SS objective, 
+In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
+model WFLO using a quadratic representation. The difference between our approach and the quadratic model
+proposed in~\cite{turner2014new} is that our model is not an approximation of the SS objective, 
 but an exact representation of the LS objective.
 
-While the main advantage of exact techniques
-is that they provide guarantees on solution quality, they may take a very long time to  find an optimal solution. 
+As we mentioned before, the main advantage of exact techniques
+is that they provide guarantees on solution quality. However, their disadvantage is
+that they may take a very long time to execute before finding an optimal solution. 
 In fact, Section~\ref{sec:QUBO4WFLO} shows that WFLO is computationally hard, a result
@@
 \paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
-quick and, hopefully, sufficiently accurate solutions. Several works solved 
+quick and sufficiently accurate solutions. Several works solved 
 WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
@@
 to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
-(i.e., they may run forever without achieving a quality criteria), and even when they do terminate (in practical cases) they do
+(i.e., they may run forever without achieving a quality criteria), and when they do terminate (in practical cases) they do
 not guarantee optimal solutions. Hence, one must often
@@
 when a stopping criteria is reached.
-The main limitation of this approach 
-%~\cite{ozturk2004heuristic}
-is that it cannot perform an action 
-that leads to an infeasible state and, therefore, quickly converges to a sub-optimal solution (local optimum) 
-that can potentially be much lower quality than the globally optimal solution~\cite{rivas2009solving}.
+The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
+that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
+that can potentially be far
+from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.
 
-To overcome this limitation, Rivas et al.\ \cite{rivas2009solving} used simulated annealing (SA), 
+To overcome this limitation, the authors of~\cite{rivas2009solving} used simulated annealing (SA), 
 a neighborhood search method that allows visits to infeasible solutions. 
 This creates a well-connected solution space that leads the algorithm to escape 
-local optima. The SA method was shown to be superior to Ozturk \& Norman~\cite{ozturk2004heuristic} 
-and to the genetic method proposed in Grady et al.\ \cite{grady2005placement}. A drawback of the SA approach 
+local optima. The SA method was shown to be superior to~\cite{ozturk2004heuristic} 
+and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach 
 is its ad-hoc nature. The various components of the algorithm
- must be tailored to the problem at hand and, as a result, small adjustments of the WFLO setting
- (e.g., adding noise considerations), would require major changes to the SA implementation. 
-Additional methods and experimental comparisons between various approximate algorithms are available in Samorani~\cite{samorani2013wind}.
+ must be tailored and geared towards WFLO. This means that small adjustments of the WFLO setting
+ (e.g., adding noise considerations), would require major changes to the SA solution. 
+Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.
 
-In our work, we also use an SA approach (similarly to Rivas et al.\ \cite{rivas2009solving}). However,
-our methodology is generic since it is based on a declarative problem model and
+In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
+our methodology is generic since it is based on a model and
 on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.
@@
 formulating WFLO using quadratic programming enables us to employ 
-recent advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
-(e.g., Fujitsu's Digital Annealer).
-%that solves quadratic problems using principles from simulated annealing). 
+advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
+(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
 A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
@@
 		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
-		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
+		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.10  & 15739.57     \\
 		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*           \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80}*           \\
 		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
-		&&40   & 21874.04 & 21643.56  & 21905.5 & \textbf{21977.17} & \textbf{21977.17}        \\
+		&&40   & 21874.04 & 21643.56  & 21905.50 & \textbf{21977.17} & \textbf{21977.17}        \\
 		\hline
 		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
-		&& 30  & \textbf{27450.99} & 27418.74 & 27442.9 & 27387.06 & 27247.5                     \\
+		&& 30  & \textbf{27450.99} & 27418.74 & 27442.90 & 27387.06 & 27247.50                     \\
 		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
@@
 		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
-		&&30   & \textbf{27975.55} & 27443.8  & 27752 & 27492.11  & 26367.2                      \\
-		&&40   & 35604.45 & 35127.32 & \textbf{35665.6}   & 35065.85 & 33451.94 \\
+		&&30   & \textbf{27975.55} & 27443.80  & 27752.00 & 27492.11  & 26367.20                      \\
+		&&40   & 35604.45 & 35127.32 & \textbf{35665.60}   & 35065.85 & 33451.94 \\
 		\bottomrule                   
@@
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8} *          \\
-		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.2 }                 \\
-		&&40   & 21874.04 & 21832.5  & 21963.81 & \textbf{21977.17} & \textbf{21977.17}        \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80} *          \\
+		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.20 }                 \\
+		&&40   & 21874.04 & 21832.50  & 21963.81 & \textbf{21977.17} & \textbf{21977.17}        \\
 		\hline
 		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19213.96 \\
-		&& 30  & \textbf{27450.99} & 27442.9 & 27442.9 & 27387.06 & 27389.21                     \\
+		&& 30  & \textbf{27450.99} & 27442.90 & 27442.90 & 27387.06 & 27389.21                     \\
 		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34856.45          \\
@@
 		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19394.58 & 19336.33 & 19354.23  & 19368.86            \\
-		&&30   & \textbf{27975.55} & 27785.74  & 27752 & 27492.11  & 27633.2                      \\
-		&&40   & 35604.45 & 35480.9 & \textbf{35665.6}   & 35065.85 & 35210.78 \\
+		&&30   & \textbf{27975.55} & 27785.74  & 27752.00 & 27492.11  & 27633.20                      \\
+		&&40   & 35604.45 & 35480.90 & \textbf{35665.60}   & 35065.85 & 35210.78 \\
 		\bottomrule                   
@@
 
-uwo\begin{figure}[t]%
+\begin{figure}[t]%
 

### 2020-04-25 paper.tex
fixes in 3, pass over 4 and 5.

--- 2020-04-22 paper.tex
@@
 %without the need for approximations and ad-hoc solutions,
-\item We prove that WFLO is $\mathcal{NP}$-hard.
+\item We prove that the QCOP variation of WFLO is $\mathcal{NP}$-hard.
 \item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
@@
 often provides solutions of similar quality two orders of magnitude faster
-than software-based approaches.
+than software-based approaches. The ability to quickly obtain competitive solutions on large WFLO instances 
+can be employed to solve
+multiple wind farm design problems.
 
@@
 The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
-naturally results in a quadratic objective function in our WFLO formulations. Such functions have been the subject of recent research advances \todo{[REFs]}, motivating the contributions here.
+naturally results in a quadratic objective function in our WFLO formulations.
+Such functions have been the subject of recent research advances~\cite{ku2017hybrid,bian2010ising}, motivating the contributions here.
 %, which is a desired property from an optimization standpoint.
@@
 Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation.
-The two resulting methods performed well
+The two resulting methods performed well when
 compared to existing approximate and exact techniques.
-\todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}
+Another non-linear optimization approach was proposed in Ulku and Alabas-Uslu\ \cite{ulku2019new}.
+Ulku and Alabas-Uslu\ \cite{ulku2019new}, replaced integer placement variables with continuous variables between $0$ and $1$,
+and approximated the constraints such that the resulting variables yield an integer placement.
+Their approach provides yet another approximation to the WFLO problem considered in~\cite{turner2014new}.
 
@@
 %proposed in
-Unlike Turner et al.\ \cite{turner2014new} our model is not an approximation of the SS objective, 
+Unlike Turner et al.\ \cite{turner2014new} nor Ulku and Alabas-Uslu\ \cite{ulku2019new}, our model is not an approximation of the SS objective, 
 but an exact representation of the LS objective.
@@
 \end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
-and enforce proximity constraints between the turbines, respectively.
+and enforce proximity constraints between the turbines, respectively. We shall refer to the QCOP
+formulation of WFLO as WFL-QCOP.
 
-The QCOP presented as Problem~\ref{QCOP} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
+WFL-QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
 Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
 \todo{Are these linear formulations ``equivalent'' as you claimed or approximations?
-If they are mathematically equivalent, why do we think a quadratic model will be any better computationally?}
+If they are mathematically equivalent, why do we think a quadratic model will be any better computationally?
+Arik: The two formulations are equivalent to WFL-QCOP.
+However, we compare against the ILP model experimentally and it does worse than WFL-QCOP.
+Hopefully, Jason's part in Section 3.5.1.
+based on Wen-Yang's thesis will explain why it does better.
+Originally, I used the quadratic formulation because it seemed more natural.
+Semantically, the linear model and the quadratic model are the equivalent imho.}
 
@@
 solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
-must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints.
+must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as WFL-QUBO.
 %These types of constraints
@@
 
-\subsection{The Computational Complexity of WFLO}
+\subsection{The Computational Complexity of WFL-QCOP}
 
@@
 %We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP).
-We therefore provide a straightforward proof of the computational complexity of the QCOP defined above.
+We therefore provide a straightforward proof of the computational complexity of WFL-QCOP.
 
 \begin{mythm}
-	The computational complexity of the QCOP defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
+	The computational complexity of WFL-QCOP defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
 \end{mythm}
 \begin{myproof}
-We prove by showing that a special case of the QCOP \eqref{QCOP}-\eqref{Vars}  is
+We prove by showing that a special case of the WFL-QCOP \eqref{QCOP}-\eqref{Vars}  is
 $\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
 are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
-Since a special case of WFLO is  $\mathcal{NP}$-hard it implies that WFLO is at least $\mathcal{NP}$-hard. On the other hand,
-WFLO is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFLO is $\mathcal{NP}$-hard.
+Since a special case of WFL-QCOP is  $\mathcal{NP}$-hard it implies that WFL-QCOP is at least $\mathcal{NP}$-hard. On the other hand,
+WFL-QCOP is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFL-QCOP is $\mathcal{NP}$-hard.
 \end{myproof}
@@
 On the other hand, we so not want to prove that QCOP in general is NPC.
-So we need to get the language precise.}
+So we need to get the language precise.
+Arik: I fixed it by referring to the problem as WFL-QCOP (and WFL-QUBO for the other one.)}
 
@@
 Our main results are: \begin{itemize}
-	\item QUBO based WFLO using the digital annealer (DA) achieves a similar performance to state-of-the-art exact methods.
-		\item The runtime for solving the QUBO model on the DA is two orders of magnitude lower than QCOP.  
-	\item Given sufficient runtime, solving the QCOP model yields superior results on the larger instances.
+	\item WFL-QUBO when solved using the digital annealer (DA) achieves a similar performance to state-of-the-art exact methods.
+		\item The runtime for solving WFL-QUBO on the DA is two orders of magnitude lower than WFL-QCOP.  
+	\item Given sufficient runtime, solving the WFL-QCOP model yields superior results on the larger instances.
 \end{itemize}
 We start by outlining the experimental design of our evaluation, followed by an overview of our main results,
-and conclude the section with a discussion of the results together with the limitations of our approach.
+and conclude the section with a discussion of the results and their practical implications.
 
@@
 
-We compared our two models, namely QCOP and QUBO, against two state of the art exact optimization methods:
+We compared our two models, namely WFL-QCOP and WFL-QUBO, against two state of the art exact optimization methods:
 the integer linear programming (ILP-LS) approach from~\cite{donovan2005wind}, and the 
@@
 reported to perform worse than the rest and was left out of the experiments.
-We used state of the art software, namely Gurobi~\cite{gurobi}, to solve QCOP, ILP-SS and QP-SS. 
-We shall refer to the QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
+We used state of the art software, namely Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-SS and QP-SS. 
+For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
 In addition, we 
-applied both Fujitsu's digital annealer and Gurobi to solve the QUBO variant of WFLO. We shall refer to the former as DA-LS, and to the latter
+applied both Fujitsu's digital annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
 as QUBO-LS.
@@
 Further, denote by $b_{w,t,a}$ denote the best solution attained before time $t$ of approach $a$ for instance $w$.
-Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.\end{equation} We can now average over the different instances at time $t$.
+Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.
+
+We can now average over the different instances at time $t$.
 The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
@@
 
-\subsection{Discussion}
+\subsection{Discussion and Practical Implications}
 
 Our result show empirically that the quadratic programming models that we
-introduced in Section~\ref{sec:QUBO4WFLO} (QCOP and QUBO) are useful in two ways.
-First, QCOP and QUBO are the basis for the QP-LS and QUBO-LS methods that couple
+introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
+First, the two problems are the basis for the QP-LS and QUBO-LS methods that couple the two
 quadratic formulations of WFLO with Gurobi optimizer. This turns out
@@
 The ability of DA-LS to quickly obtain competitive solutions on large WFLO instances 
-can be employed to solve
-multiple wind farm design problems.
-This allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that came
-(so far) with re-solving the problem. In addition, the performance loss due to the use of an approximate method 
-that uses advanced technology is negligible. This balance (instead of trade-off) between
-runtime and total energy, points at the usefulness of the proposed quadratic formulations of WFLO.
-\todo{maybe we should say some of these things in the intro and bring citations to the fact that demand for wind farms increases.
-I found some papers about surge in China and the UK.}
+is crucial when solving
+multiple wind farm layout problems.
+Moreover, it allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that came
+with (re-)solving the problem using exact methods. In addition, the performance loss due to the use of an approximate method 
+that exploits advanced technology is negligible. This balance (instead of trade-off) between
+runtime and total energy, points at the potential contribution of the proposed quadratic framework for solving WFLO.
 
@@
 wind farm layout optimization. 
-The framework is based on two models: a constrained model (QCOP) and an unconstrained binary
-model (QUBO). The former can be used to solve WFLO with advanced optimization software,
-while the latter can be mounted onto quantum-inspired dedicated hardware that can quickly (yet approximately) solve WFLO. We used the constrained quadratic formulation to prove 
+The framework is based on two models: a constrained model (WFL-QCOP) and an unconstrained binary
+model (WFL-QUBO). The former can be used to solve WFLO using advanced optimization software,
+while the latter can be mounted onto quantum-inspired machines that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
 that the computational complexity of WFLO is $\mathcal{NP}$-hard, a pessimistic result that 
 was also noticed experimentally in~\cite{Zhang2014}. However,
-our empirical evaluation shows that when advanced exact solvers like Gurobi~\cite{gurobi} are
-coupled with our QCOP and QUBO models they can solve small WFLO instances to optimality in less than 3600 seconds.
-We also show that when an advanced hardware is used with the QUBO representation of WFLO, even large instances can be solved quickly with no significant decrease in the total energy compared to state-of-the-art exact methods.
-This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our proposed quadratic framework useful for designing new wind farms, while multiple input parameters.
+our empirical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
+coupled with our WFL-QCOP and WFL-QUBO models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
+We also show that when an advanced hardware is used with the QUBO representation, large instances WFLO can be solved quickly with no significant degradation in the total energy compared to exact methods.
+This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our quadratic framework useful for designing new wind farms, while varying multiple input parameters.
 
@@
 Moreover, we wish to explore real-world wind farm design problems 
-that may be larger and more complex than the 12 standard instances, and solve them 
+that may be larger and more complex than the 12 standard instances 
+and solve them 
 using our quadratic programming approach.

### 2020-04-29 paper.tex
QP subsection + Machine specs + Evaluation - Jason

@@
+Skip to content
+Search or jump to…
+
+Pull requests
+Issues
+Marketplace
+Explore
+
+@JasonZhangjc 
+Learn Git and GitHub without any code!
+Using the Hello World guide, you’ll start a branch, write comments, and open a pull request.
+
+ArikSenderovich1983
+/
+WFLO
+1
+00
+ Code Issues 0 Pull requests 0 Actions Projects 0 Wiki Security 0 Insights
+WFLO/paper.tex
+@ArikSenderovich1983 ArikSenderovich1983 fixes in 3, pass over 4 and 5.
+8cbc7fe 8 hours ago
+@ArikSenderovich1983@jcb-uoft@JasonZhangjc
+908 lines (744 sloc)  54.1 KB
+
 \documentclass[preprint,12pt]{elsarticle}
@@
 \subsubsection{QCOP solving methods}
+WFLO is a type of strictly convex integer quadratically constrained optimization problem (IQCOP), which is known to be NP-hard \cite{van1981another}. The common exact methods for solving
+IQCOPs are tree search \cite{land2010automatic} type algorithms with a variety of techniques to prune sub-trees that definitely do not contain the optimal solution.
 
-\todo{Jason's part}
+In operations research, the generic exact approach for solving IQCOPs are the branch-and-cut based mixed integer programming (MIP) solvers \cite{bonami2008algorithmic}.
+By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.
+
+Constraint programming (CP) \cite{rossi2006handbook} is another tree search framework that can be applied to solve IQCOPs.
+At each node in the search tree, inference algorithms that are often encapsulated in the global constraints \cite{katriel20066}, conduct domain filtering on each constraint.
+Domain filtering removes possible values from variable domains by proving that a value cannot satisfy the constraint and hence cannot appear in a global solution, hence pruning sub-trees with variable assignments that are not part of any feasible solution.
+
+In the community of communications, IQCOPs are commonly solved via discrete ellipsoid-based search (DEBS), which is built on enumeration of integer points in the hyper-ellipsoid defining the feasible region \cite{agrell2002closest}.
+DEBS can be interpreted as a form of CP, where the ellipsoid geometry induces for each variable an interval domain, producing an inference algorithm removing values that do not belong to the ellipsoid.
+However, DEBS cannot be applied to general IQCOPs due to its special nature.
+This drawback has been partially overcome by Wen-Yang \cite{ku2017hybrid}, who has proposed a competitive hybrid algorithm that combines DEBS and MIP to solve IQCOPs.
+
+% In addition to exact methods, heuristics that are designed to find good but not provably optimal solutions fast can also be used to solve IQCOPs.
+
+\todo{@Arik and Chris: please correct Jason's writing errors}
 
@@
 values (SS in QP-SS and LS in the rest of the methods).
-For the digital annealer, we used a single configuration that was found to work best across the different instances \todo{Jason, please insert details about the DA config.
-here} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime
+For the digital annealer, we used a single configuration that was found to work best across the different instances.
+The initial temperature is 500000, the temperature mode is 0, the  interval for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically.
+\todo{Eldan and Arik: Do I need to write the formula for computing temperature decay rate?
+-- Jason} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime
 of the procedure was recorded.
@@
 
-The exact methods were all solved using Gurobi's quadratic programming solver on \todo{Jason, please add machine details here}.
+The exact methods were all solved using Gurobi's quadratic programming solver on a Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
 Furthermore, QUBO-LS was solved on DA version 1 \todo{Eldan: any additional info that we need to add here?}.
@@
 We shall present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
-\todo{Jason: please check that this is consistent with how you calculated MRE in the experiment.}
 
@@
 		\toprule
-		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
 		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.10  & 15739.57     \\
@@
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80}*           \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
 		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method.
-* denotes optimal solution.
-\todo{@Jason: could you please check if my marking of * for optimal solutions is correct?
-I think that optimality was proven only for $m=20$ and single wind direction, right?} }\label{tab:results1}
+	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method. \todo{@Arik: could you please why we do not mention any optimality here?} }\label{tab:results1}
 \end{table}
@@
 		\toprule
-		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
 		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
@@
 		\cline{2-8}
-		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80}*  & \textbf{11404.80} *          \\
+		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
 		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.20 }                 \\
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method. * denotes optimal solution.}\label{tab:results2}
+	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method.}\label{tab:results2}
 \end{table}
@@
 %\end{table}
+
+© 2020 GitHub, Inc.
+Terms
+Privacy
+Security
+Status
+Help
+Contact GitHub
+Pricing
+API
+Training
+Blog
+About

### 2020-04-29 paper.tex
minor

@@
-Skip to content
-Search or jump to…
-
-Pull requests
-Issues
-Marketplace
-Explore
-
-@JasonZhangjc 
-Learn Git and GitHub without any code!
-Using the Hello World guide, you’ll start a branch, write comments, and open a pull request.
-
-ArikSenderovich1983
-/
-WFLO
-1
-00
- Code Issues 0 Pull requests 0 Actions Projects 0 Wiki Security 0 Insights
-WFLO/paper.tex
-@ArikSenderovich1983 ArikSenderovich1983 fixes in 3, pass over 4 and 5.
-8cbc7fe 8 hours ago
-@ArikSenderovich1983@jcb-uoft@JasonZhangjc
-908 lines (744 sloc)  54.1 KB
-
 \documentclass[preprint,12pt]{elsarticle}

### 2020-04-29 paper.tex
minor edits.

@@
 \subsection{The Computational Complexity of WFL-QCOP}
-
+\label{sec:computational}
 %(linear) problems were formulated and solved using  exact solvers.
@@
 \subsubsection{QCOP solving methods}
-WFLO is a type of strictly convex integer quadratically constrained optimization problem (IQCOP), which is known to be NP-hard \cite{van1981another}. The common exact methods for solving
+WFL-QCOP is a type of strictly convex integer 
+quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}. In fact,
+we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard. The common exact methods for solving
 IQCOPs are tree search \cite{land2010automatic} type algorithms with a variety of techniques to prune sub-trees that definitely do not contain the optimal solution.
@@
 
-\todo{@Arik and Chris: please correct Jason's writing errors}
+\todo{@Chris and Jason: I still feel that we need to say how Gurobi do it (or at least how we think Gurobi does that)}
 
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method. \todo{@Arik: could you please why we do not mention any optimality here?} }\label{tab:results1}
+	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method.}\label{tab:results1}
 \end{table}

### 2020-04-29 paper.tex
Edited abstract and introduction

@@
 %role as a source of renewable energy and 
-  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference among turbines.
+  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and a model of the wind patterns, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference among turbines.
 Given the pairwise wake interactions, it is natural
@@
 %that approximate wake dependencies,
-as, until recent years, solving quadratic programming problems
-was computationally challenging compared to linear models.
+as solving quadratic programming problems
+is computationally challenging compared to linear models.
 
@@
 model represents WFLO as a quadratic constrained optimization
-problem (QCOP).  The QCOP can be solved quickly using state-of-the-art mathematical solvers
+problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical solvers
 that have been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
@@
 %, often outperforming existing solutions when provided sufficient run-time.
-Furthermore, we observe that the DA 
-often provides solutions of similar quality two orders of magnitude faster
-than software-based approaches. The ability to quickly obtain competitive solutions on large WFLO instances 
-can be employed to solve
-multiple wind farm design problems.
+Furthermore, we observe that the DA often provides solutions of
+similar quality to software-based approaches but with two orders of
+magnitude lower run-time.
+%The ability to quickly obtain competitive
+%solutions on large WFLO instances can be employed to solve multiple
+%wind farm design problems.
 

### 2020-04-29 paper.tex
Half way through editting of Section 2.

@@
 %\paragraph{Wind Farm as a Grid}
-We consider a square-shaped wind farm that we model as an $n\times n$
+We consider a square wind farm that we model as an $n\times n$
 grid with $n$ being the number of cells on each axis.  The cells have
@@
 We assume that the number of identical turbines to be placed is known 
-in advance and denote the number of turbines by $m$. To place these $m$ turbines, 
+in advance and denoted by $m$. To place these $m$ turbines, 
 we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
@@
   Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
-  However, the expression is challenging to exact mathematical optimization techniques due to its complexity (i.e., the cube of a square root).
+  However, the expression is challenging for exact mathematical optimization techniques due to its complexity (i.e., the cube of a square root).
   %since it is a squared root function of a cubic expression. 
@@
 Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
-The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
-naturally results in a quadratic objective function in our WFLO formulations.
+The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO}, it 
+ results in a quadratic objective function in our WFLO formulations.
 Such functions have been the subject of recent research advances~\cite{ku2017hybrid,bian2010ising}, motivating the contributions here.
@@
 compared to existing approximate and exact techniques.
-Another non-linear optimization approach was proposed in Ulku and Alabas-Uslu\ \cite{ulku2019new}.
-Ulku and Alabas-Uslu\ \cite{ulku2019new}, replaced integer placement variables with continuous variables between $0$ and $1$,
+Another non-linear optimization approach \cite{ulku2019new}, relaxed the binary placement variables to continuous values in $[0,1]$, 
 and approximated the constraints such that the resulting variables yield an integer placement.
-Their approach provides yet another approximation to the WFLO problem considered in~\cite{turner2014new}.
+Their approach provides another approximation to the WFLO problem considered by Turner et al.\ \cite{turner2014new}.
 
-The most recent attempt to solve the wind farm variation that we presented 
-in this section is given in Zhang et al.\ \cite{Zhang2014}. The authors compared
-previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
-is an optimization technique that does not make assumptions on the functional form of the objective.
-However, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.
+Zhang et al.\ \cite{Zhang2014} developed a constraint programming (CP) approach to the problem variation presented above.
+CP is an optimization technique that has grown from the AI literature that does not make the assumptions on the functional form of the constraints or objective \cite{rossi2006handbook}.
+The authors compared existing MILP and QP models to their approach for directly optimizing the SS objective and showed that, while promising for smaller problems, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.
+
+%% The most recent attempt to solve the wind farm variation that we presented 
+%% in this section is given in Zhang et al.\ \cite{Zhang2014}. The authors compared
+%% previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
+%% is an optimization technique that does not make assumptions on the functional form of the objective.
+%% However, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.
 
@@
 %proposed in
-Unlike Turner et al.\ \cite{turner2014new} nor Ulku and Alabas-Uslu\ \cite{ulku2019new}, our model is not an approximation of the SS objective, 
-but an exact representation of the LS objective.
+Unlike Turner et al.\ \cite{turner2014new} and Ulku and Alabas-Uslu\ \cite{ulku2019new}, our model is not an approximation of the SS objective, 
+but an rather an exact representation of the LS objective.
 

### 2020-04-29 paper.tex
Added comment on the approximations.
Finished edit of Section 2

@@
 Unlike Turner et al.\ \cite{turner2014new} and Ulku and Alabas-Uslu\ \cite{ulku2019new}, our model is not an approximation of the SS objective, 
-but an rather an exact representation of the LS objective.
+but rather an exact representation of the LS objective.
+That is, while the previous work has approximated a more accurate physical model, we are exactly representing the less accurate physical model: the approximation arises at a different stage of the modeling pipeline.
 
@@
 is that they provide guarantees on solution quality, they may take a very long time to  find an optimal solution. 
-In fact, Section~\ref{sec:QUBO4WFLO} shows that WFLO is computationally hard, a result
+In fact, Section~\ref{sec:QUBO4WFLO} shows formally that WFLO is computationally hard, a result
 that we have not yet encountered in the WFLO literature. Therefore,
-when one considers solving multiple wind farm design problems, a fast solution is of essence,
-and hence one may turn to approximate WFLO.
+when one considers solving multiple wind farm design problems, if a fast solution is of essence,
+one may turn to approximate optimization techniques.
 
 \paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
-quick and, hopefully, sufficiently accurate solutions. Several works solved 
+quick and, hopefully, sufficiently accurate solutions. Several works have solved 
 WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
-These methods use two principles (cross-over and mutation)
-to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
+%These methods use two principles (cross-over and mutation)
+%to change existing solutions and improve them over time.
+These methods do not guarantee termination 
 (i.e., they may run forever without achieving a quality criteria), and even when they do terminate (in practical cases) they do
@@
 thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
-Furthermore, there is no natural way of introducing constraints into genetic methods, which
+Furthermore, there is no natural way of introducing constraints into genetic methods (see e.g., \cite{sorkhabi2018constrained}), which
 leads to a potentially costly feasibility check for every solution found.
@@
 The first attempt to address these limitations 
-was made using a local search approach for WFLO~\cite{ozturk2004heuristic}. 
+was made using a local search approach~\cite{ozturk2004heuristic}. 
 The solution does not guarantee optimality, 
@@
 is that it cannot perform an action 
-that leads to an infeasible state and, therefore, quickly converges to a sub-optimal solution (local optimum) 
+that leads to an infeasible state and, therefore, often quickly converges to a sub-optimal solution (local optimum) 
 that can potentially be much lower quality than the globally optimal solution~\cite{rivas2009solving}.
@@
 This creates a well-connected solution space that leads the algorithm to escape 
-local optima. The SA method was shown to be superior to Ozturk \& Norman~\cite{ozturk2004heuristic} 
+local optima. The SA method was shown to be superior to the approach of Ozturk \& Norman~\cite{ozturk2004heuristic} 
 and to the genetic method proposed in Grady et al.\ \cite{grady2005placement}. A drawback of the SA approach 
@@
 
-In our work, we also use an SA approach (similarly to Rivas et al.\ \cite{rivas2009solving}). However,
+In our work, we also use an annealing based approach (similar to Rivas et al.\ \cite{rivas2009solving}). However,
 our methodology is generic since it is based on a declarative problem model and

### 2020-04-29 paper.tex
Edit to end of Sect 3.2

@@
 in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
-These results lead to the conclusion that both types of methods (exact and approximate) should be considered and, therefore, we introduce a unified paradigm, namely quadratic programming, that enables
+These results lead to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
 both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
@@
 with and without interference from other turbines due to wake effects, respectively. 
-Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.
+Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.\todo{JCB: This is an unnecessary change.
+Go back and reformulate Section~\ref{sec:related} to refer to the potential position $i$ rather than turbine $i$ so that we do not have this inconsistency.}
 Further, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
@@
 Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
-\todo{Are these linear formulations ``equivalent'' as you claimed or approximations?
+\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation.}
+%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations?
 If they are mathematically equivalent, why do we think a quadratic model will be any better computationally?

### 2020-04-29 paper.tex
Finished edit through Sect 3.5.1

@@
 The set of upstream turbines, $\mathcal{U}_{id}$ is computed
-by rotating the grid w.r.t.\ to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$.
+by rotating the grid wrt to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$.
 The expression in Eq.
@@
 with and without interference from other turbines due to wake effects, respectively. 
-Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.\todo{JCB: This is an unnecessary change.
-Go back and reformulate Section~\ref{sec:related} to refer to the potential position $i$ rather than turbine $i$ so that we do not have this inconsistency.}
+Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.
 Further, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
@@
 Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
-\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation.}
+\todo{JCB: We need to formally present the ILP model in the experiment section.
+We can then discuss the fact that the model is identical - because the linearization gives and exact representation.
+Should put a forward reference to that model here.}
 %\todo{Are these linear formulations ``equivalent'' as you claimed or approximations?
@@
 
-Quadratic unconstrained binary optimization problems, as their name implies, do not allow the direct representation of hard constraints. Instead, 
+Quadratic unconstrained binary optimization problems (QUBOs), as their name implies, do not allow the direct representation of hard constraints. Instead, 
 %they only 
@@
 with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
-penalizing solutions that violate the exactly $m$ turbines constrain and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
+penalizing solutions that violate the exactly $m$ turbines constraint and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
 solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
@@
 
-\todo{The proof is a bit awkward as it goes back-and-forth between QCOP and WFLO.
+%\todo{The proof is a bit awkward as it goes back-and-forth between QCOP and WFLO.
 Technically, the problem as defined in \eqref{QCOP}-\eqref{Vars} but, given the different (not mathematically equivalent) models of WFLO, this is different from claiming that WFLO is NPC.
@@
 \subsubsection{QCOP solving methods}
-WFL-QCOP is a type of strictly convex integer 
-quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}. In fact,
-we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard. The common exact methods for solving
-IQCOPs are tree search \cite{land2010automatic} type algorithms with a variety of techniques to prune sub-trees that definitely do not contain the optimal solution.
 
-In operations research, the generic exact approach for solving IQCOPs are the branch-and-cut based mixed integer programming (MIP) solvers \cite{bonami2008algorithmic}.
-By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.
+WFL-QCOP is a strictly convex integer 
+quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}.
+%In fact, we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard.
+The common exact methods for solving
+IQCOPs are tree search \cite{land2010automatic} algorithms with a variety of sophisticated techniques to limit the search necessary to find and prove optimality.
+In operations research, the generic exact approach for solving IQCOPs is the branch-and-cut based mixed integer programming (MIP) \cite{bonami2008algorithmic}.
+A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function.
+This relaxation is used to generate cuts, implied constraints that tighten the relaxation, to identify sub-trees that can be discarded without search because they are guaranteed not to contain a solution better than the current incumbent, and to drive the heuristic branching that systematically searches for a solution.
+Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.
 
-Constraint programming (CP) \cite{rossi2006handbook} is another tree search framework that can be applied to solve IQCOPs.
+Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements delivering orders of magnitude improved performance \cite{Bixby07a}.
+More recently the development of such solvers has focused on improved performance for quadratic problems such as WFLO-QCOP \cite{Furini19a}.
+%By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.
+
+%Constraint programming (CP) \cite{rossi2006handbook} is another tree search framework that can be applied to solve IQCOPs.
 At each node in the search tree, inference algorithms that are often encapsulated in the global constraints \cite{katriel20066}, conduct domain filtering on each constraint.
@@
 
-In the community of communications, IQCOPs are commonly solved via discrete ellipsoid-based search (DEBS), which is built on enumeration of integer points in the hyper-ellipsoid defining the feasible region \cite{agrell2002closest}.
+%In the community of communications, IQCOPs are commonly solved via discrete ellipsoid-based search (DEBS), which is built on enumeration of integer points in the hyper-ellipsoid defining the feasible region \cite{agrell2002closest}.
 DEBS can be interpreted as a form of CP, where the ellipsoid geometry induces for each variable an interval domain, producing an inference algorithm removing values that do not belong to the ellipsoid.
@@
 
-\todo{@Chris and Jason: I still feel that we need to say how Gurobi do it (or at least how we think Gurobi does that)}
+%\todo{@Chris and Jason: I still feel that we need to say how Gurobi do it (or at least how we think Gurobi does that)}
 
@@
 	\includegraphics[width=\textwidth]{energy_50.png}  
-	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (after 50 seconds).}%
+	\caption{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 50 seconds).}%
 
@@
 	\includegraphics[width=\textwidth]{energy_3600.png}   \\%\\
-	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (after 3600 seconds).}%
+	\caption{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 3600 seconds).}%
 

### 2020-04-30 paper.tex
ILP in Section 4.

@@
 \subsubsection{QCOP solving methods}
-
+\label{quadratic}
 WFL-QCOP is a strictly convex integer 
@@
 2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
-and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). 
+and 3 variations for number of turbines ($m \in \{20, 30, 40\}$). 
 For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}.
@@
 We compared our two models, namely WFL-QCOP and WFL-QUBO, against two state of the art exact optimization methods:
-the integer linear programming (ILP-LS) approach from~\cite{donovan2005wind}, and the 
-model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. The CP approach presented in~\cite{Zhang2014} was
+the integer linear programming (ILP-LS) approach from~\cite{donovan2005wind,Zhang2014}, and the 
+model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. 
+The ILP model is given in the following set of equations, which we denote WFL-ILP, \begin{eqnarray} \label{ILP}
+&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)y_{i,j})   \\
+&s.t.:& \nonumber\\
+&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality2}\\
+&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid2}\\
+&\mbox{       }& x_i + x_j -1 \leq y_{i,j},  \forall i,j \in \mathcal{N}, \label{Linearization}\\
+&\mbox{       }& y_{i,j} \geq 0,     \forall i,j \in \mathcal{N}, \label{Yij}\\ 
+&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars2} 
+\end{eqnarray} The variables $y_{i,j}$ are continuous and can take only non negative values (Eq.(\ref{Yij})). However, due to Eq.~(\ref{Yij}) their values must
+be either zero or one. If a turbine is placed in both $i$ and $j$ ($x_i = x_j = 1$), then $y_{i,j}$ will be $1$ and a wake reduction 
+will take place in the objective function. If a turbine is placed in either $x_i$ or $x_j$ (but not in both), $y_{i,j}$ will
+take the value of $0$ and no wake reduction will occur. The WFL-ILP is equivalent to WFL-QCOP: the interaction
+variables $y_{i,j}$ of the ILP model are replaced by multiplicative expressions $x_{i}x_{j}$, which makes Constraint (\ref{Yij}) redundant. 
+However, the ILP is solved using linear optimization approaches, while WFL-QCOP is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
+we can expect a difference in the performance of optimization software when solving the two problems.
+
+The CP approach presented in~\cite{Zhang2014} was
 reported to perform worse than the rest and was left out of the experiments.
@@
 %\end{table}
-
-© 2020 GitHub, Inc.
-Terms
-Privacy
-Security
-Status
-Help
-Contact GitHub
-Pricing
-API
-Training
-Blog
-About

### 2020-04-30 paper.tex
Edits to end of Sect 4.1

--- 2020-04-29 paper.tex
@@
 Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
-\todo{JCB: We need to formally present the ILP model in the experiment section.
+%\todo{JCB: We need to formally present the ILP model in the experiment section.
 We can then discuss the fact that the model is identical - because the linearization gives and exact representation.
@@
 To test and benchmark our models, we used 12 standard benchmark WFLO instances
-from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement}. 
+from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement} that we denote with the set $\mathcal{W}$. 
 The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
@@
 For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}.
-We denote $\mathcal{W}$ the set of 12 instances.
-Note that instances with $36$ wind regimes are considered computationally harder~\cite{Zhang2014}.
+%We denote $\mathcal{W}$ the set of 12 instances.
+Instances with $36$ wind regimes are considered computationally more challenging~\cite{Zhang2014}.
 
@@
 
-We compared our two models, namely WFL-QCOP and WFL-QUBO, against two state of the art exact optimization methods:
-the integer linear programming (ILP-LS) approach from~\cite{donovan2005wind}, and the 
-model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. The CP approach presented in~\cite{Zhang2014} was
-reported to perform worse than the rest and was left out of the experiments.
-We used state of the art software, namely Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-SS and QP-SS. 
+We compared our two models, WFL-QCOP and WFL-QUBO, against two state-of-the-art exact optimization methods:
+the integer linear programming (ILP-LS) approach~\cite{donovan2005wind,Zhang2014} and the 
+model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. 
+The ILP model is a standard, exact linearization of the WFL-QCOP model.
+%% given in the following set of equations, which we denote WFL-ILP, \begin{eqnarray} \label{ILP}
+%% &\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)y_{i,j})   \\
+%% &s.t.:& \nonumber\\
+%% &\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality2}\\
+%% &\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid2}\\
+%% &\mbox{       }& x_i + x_j -1 \leq y_{i,j},  \forall i,j \in \mathcal{N}, \label{Linearization}\\
+%% &\mbox{       }& y_{i,j} \geq 0,     \forall i,j \in \mathcal{N}, \label{Yij}\\ 
+%% &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars2} 
+%% \end{eqnarray} The variables $y_{i,j}$ are continuous and can take only non negative values (Eq.(\ref{Yij})). However, due to Eq.~(\ref{Yij}) their values must
+%% be either zero or one. If a turbine is placed in both $i$ and $j$ ($x_i = x_j = 1$), then $y_{i,j}$ will be $1$ and a wake reduction 
+%% will take place in the objective function. If a turbine is placed in either $x_i$ or $x_j$ (but not in both), $y_{i,j}$ will
+%% take the value of $0$ and no wake reduction will occur. The WFL-ILP is equivalent to WFL-QCOP: the interaction
+%% variables $y_{i,j}$ of the ILP model are replaced by multiplicative expressions $x_{i}x_{j}$, which makes Constraint (\ref{Yij}) redundant. 
+However, the ILP is solved using linear optimization approaches, while WFL-QCOP is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
+we can expect a difference in the performance of optimization software when solving the two problems.
+
+The existing constraint programming approach was
+reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
+We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-SS and QP-SS. 
 For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
 In addition, we 
-applied both Fujitsu's digital annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
+used both Fujitsu's Digital Annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
 as QUBO-LS.
@@
 we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
-until that point in time. The best solution in terms of energy is the solution that attained the highest objective
-values (SS in QP-SS and LS in the rest of the methods).
-For the digital annealer, we used a single configuration that was found to work best across the different instances.
+found until that point in time.
+%The best solution in terms of energy is the solution that attained the highest objective
+%values (SS in QP-SS and LS in the rest of the methods).
+For the DA, we used a single configuration that was found to work best across the different instances.
 The initial temperature is 500000, the temperature mode is 0, the  interval for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically.
 \todo{Eldan and Arik: Do I need to write the formula for computing temperature decay rate?
--- Jason} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime
+-- Jason} \todo{JCB: Yes, if it is not too complicated.
+You should mention if this is the standard thing that DA does or something you came up with.} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime
 of the procedure was recorded.
 \todo{@Chris: do we want to share the code that ran the experiments in batches?
-perhaps it should be separated from the problem construction and the two need to be shared separately}
+perhaps it should be separated from the problem construction and the two need to be shared separately.} \todo{JCB: As I requested last fall, we should have separate problem instance files from the code and make at least the problem instances available.}
 
 The exact methods were all solved using Gurobi's quadratic programming solver on a Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
-Furthermore, QUBO-LS was solved on DA version 1 \todo{Eldan: any additional info that we need to add here?}.
+Furthermore, QUBO-LS was solved on DA version 1.
+\todo{Eldan: any additional info that we need to add here?}
 
@@
 
-We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in kW.}.
-The only method that optimizes on SS is QP-SS. Hence, for the other methods, we converted solutions
-LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in~\cite{Zhang2014}.
+We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{The total mean energy is presented in kW.}.
+The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
+LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}.
 Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
@@
 
-Next, we tested the performance of our approach as function of solution time. To this end, we 
+Next, we tested the performance of our approach as function of solution time. We 
 compared the average performance of all WFLO methods on all 12 WFLO instances.
-Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specs) we used the \emph{mean relative error} (MRE) measure.
+Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specifications) we used the \emph{mean relative error} (MRE) measure.
 The MRE is defined with respect to the best solution obtained per instance and can be computed over time.
@@
 Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
-Moreover, we let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
+Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
 across the different methods.
-For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
+%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
 Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
 Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$.
-Further, denote by $b_{w,t,a}$ denote the best solution attained before time $t$ of approach $a$ for instance $w$.
+Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$.
 Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.
@@
 The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
-We shall present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
+We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
 

### 2020-04-30 paper.tex
Edits for Section 4.2

@@
 
-We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{The total mean energy is presented in kW.}.
+We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss})).
+%\footnote{The total mean energy is presented in kW.}.
 The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
@@
 
-In this part, we present the main findings of our empirical evaluation. 
-We shall start by showing the best attained solutions by all methods.
+%% In this part, we present the main findings of our empirical evaluation. 
+%% We shall start by showing the best attained solutions by all methods.
 Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
@@
 \begin{table}[t!]
-
+	\small
 	\begin{tabular}{| c | c | c | c | c | c | c | c |}
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method.}\label{tab:results1}
+	\caption{Total Expected Power (kW) after 100 seconds for the exact techniques and by 10 seconds for DA-LS.}\label{tab:results1}
 \end{table}
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method.}\label{tab:results2}
+	\caption{Total Expected Power (kW) after 3600 seconds and by 10 seconds for DA-LS.}\label{tab:results2}
 \end{table}
@@
 the three quadratic LS-based approaches (DA-LS, QUBO-LS, and QP-LS) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
-the linear approach (ILP-LS) finds better solutions \todo{@Chris, Jason: do we have a good explanation for this?}.
+the linear approach (ILP-LS) finds better solutions.
+%\todo{@Chris, Jason: do we have a good explanation for this?}. NO
 Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds.
-We see that the DA is still competitive with the other approaches (recall, its solution was found after $10$ seconds),
-while QP-LS finds the best solutions in 10 out of the 12 scenarios. For the 6 hard scenarios (12 wind directions), our quadratic solutions
-dominate all 6. \todo{@Chris: I am not sure that the first table contributes anything beyond the second table}
+For the exact technique, we can see that the improvement between 100 and 3600 seconds of run-time is marginal.
+We also observe that the DA is, therefore, still competitive with the other approaches (recall, its solution was found after $10$ seconds),
+while QP-LS finds the best solutions in 10 out of the 12 scenarios.
+Our quadratic approaches dominate all six hard scenarios (12 wind directions).
+%\todo{@Chris: I am not sure that the first table contributes anything beyond the second table.
+JCB: ``We include the second table to demonstrate that only minor improvements are achieved by the exact techniques between 100 seconds of run-time.''}
 
-We now turn to present the mean relative error over time per approach, i.e., we are presenting $MRE(a,t)$). Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
-presents the MRE after $50$ seconds of running the solutions focusing runtime performance, while Figure~\ref{fig:results2} corresponds
-to the full run (after $3600$ seconds). The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
-correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.
+We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
+presents the MRE after $50$ seconds, while Figure~\ref{fig:results2} corresponds
+to the full run (after $3600$ seconds).
+%The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
+%correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.
 
 The results show that on average (across the 12 instances),
-the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS (after around 1250 seconds).
+the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS after around 1250 seconds.
 
-The two methods that are based on advanced software (QP-LS) and hardware (DA-LS) solutions
-outperform the rest both in terms of runtime (best results in short time), and in terms of overall performance.
+Our proposed quadratic models, QP-LS and DA-LS,
+outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
 Specifically, for most instances, QP-LS gives the best solution at 3600 seconds, while the DA-based approach
@@
 Furthermore, QP-LS converges to the best solution faster than all other exact methods.
-In fact, it is inferior only to the approximate DA-LS method, which is geared mainly toward runtime.
-This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
+In fact, it is inferior only to the approximate DA-LS method.
+%This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
 

### 2020-04-30 paper.tex
Finished edit of Section 5

@@
 introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
-First, the two problems are the basis for the QP-LS and QUBO-LS methods that couple the two
-quadratic formulations of WFLO with Gurobi optimizer. This turns out
+First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi. This turns out
 to be a powerful method as it gives the best performance after 3600 seconds for most instances (7 out of 12),
 and especially,
-for the hard instances with 36 wind directions. In addition, the runtime performance of QP-LS is second
-only to the digital annealer.
-Second, the QUBO model can be mounted onto a quantum-inspired optimization hardware (e.g., Fujitsu's digital annealer) demonstrated by the DA-LS approach in our evaluation.
-DA-LS yields the best results in 7 out of 12
-instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime.
+for the hard instances with 36 wind directions.
+\todo{Avoid ``this'' as a pronoun.
+The previous sentence is a good example.
+I do not know what ``this'' refers to, especially as two models are referred to.
+Also how do you get 7/12?} In addition, the runtime performance of QP-LS is second
+only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
+instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime by two orders of magnitude.\todo{JCB: Is ``two orders of magnitude'' accurate?
+If you compare to 100, it is only 1 order of magnitude to achieve the ``same'' performance.}
 
@@
 multiple wind farm layout problems.
-Moreover, it allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that came
-with (re-)solving the problem using exact methods. In addition, the performance loss due to the use of an approximate method 
-that exploits advanced technology is negligible. This balance (instead of trade-off) between
+Moreover, it allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that come
+with (re-)solving the problem using exact methods. The quality loss due to the use of an approximate method 
+that exploits advanced technology is negligible. This balance between
 runtime and total energy, points at the potential contribution of the proposed quadratic framework for solving WFLO.
@@
 wind farm layout optimization. 
-The framework is based on two models: a constrained model (WFL-QCOP) and an unconstrained binary
-model (WFL-QUBO). The former can be used to solve WFLO using advanced optimization software,
+The framework is embodied in two models: constrained (WFL-QCOP) and unconstrained binary
+models (WFL-QUBO). The former can be used to solve WFLO using advanced optimization software,
 while the latter can be mounted onto quantum-inspired machines that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
-that the computational complexity of WFLO is $\mathcal{NP}$-hard, a pessimistic result that 
-was also noticed experimentally in~\cite{Zhang2014}. However,
+that the computational complexity of WFLO is $\mathcal{NP}$-hard. However,
 our empirical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
 coupled with our WFL-QCOP and WFL-QUBO models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
-We also show that when an advanced hardware is used with the QUBO representation, large instances WFLO can be solved quickly with no significant degradation in the total energy compared to exact methods.
+We also show that when an advanced hardware is used with the QUBO representation, large WFLO instances can be solved quickly with no significant degradation in the total energy compared to exact methods.
 This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our quadratic framework useful for designing new wind farms, while varying multiple input parameters.
@@
 In future work, we aim at extending our quadratic models to include additional 
-WFLO aspects, such as noise, non-flat terrain and multiple objectives. 
+WFLO aspects, such as noise, non-flat terrain, and multiple objectives. 
 Moreover, we wish to explore real-world wind farm design problems 

### 2020-04-30 paper.tex
evaluation.

@@
 \todo{@Chris: do we want to share the code that ran the experiments in batches?
-perhaps it should be separated from the problem construction and the two need to be shared separately.} \todo{JCB: As I requested last fall, we should have separate problem instance files from the code and make at least the problem instances available.}
+perhaps it should be separated from the problem construction and the two need to be shared separately.} \todo{JCB: As I requested last fall, we should have separate problem instance files from the code and make at least the problem instances available.
+Arik: OK, we will upload the experiment to the github and share the link.}
 
@@
 introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
-First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi. This turns out
-to be a powerful method as it gives the best performance after 3600 seconds for most instances (7 out of 12),
+First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi. These methods turn out
+to be a powerful approach to solving WFLO as it gives the best performance after 3600 seconds for most instances (7 out of 12), \todo{@Chris: I counted the number of instances for which the two methods show best results after 3600.
+is that ok?}
 and especially,
-for the hard instances with 36 wind directions.
-\todo{Avoid ``this'' as a pronoun.
-The previous sentence is a good example.
-I do not know what ``this'' refers to, especially as two models are referred to.
-Also how do you get 7/12?} In addition, the runtime performance of QP-LS is second
+for the hard instances with 36 wind directions. In addition, the runtime performance of QP-LS is second
 only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
 instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime by two orders of magnitude.\todo{JCB: Is ``two orders of magnitude'' accurate?
-If you compare to 100, it is only 1 order of magnitude to achieve the ``same'' performance.}
+If you compare to 100, it is only 1 order of magnitude to achieve the ``same'' performance.
+Arik: I refer to 10 seconds for the DA, hence 2 orders.}
 

### 2020-05-01 paper.tex
Eldan's part.

@@
 
-\todo{Eldan's part}
+Recent years have seen significant advancement in the use of specialized hardware platforms, such as adiabatic and gate-based quantum computers, Digital/CMOS annealers, and neuromorphic computers \cite{coffrin2019evaluating}, to solve hard combinatorial optimization problems.
+Interestingly, a large number of these platforms support the optimization ofquadratic functions over binary variables by using Ising models \cite{johnson2011quantum}, an equivalent representation to QUBO \cite{bian2010ising}, as their mathematical abstraction.
+
+In this work, we use the first generation of the Fujitsu Digital Annealer (DA) \cite{aramon2019physics}, a specialized CMOS hardware capable of solving QUBO problems with up to 1024 binary variables.
+The algorithm used by the DA is based on simulated annealing \cite{kirkpatrick1983optimization}, however it takes advantage of the massive parallelization provided by the custom CMOS hardware and uses a dynamic offset mechanism to escape local minima \cite{aramon2019physics}.
+The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising}, communication \cite{naghsh2019digitally,rahman2019ising}, etc.
+
+	Unlike %exact techninques
+	standard exact optimization solvers that run until obtaining an optimal solution, the DA runs until reaching a time limit and returns the best found solution.
+{\color{purple} In our experiments, we analyze the quality of obtained solutions for different time limits.} \todo{EC: can be phrased instead as : In our experiments, we run the DA with increasingly large time limits until we observe no further improvement in objective.
+(see my point in Section 1.4)}
 
@@
 of the procedure was recorded.
+\todo{\\EC: I am not sure what does it mean for the DA to run until convergence.
+You may want to specify that the DA was run with increasingly large time limits until no further improvement in objective.\\}
+
 \todo{@Chris: do we want to share the code that ran the experiments in batches?

### 2020-05-04 paper.tex
Add formula for temperature decay rate.

@@
 For the DA, we used a single configuration that was found to work best across the different instances.
-The initial temperature is 500000, the temperature mode is 0, the  interval for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically.
-\todo{Eldan and Arik: Do I need to write the formula for computing temperature decay rate?
--- Jason} \todo{JCB: Yes, if it is not too complicated.
+The initial temperature \text{tmp\_st} is 500000, the temperature mode is 0, the interval \text{tmp\_interval} for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically, i.e., $ \text{tmp\_decay} = (0.001 / \text{tmp\_st})^{(1.0 / (\text{num\_iterations} / \text{tmp\_interval}))} $.
+This formula of decay rate is for better convergence and is provided by DA developers.
+\todo{JCB: Yes, if it is not too complicated.
 You should mention if this is the standard thing that DA does or something you came up with.} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime

### 2020-05-07 paper.tex
minor.

--- 2020-05-01 paper.tex
@@
 standard exact optimization solvers that run until obtaining an optimal solution, the DA runs until reaching a time limit and returns the best found solution.
-{\color{purple} In our experiments, we analyze the quality of obtained solutions for different time limits.} \todo{EC: can be phrased instead as : In our experiments, we run the DA with increasingly large time limits until we observe no further improvement in objective.
-(see my point in Section 1.4)}
+In our experiments, we analyze the quality of obtained solutions for different time limits.
 
@@
 	\item WFL-QUBO when solved using the digital annealer (DA) achieves a similar performance to state-of-the-art exact methods.
-		\item The runtime for solving WFL-QUBO on the DA is two orders of magnitude lower than WFL-QCOP.  
-	\item Given sufficient runtime, solving the WFL-QCOP model yields superior results on the larger instances.
+		\item The runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
+	\item Given sufficient runtime, solving the WFL-QCOP model yields superior results on most of the instances.
 \end{itemize}
@@
 the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
-ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.
+ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}.
 
@@
 we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
-found until that point in time.
+found until that point in time. The exact methods were all solved using Gurobi's quadratic programming solver on a 
+Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
+
+For the DA, we used a single configuration that was found to work best across the different instances. 
+The DA was run for increasingly large time limits ranging from $0.7$ seconds to $65$ seconds where we no longer observe an improvement in 
+performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~(\ref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
+is set to the highest value that the Digital Annealer allows\footnote{The DA has an upper limit on the size of objective function coefficients.}.
+
 %The best solution in terms of energy is the solution that attained the highest objective
 %values (SS in QP-SS and LS in the rest of the methods).
-For the DA, we used a single configuration that was found to work best across the different instances.
-The initial temperature is 500000, the temperature mode is 0, the  interval for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically.
-\todo{Eldan and Arik: Do I need to write the formula for computing temperature decay rate?
--- Jason} \todo{JCB: Yes, if it is not too complicated.
+
+%
+%The initial temperature \text{tmp\_st} is 500000, the temperature mode is 0, the interval \text{tmp\_interval} for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically, i.e., $ \text{tmp\_decay} = (0.001 / \text{tmp\_st})^{(1.0 / (\text{num\_iterations} / \text{tmp\_interval}))} $.
+This formula of decay rate is for better convergence and is provided by DA developers.
+\todo{JCB: Yes, if it is not too complicated.
 You should mention if this is the standard thing that DA does or something you came up with.} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime
-of the procedure was recorded.
+%of the procedure was recorded.
 \todo{\\EC: I am not sure what does it mean for the DA to run until convergence.
 You may want to specify that the DA was run with increasingly large time limits until no further improvement in objective.\\}
-
-\todo{@Chris: do we want to share the code that ran the experiments in batches?
+%
+%
+%\todo{@Chris: do we want to share the code that ran the experiments in batches?
 perhaps it should be separated from the problem construction and the two need to be shared separately.} \todo{JCB: As I requested last fall, we should have separate problem instance files from the code and make at least the problem instances available.
 Arik: OK, we will upload the experiment to the github and share the link.}
-
-The exact methods were all solved using Gurobi's quadratic programming solver on a Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
-Furthermore, QUBO-LS was solved on DA version 1.
-\todo{Eldan: any additional info that we need to add here?}
 
@@
 yields competitive solutions after less than 10 seconds. 
-Furthermore, QP-LS converges to the best solution faster than all other exact methods.
-In fact, it is inferior only to the approximate DA-LS method.
+Furthermore, 
+QP-LS converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate DA-LS method.
+
+To demonstrate the speedup gained from using the DA, we compare the point in time when the DA reaches 
+a competitive solution (10 seconds) to the point in time when QP-LS reaches the same solution in terms of energy.
+Table~\ref{speedup} summarizes the performance of QP-LS when compared to that of the DA after 10 seconds.
+We see a speedup factor of X. \todo{discuss with Chris}
 %This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
@@
 introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
-First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi. These methods turn out
-to be a powerful approach to solving WFLO as it gives the best performance after 3600 seconds for most instances (7 out of 12), \todo{@Chris: I counted the number of instances for which the two methods show best results after 3600.
-is that ok?}
+First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi.
+The two methods, namely QP-LS and QUBO-LS, turn out to be powerful approaches to solving WFLO.
+Specifically, QP-LS yields the best performance (after 3600 seconds) for most problem instances
 and especially,
@@
 only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
-instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime by two orders of magnitude.\todo{JCB: Is ``two orders of magnitude'' accurate?
-If you compare to 100, it is only 1 order of magnitude to achieve the ``same'' performance.
-Arik: I refer to 10 seconds for the DA, hence 2 orders.}
+instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (\todo{speed-up ratio of ...})
 

### 2020-05-14 paper.tex
speedups

@@
 the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
-ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}.
+ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}
 
@@
 performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~(\ref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
-is set to the highest value that the Digital Annealer allows\footnote{The DA has an upper limit on the size of objective function coefficients.}.
+is set to the highest value that the Digital Annealer allows.\footnote{The DA has an upper limit on the size of objective function coefficients.}
 
@@
 Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.
-
 We can now average over the different instances at time $t$.
@@
 We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
+
+\todo{@Chris - new part:} The last measure that we introduce is the speedup factor $S(a_1, a_2, t, w)$. The measure compares the time that it takes
+approach $a_1 \in \mathcal{A}$ to
+achieve the same energy level as the energy achieved by approach $a_2 \in \mathcal{A}$
+within $t$ seconds when solving instance $w$ (energy levels are measured in terms of the sum-of-squares energy expression). 
+Formally, $$S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, $$
+with $E_{SS}(a, t, w)$ being the SS energy value under approach $a$ for instance $w$ by time $t$. The speedup factor 
+is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
+For example, $$S(QP-LS, DA-LS, 10sec, w_1)$$ is computed as the ratio between the time that it takes QP-LS to 
+reach the same quality of solution as DA-LS achieves after $10$ seconds (over problem $w_1$). Since
+we limit the runtime of the exact methods by $3600$ seconds, we may get a lower bound on the speedup factor. 
+Therefore, if for the example above the result is $3600/10$, we shall interpret it as a speedup factor of \emph{at least} $360$. \todo{@Chris - new part ends here.}
 
@@
 
-To demonstrate the speedup gained from using the DA, we compare the point in time when the DA reaches 
-a competitive solution (10 seconds) to the point in time when QP-LS reaches the same solution in terms of energy.
-Table~\ref{speedup} summarizes the performance of QP-LS when compared to that of the DA after 10 seconds.
-We see a speedup factor of X. \todo{discuss with Chris}
+\todo{@Chris - new part:} To demonstrate the gain in runtime from using the DA, we compare the average speedup factor between
+the QP-LS (the dominant exact method) to a DA-LS solutions after $1$ and $10$ seconds.
+In other words, $w \in \mathcal{W}$ we computed $S(QP-LS, DA-LS, 1sec, w)$ and $S(QP-LS, DA-LS, 10sec, w)$
+for every $w$ 
+and averaged the speedup factors across the instances.
+
+As a result, the average speedup factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively.\footnote{The speedup factors  were rounded to have 0 values after the floating point.} These results show that the DA achieves competitive solutions after running for $1$ second.
+Note that since for many of the instances QP-LS did not attain
+the energy level of DA-LS at time $t$ (1 or 10 seconds), we consider the speedup factor to be a lower bound, meaning
+that the DA-LS gives potentially an even higher speedup factor, which we did not compute due to the time limit of $3600$. \todo{@Chris - new part ends here}
 %This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
@@
 \subsection{Discussion and Practical Implications}
-
+\todo{@Chris - please check discussion:}
 Our result show empirically that the quadratic programming models that we
@@
 only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
-instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (\todo{speed-up ratio of ...})
+instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (speedup factor of 2 orders of magnitude when the DA runs for $1$ second).
 

### 2020-05-14 paper.tex
Edits for Section 4 and 5. Comments left for Arik.

--- 2020-05-07 paper.tex
@@
 Our main results are: \begin{itemize}
-	\item WFL-QUBO when solved using the digital annealer (DA) achieves a similar performance to state-of-the-art exact methods.
-		\item The runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
+	\item WFL-QUBO when solved using the digital annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
+		\item To achieve such a solution quality, the runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
 	\item Given sufficient runtime, solving the WFL-QCOP model yields superior results on most of the instances.
+\todo{Arik: Is this correct?
+Table 2 shows DA@10seconds with the best solution found for 7/12 instances and QP-LS@3600 for 6/12.
+And 4/6 vs.
+3/6 for the WR36).
+Are you comparing QP-LS with QP-SS and ILP-LS?
+We seem to be underplaying DA performance in this comment.}
 \end{itemize}
+
 We start by outlining the experimental design of our evaluation, followed by an overview of our main results,
@@
 and 3 variations for number of turbines ($m \in \{20, 30, 40\}$). 
-For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}.
+For the $36$ wind directions setting, the velocities come from the probability distribution presented in Figure~\ref{fig:prob_wind}.
 %We denote $\mathcal{W}$ the set of 12 instances.
@@
 as QUBO-LS.
+\todo{Arik: Is it necessary to change the names WFL-QUBO -> QUBO-LS, WFL-QCOP -> QP-LS?
+This makes it confusing for someone who skims over this paragraph.Can't we use the same names (one or the other) through-out?}
 
@@
 The DA was run for increasingly large time limits ranging from $0.7$ seconds to $65$ seconds where we no longer observe an improvement in 
-performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~(\ref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
-is set to the highest value that the Digital Annealer allows\footnote{The DA has an upper limit on the size of objective function coefficients.}.
+performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~\eqref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
+is set to the highest value that the Digital Annealer allows.\footnote{The DA has an upper limit on the size of objective function coefficients.}
 
@@
 Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
-Then, the SS based energy expression can be written as, 
+Then, the SS energy expression can be written as, 
 $$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$
@@
 compared the average performance of all WFLO methods on all 12 WFLO instances.
-Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specifications) we used the \emph{mean relative error} (MRE) measure.
-The MRE is defined with respect to the best solution obtained per instance and can be computed over time.
-Then, the
-time-varying MRE can be averaged across the 12 scenarios and presented over time. In what follows, we demonstrate the construction of the MRE measure.
+Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specifications) we used the \emph{mean relative error} (MRE) measure with respect to the best solution obtained per instance.
+%Then, the
+%time-varying MRE can be averaged across the 12 scenarios and presented over time.
+%In what follows, we demonstrate the construction of the MRE measure.
 
@@
 Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
-Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$.
+Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \mbox{\emph{QUBO-LS}}, \mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, \mbox{\emph{ILP-LS}}, \mbox{\emph{QP-SS}}\}$.
 Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$.
-Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.
-
+Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
+\begin{equation}
+  RE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.
+\end{equation} Note that the expression is always positive, since we are solving a maximization problem.
 We can now average over the different instances at time $t$.
-The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
+The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} RE(w,t,a).\end{equation}
 We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
+
+The last measure that we introduce is the speed-up factor $S(a_1, a_2, t, w)$. The measure compares the time that it takes
+approach $a_1 \in \mathcal{A}$ to
+achieve the same energy level as the energy achieved by approach $a_2 \in \mathcal{A}$
+within $t$ seconds when solving instance $w$ (energy levels are measured in terms of the sum-of-squares energy expression). 
+Formally, $$S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, $$
+with $E_{SS}(a, t, w)$ being the SS energy value under approach $a$ for instance $w$ at time $t$. The speed-up factor 
+is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
+For example, $$S(\mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, 10, w_1)$$ is the ratio between the time that it takes QP-LS to 
+reach the same quality of solution as DA-LS achieves after $10$ seconds on problem $w_1$. Since
+we limit the runtime of the exact methods by $3600$ seconds, $S$ is a lower bound on the true speed-up factor. 
+%Therefore, if for the example above the result is $3600/10$, we shall interpret it as a speed-up factor of \emph{at least} $360$.
 
@@
 For the exact technique, we can see that the improvement between 100 and 3600 seconds of run-time is marginal.
-We also observe that the DA is, therefore, still competitive with the other approaches (recall, its solution was found after $10$ seconds),
+We also observe that the DA at 10 seconds is, therefore, still competitive with the other approaches,
 while QP-LS finds the best solutions in 10 out of the 12 scenarios. 
@@
 %correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.
-
-The results show that on average (across the 12 instances),
-the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS after around 1250 seconds.
 
@@
 
-To demonstrate the speedup gained from using the DA, we compare the point in time when the DA reaches 
-a competitive solution (10 seconds) to the point in time when QP-LS reaches the same solution in terms of energy.
-Table~\ref{speedup} summarizes the performance of QP-LS when compared to that of the DA after 10 seconds.
-We see a speedup factor of X. \todo{discuss with Chris}
+The results also show that on average 
+the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS after around 1250 seconds.
+
+To demonstrate the gain in runtime from using the DA, we compare the average speed-up factor between
+the QP-LS (the dominant exact method) to a DA-LS solutions after $1$ and $10$ seconds.
+For each $w \in \mathcal{W}$ we computed the mean $S(\mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, t, w)$ for $t \in \{1,10\}$. 
+%and $S(QP-LS, DA-LS, 10sec, w)$
+%for every $w$ 
+%and averaged the speed-up factors across the instances. 
+The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively.
+\todo{Arik: What about a table -- perhaps in an Appendix?
+If not, you should at least quote the min and max speed-up factors and comment on the speed-up comparing DA with the other techniques.}
+%\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}
+
+These results show that the DA achieves competitive solutions after running for $1$ second.
+Note that since for many of the instances QP-LS did not attain
+the energy level of DA-LS at time $t$ (1 or 10 seconds),
+%we consider the speed-up factor to be a lower bound, meaning
+that
+the DA-LS speed-up is actually even higher.
+%speed-up factor, which we did not compute due to the time limit of $3600$. 
 %This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
@@
 
+\todo{Arik: Should this be Sectiobn 5 rather than 4.3?}
+
 Our result show empirically that the quadratic programming models that we
 introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
-First, the two problems are the basis for the QP-LS and QUBO-LS methods that solved with Gurobi.
-The two methods, namely QP-LS and QUBO-LS, turn out to be powerful approaches to solving WFLO.
+First, the two problems are the basis for the QP-LS and QUBO-LS methods solved with Gurobi.
+The two methods are powerful approaches to solving WFLO.
 Specifically, QP-LS yields the best performance (after 3600 seconds) for most problem instances
@@
 only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
-instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (\todo{speed-up ratio of ...})
+instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).
 
@@
 
+Finally, though we have not emphasized it here, all the approaches presented, including those from the literature, have the advantage of being represented with a declarative mathematical model.
+Such a representation makes problem extensions significantly easier and allows formal analysis.
+While model-based  approaches are common in exact optimization such as mixed-integer programming, they are rare for heuristic techniques such as genetic algorithms and meta-heuristics.
+Many of the state-of-the-art implementations of heuristic techniques are specialized for the particular problem being solved.
+A change in the problem definition (e.g., the addition of a new constraint) can, therefore, require substantial software changes.
+Conversely, model-based techniques can be adapted to a revised problem definition via the declarative addition of the new constraint to the model.
+Further, amenability to formal analysis means that substantial insight can be gained from mathematical treatments of problem classes, leading, as can be seen in mixed integer programming \cite{Bixby07a}, to substantial improvements in problem solving performance.
+
+%State-of-the-art solvers still make use of advanced data structures but they are specialized for the class of problems expressible in the mathematical formalism rather than the problem at hand.
+Additionally, the mathematical representation of the model allows formal results to be proved.
+Model-based approaches can therefore provide
+
 \section{Conclusion}
@@
 The framework is embodied in two models: constrained (WFL-QCOP) and unconstrained binary
-models (WFL-QUBO). The former can be used to solve WFLO using advanced optimization software,
+ (WFL-QUBO). The former can be used to solve WFLO using advanced optimization software,
 while the latter can be mounted onto quantum-inspired machines that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
-that the computational complexity of WFLO is $\mathcal{NP}$-hard. However,
+that the computational complexity of WFLO is $\mathcal{NP}$-hard. Nonetheless,
 our empirical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are

### 2020-05-15 paper.tex
replies and edits.

@@
 		\item To achieve such a solution quality, the runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
-	\item Given sufficient runtime, solving the WFL-QCOP model yields superior results on most of the instances.
+	\item Given sufficient runtime, solving the WFL-QCOP model yields competitive results on most of the instances.
 \todo{Arik: Is this correct?
@@
 Are you comparing QP-LS with QP-SS and ILP-LS?
-We seem to be underplaying DA performance in this comment.}
+We seem to be underplaying DA performance in this comment.
+Chris: I changed to competitive results.
+You're right, I'd say that the DA is not `losing' to QP-LS.}
 \end{itemize}
@@
 reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
-We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-SS and QP-SS. 
+We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-LS and QP-SS. 
 For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
@@
 \todo{Arik: Is it necessary to change the names WFL-QUBO -> QUBO-LS, WFL-QCOP -> QP-LS?
-This makes it confusing for someone who skims over this paragraph.Can't we use the same names (one or the other) through-out?}
+This makes it confusing for someone who skims over this paragraph.Can't we use the same names (one or the other) through-out?
+Chris: The issue is that WFL-QUBO and WFL-QCOP are problem names without mentioning of the objective - I need to differentiate QP-SS from QP-LS.
+Should I change everything backwards to the XX-LS notation?
+I could do that.}
 
@@
 Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
-Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \mbox{\emph{QUBO-LS}}, \mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, \mbox{\emph{ILP-LS}}, \mbox{\emph{QP-SS}}\}$.
+Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \mbox{
+	\{\emph{QUBO-LS}}, \mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, \mbox{\emph{ILP-LS}}, \mbox{\emph{QP-SS}}\}$.
 Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$.
@@
 The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
+Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds, meaning that for some of the instances DA-LS was actually
+slower than QP-LS. The maximum speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
+QP-LS was not able to find a comparable solution to DA-LS within the $3600$ time limit.
 \todo{Arik: What about a table -- perhaps in an Appendix?
-If not, you should at least quote the min and max speed-up factors and comment on the speed-up comparing DA with the other techniques.}
+If not, you should at least quote the min and max speed-up factors and comment on the speed-up comparing DA with the other techniques.
+Chris: I added min/max and comparison to the two other LS-based methods.}
 %\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}
@@
 that
-the DA-LS speed-up is actually even higher.
+the DA-LS speed-up is actually even higher. When compared
+to the other LS-based methods, the DA speeds up by a higher factor compared to its speedup w.r.t QP-LS. At $1$ ($10$) seconds, the speedup factor compared to ILP-LS 
+is $1385$ ($180$) and for the QUBO-LS it is $900$ ($152$). Here, we compare only LS based methods as they optimize on the same objective as the DA.  
 %speed-up factor, which we did not compute due to the time limit of $3600$. 
@@
 
-\todo{Arik: Should this be Sectiobn 5 rather than 4.3?}
+\todo{Arik: Should this be Sectiobn 5 rather than 4.3? Chris: I am fine with both. I felt that it was strongly tied to 4.2, but I can see it becoming a short standalone section.}
 

### 2020-05-15 paper.tex
Addressed todos, edited some of the recently added text.

@@
 Recent years have seen significant advancement in the use of specialized hardware platforms, such as adiabatic and gate-based quantum computers, Digital/CMOS annealers, and neuromorphic computers \cite{coffrin2019evaluating}, to solve hard combinatorial optimization problems.
-Interestingly, a large number of these platforms support the optimization ofquadratic functions over binary variables by using Ising models \cite{johnson2011quantum}, an equivalent representation to QUBO \cite{bian2010ising}, as their mathematical abstraction.
+Interestingly, a large number of these platforms support the optimization of quadratic functions over binary variables by using Ising models \cite{johnson2011quantum}, an equivalent representation to QUBO \cite{bian2010ising}, as their mathematical abstraction.
 
@@
 		\item To achieve such a solution quality, the runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
-	\item Given sufficient runtime, solving the WFL-QCOP model yields competitive results on most of the instances.
-\todo{Arik: Is this correct?
+	\item Given sufficient runtime, solving the WFL-QCOP model yields strong results on most of the instances.
+Compared to previous exact models (QP-SS and ILP-LS), WFL-QCOP is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios.
+%  \todo{Arik: Is this correct?
 Table 2 shows DA@10seconds with the best solution found for 7/12 instances and QP-LS@3600 for 6/12.
@@
 as QUBO-LS.
-\todo{Arik: Is it necessary to change the names WFL-QUBO -> QUBO-LS, WFL-QCOP -> QP-LS?
-This makes it confusing for someone who skims over this paragraph.Can't we use the same names (one or the other) through-out?
-Chris: The issue is that WFL-QUBO and WFL-QCOP are problem names without mentioning of the objective - I need to differentiate QP-SS from QP-LS.
-Should I change everything backwards to the XX-LS notation?
-I could do that.}
+\todo{Arik: I think it is important to use consistent names through-out.
+The problem is that we do not have a unified naming format.
+The problems can be QC-LS, QU-LS (for Quadratic Constraint/Unconstrained Linear Superposition) instead of WFL-QCOP QFL-QUBO.
+Then the results need to include the solver and so we have QU-LS(DA), QU-LS(GRB), QC-LS(GRB), QC-SS(GRB), and LC-LS(GRB).
+(LC = Linear Constrained).
+I suggest you use Latex macros so that we can easily change then all again if necessary.}
 
@@
 To demonstrate the gain in runtime from using the DA, we compare the average speed-up factor between
-the QP-LS (the dominant exact method) to a DA-LS solutions after $1$ and $10$ seconds.
+ QP-LS and a DA-LS solutions after $1$ and $10$ seconds.
 For each $w \in \mathcal{W}$ we computed the mean $S(\mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, t, w)$ for $t \in \{1,10\}$. 
@@
 The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
-Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds, meaning that for some of the instances DA-LS was actually
-slower than QP-LS. The maximum speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
-QP-LS was not able to find a comparable solution to DA-LS within the $3600$ time limit.
-\todo{Arik: What about a table -- perhaps in an Appendix?
-If not, you should at least quote the min and max speed-up factors and comment on the speed-up comparing DA with the other techniques.
-Chris: I added min/max and comparison to the two other LS-based methods.}
+Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances QP-LS achieved the same quality as DA-LS in less time.
+The maximum speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
+QP-LS was not able to find a comparable solution to DA-LS within the $3600$ time limit. As a consequence, 
+the DA-LS speed-up is actually even higher than computed above. \todo{Arik: I think a table of speed-ups per problem instance is still necessary in an Appendix.}
 %\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}
 
-These results show that the DA achieves competitive solutions after running for $1$ second.
-Note that since for many of the instances QP-LS did not attain
-the energy level of DA-LS at time $t$ (1 or 10 seconds),
+%These results show that the DA achieves competitive solutions after running for $1$ second.
+%Many of the instances running QP-LS for 3600 seconds did not result in as good a solution as DA-LS at time $t$ (1 or 10 seconds).
 %we consider the speed-up factor to be a lower bound, meaning
-that
-the DA-LS speed-up is actually even higher. When compared
-to the other LS-based methods, the DA speeds up by a higher factor compared to its speedup w.r.t QP-LS. At $1$ ($10$) seconds, the speedup factor compared to ILP-LS 
-is $1385$ ($180$) and for the QUBO-LS it is $900$ ($152$). Here, we compare only LS based methods as they optimize on the same objective as the DA.  
+When compared
+to the other LS-based methods, the DA speed-up is even larger than when compared to QP-LS.  At $1$ ($10$) seconds, the speedup factor compared to ILP-LS 
+is $1385$ ($180$) and for the QUBO-LS it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
 %speed-up factor, which we did not compute due to the time limit of $3600$. 
@@
 
-\subsection{Discussion and Practical Implications}
-
-\todo{Arik: Should this be Sectiobn 5 rather than 4.3? Chris: I am fine with both. I felt that it was strongly tied to 4.2, but I can see it becoming a short standalone section.}
+\section{Discussion and Practical Implications}
 

### 2020-05-18 paper.tex
speedup results and renaming.

@@
 %\newcommand{\myauthorhyperref}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}
+\newcommand{\qcls}{{\sf {\small QC-LS\xspace}}}
+\newcommand{\ilpls}{{\sf {\small ILP-LS\xspace}}}
+\newcommand{\quls}{{\sf {\small QU-LS\xspace}}}
+\newcommand{\qcss}{{\sf {\small QC-SS\xspace}}}
+
+\newcommand{\qclsgrb}{{\sf {\small QC-LS(GRB)\xspace}}}
+\newcommand{\ilplsgrb}{{\sf {\small ILP-LS(GRB)\xspace}}}
+\newcommand{\qulsgrb}{{\sf {\small QU-LS(GRB)\xspace}}}
+\newcommand{\qcssgrb}{{\sf {\small QC-SS(GRB)\xspace}}}
+\newcommand{\qulsda}{{\sf {\small QU-LS(DA)\xspace}}}
 
@@
 and enforce proximity constraints between the turbines, respectively. We shall refer to the QCOP
-formulation of WFLO as WFL-QCOP.
-
-WFL-QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
+formulation of WFLO as \qcls{}.
+
+\qcls{} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
 Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
@@
 solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
-must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as WFL-QUBO.
+must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as \quls.
 %These types of constraints
@@
 
-\subsection{The Computational Complexity of WFL-QCOP}
+\subsection{The Computational Complexity of \qcls{}}
 \label{sec:computational}
@@
 %We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP).
-We therefore provide a straightforward proof of the computational complexity of WFL-QCOP.
+We therefore provide a straightforward proof of the computational complexity of \qcls{}.
 
 \begin{mythm}
-	The computational complexity of WFL-QCOP defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
+	The computational complexity of \qcls{} defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
 \end{mythm}
 \begin{myproof}
-We prove by showing that a special case of the WFL-QCOP \eqref{QCOP}-\eqref{Vars}  is
+We prove by showing that a special case of the \qcls{} \eqref{QCOP}-\eqref{Vars}  is
 $\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
 are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
-Since a special case of WFL-QCOP is  $\mathcal{NP}$-hard it implies that WFL-QCOP is at least $\mathcal{NP}$-hard. On the other hand,
-WFL-QCOP is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFL-QCOP is $\mathcal{NP}$-hard.
+Since a special case of \qcls{} is  $\mathcal{NP}$-hard it implies that \qcls{} is at least $\mathcal{NP}$-hard. On the other hand,
+\qcls{} is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that \qcls{} is $\mathcal{NP}$-hard.
 \end{myproof}
@@
 \label{quadratic}
-WFL-QCOP is a strictly convex integer 
+\qcls{} is a strictly convex integer 
 quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}.
@@
 Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements delivering orders of magnitude improved performance \cite{Bixby07a}.
-More recently the development of such solvers has focused on improved performance for quadratic problems such as WFLO-QCOP \cite{Furini19a}.
+More recently the development of such solvers has focused on improved performance for quadratic problems such as \qcls{} \cite{Furini19a}.
 %By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.
@@
 Our main results are: \begin{itemize}
-	\item WFL-QUBO when solved using the digital annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
-		\item To achieve such a solution quality, the runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
-	\item Given sufficient runtime, solving the WFL-QCOP model yields strong results on most of the instances.
-Compared to previous exact models (QP-SS and ILP-LS), WFL-QCOP is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios.
+	\item \quls{} when solved using the digital annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
+		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  
+	\item Given sufficient runtime, solving the \qcls{} model yields strong results on most of the instances.
+Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios.
 %  \todo{Arik: Is this correct?
-Table 2 shows DA@10seconds with the best solution found for 7/12 instances and QP-LS@3600 for 6/12.
+Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12.
 And 4/6 vs.
 3/6 for the WR36).
-Are you comparing QP-LS with QP-SS and ILP-LS?
+Are you comparing \qcls{}grb with \qcssgrb and ILP-LS?
 We seem to be underplaying DA performance in this comment.
 Chris: I changed to competitive results.
-You're right, I'd say that the DA is not `losing' to QP-LS.}
+You're right, I'd say that the DA is not `losing' to \qcls{}grb.}
 \end{itemize}
@@
 
-We compared our two models, WFL-QCOP and WFL-QUBO, against two state-of-the-art exact optimization methods:
-the integer linear programming (ILP-LS) approach~\cite{donovan2005wind,Zhang2014} and the 
-model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. 
-The ILP model is a standard, exact linearization of the WFL-QCOP model.
+We compared our two models, \qcls{} and \quls{}, against two state-of-the-art exact optimization methods:
+the integer linear programming (denoted \ilpls) approach~\cite{donovan2005wind,Zhang2014} and the 
+model that uses a quadratic approximation (\qcss)~\cite{turner2014new}. 
+The ILP model is a standard, exact linearization of the \qcls{} model.
 %% given in the following set of equations, which we denote WFL-ILP, \begin{eqnarray} \label{ILP}
@@
 %% variables $y_{i,j}$ of the ILP model are replaced by multiplicative expressions $x_{i}x_{j}$, which makes Constraint (\ref{Yij}) redundant. 
-However, the ILP is solved using linear optimization approaches, while WFL-QCOP is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
+However, the ILP is solved using linear optimization approaches, while \qcls{} is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
 we can expect a difference in the performance of optimization software when solving the two problems.
@@
 reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
-We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-LS and QP-SS. 
-For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
+We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve \qcls{}, \ilpls{} and \qcss. 
+We shall denote the problems using their names and the solvers used as follows: \qclsgrb{} (GRB stands for Gurobi), \ilplsgrb{}, and \qcssgrb{}.
 In addition, we 
-used both Fujitsu's Digital Annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
-as QUBO-LS.
-\todo{Arik: I think it is important to use consistent names through-out.
-The problem is that we do not have a unified naming format.
-The problems can be QC-LS, QU-LS (for Quadratic Constraint/Unconstrained Linear Superposition) instead of WFL-QCOP QFL-QUBO.
-Then the results need to include the solver and so we have QU-LS(DA), QU-LS(GRB), QC-LS(GRB), QC-SS(GRB), and LC-LS(GRB).
-(LC = Linear Constrained).
-I suggest you use Latex macros so that we can easily change then all again if necessary.}
+used both Fujitsu's Digital Annealer and Gurobi to solve \quls: we shall refer to the former as \qulsda{}, and to the latter
+as \qulsgrb.
 
@@
 
-For exact methods (QP-SS, ILP-LS, QP-LS and QUBO-LS),
+For exact methods (\qcss{}, \ilpls{}, \qcls{} and \quls{}),
 we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
@@
 %The best solution in terms of energy is the solution that attained the highest objective
-%values (SS in QP-SS and LS in the rest of the methods).
+%values (SS in \qcssgrb and LS in the rest of the methods).
 
@@
 %\footnote{The total mean energy is presented in kW.}.
-The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
+The only method that directly optimizes SS is \qcssgrb{}. Hence, for the other methods, we converted solutions
 LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}.
@@
 across the different methods.
-%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
+%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by \qulsgrb and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
 Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$.
-Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \mbox{
-	\{\emph{QUBO-LS}}, \mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, \mbox{\emph{ILP-LS}}, \mbox{\emph{QP-SS}}\}$.
-Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$.
+Let $\mathcal{A}$ be the set of solution approaches, $$\mathcal{A}= \mbox{
+	\{\emph{\qulsgrb}}, \mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, \mbox{\emph{\ilplsgrb}}, \mbox{\emph{\qcssgrb}}\}.$$ Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$.
 Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
@@
 within $t$ seconds when solving instance $w$ (energy levels are measured in terms of the sum-of-squares energy expression). 
-Formally, $$S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, $$
+Formally, \begin{equation} S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, \label{eq:speedup} \end{equation}
 with $E_{SS}(a, t, w)$ being the SS energy value under approach $a$ for instance $w$ at time $t$. The speed-up factor 
 is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
-For example, $$S(\mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, 10, w_1)$$ is the ratio between the time that it takes QP-LS to 
-reach the same quality of solution as DA-LS achieves after $10$ seconds on problem $w_1$. Since
+For example, $$S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, 10, w_1)$$ is the ratio between the time that it takes \qclsgrb{} to 
+reach the same quality of solution as \qulsda{} achieves after $10$ seconds on problem $w_1$. Since
 we limit the runtime of the exact methods by $3600$ seconds, $S$ is a lower bound on the true speed-up factor. 
@@
 for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
-The runtime for DA-LS was always less than $10$ seconds.
+The runtime for the DA was always less than $10$ seconds.
 
@@
 		\toprule
-		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
+		Wind Directions  & n  & m  & \qulsda{}  & \qulsgrb  & \qcls & \qcss  & \ilpls  \\
 		\toprule
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 100 seconds for the exact techniques and by 10 seconds for DA-LS.}\label{tab:results1}
+	\caption{Total Expected Power (kW) after 100 seconds for the exact techniques and by 10 seconds for \qulsda{}.}\label{tab:results1}
 \end{table}
@@
 		\toprule
-		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
+		Wind Directions  & n  & m  & \qulsda  & \qulsgrb  & \qcls & \qcss  & \ilpls  \\
 		\toprule
@@
 	\vspace{0.5em}
-	\caption{Total Expected Power (kW) after 3600 seconds and by 10 seconds for DA-LS.}\label{tab:results2}
+	\caption{Total Expected Power (kW) after 3600 seconds and by 10 seconds for \qulsda{}.}\label{tab:results2}
 \end{table}
@@
 comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
-the three quadratic LS-based approaches (DA-LS, QUBO-LS, and QP-LS) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
-the linear approach (ILP-LS) finds better solutions.
+the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
+the linear approach (\ilplsgrb) finds better solutions.
 %\todo{@Chris, Jason: do we have a good explanation for this?}. NO
@@
 We also observe that the DA at 10 seconds is, therefore, still competitive with the other approaches,
-while QP-LS finds the best solutions in 10 out of the 12 scenarios. 
+while \qclsgrb{} finds the best solutions in 10 out of the 12 scenarios. 
 Our quadratic approaches dominate all six hard scenarios (12 wind directions).
@@
 
-Our proposed quadratic models, QP-LS and DA-LS,
+Our proposed quadratic models, \qcls{} and \quls{},
 outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
-Specifically, for most instances, QP-LS gives the best solution at 3600 seconds, while the DA-based approach
+Specifically, for most instances, \qclsgrb{} gives the best solution at 3600 seconds, while the DA-based approach
 yields competitive solutions after less than 10 seconds. 
 Furthermore, 
-QP-LS converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate DA-LS method.
+\qclsgrb{} converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method.
 
 The results also show that on average 
-the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS after around 1250 seconds.
-
-To demonstrate the gain in runtime from using the DA, we compare the average speed-up factor between
- QP-LS and a DA-LS solutions after $1$ and $10$ seconds.
-For each $w \in \mathcal{W}$ we computed the mean $S(\mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, t, w)$ for $t \in \{1,10\}$. 
-%and $S(QP-LS, DA-LS, 10sec, w)$
+the \ilplsgrb{} method starts with poor solutions, yet ends up improving over \qcssgrb{} and \qulsgrb{} after around 1250 seconds.
+
+\begin{sloppypar}To demonstrate the gain in runtime from using the DA, we compare the average speed-up factor between
+ \qclsgrb{} and a \qulsda{} solutions after $1$ and $10$ seconds.
+For each $w \in \mathcal{W}$ we computed the mean $S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, t, w)$ for $t \in \{1,10\}$. 
+%and $S(\qclsgrb, \qulsda, 10sec, w)$
 %for every $w$ 
@@
 The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
-Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances QP-LS achieved the same quality as DA-LS in less time.
+Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
 The maximum speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
-QP-LS was not able to find a comparable solution to DA-LS within the $3600$ time limit. As a consequence, 
-the DA-LS speed-up is actually even higher than computed above. \todo{Arik: I think a table of speed-ups per problem instance is still necessary in an Appendix.}
+\qclsgrb{} was not able to find a comparable solution to \qulsda{} within the $3600$ time limit. As a consequence, 
+the \qulsda{} speed-up is actually even higher than computed above. \end{sloppypar}
 %\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}
@@
 %These results show that the DA achieves competitive solutions after running for $1$ second.
-%Many of the instances running QP-LS for 3600 seconds did not result in as good a solution as DA-LS at time $t$ (1 or 10 seconds).
+%Many of the instances running \qclsgrb for 3600 seconds did not result in as good a solution as \qulsda at time $t$ (1 or 10 seconds).
 %we consider the speed-up factor to be a lower bound, meaning
 When compared
-to the other LS-based methods, the DA speed-up is even larger than when compared to QP-LS.  At $1$ ($10$) seconds, the speedup factor compared to ILP-LS 
-is $1385$ ($180$) and for the QUBO-LS it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
+to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the speedup factor compared to \ilplsgrb{} 
+is $1385$ ($180$) and for the \qulsgrb{} it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
+For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in the Section~\ref{app1}.
 %speed-up factor, which we did not compute due to the time limit of $3600$. 
@@
 Our result show empirically that the quadratic programming models that we
-introduced in Section~\ref{sec:QUBO4WFLO} (WFL-QCOP and WFL-QUBO) are useful in two ways.
-First, the two problems are the basis for the QP-LS and QUBO-LS methods solved with Gurobi.
-The two methods are powerful approaches to solving WFLO.
-Specifically, QP-LS yields the best performance (after 3600 seconds) for most problem instances
+introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
+First, the two methods are powerful approaches to solving WFLO. Specifically, \qcls{} yields the best performance (after 3600 seconds) for most problem instances
 and especially,
-for the hard instances with 36 wind directions. In addition, the runtime performance of QP-LS is second
-only to the Digital Annealer.  Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the DA-LS approach. DA-LS yields the best results in 7 out of 12 
+for the hard instances with 36 wind directions. In addition, the runtime performance of \qcls{} is second
+only to the Digital Annealer.
+Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the \qulsda{} approach.
+\qulsda{} yields the best results in 7 out of 12
 instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).
 
-The ability of DA-LS to quickly obtain competitive solutions on large WFLO instances 
+The ability of \qulsda{} to quickly obtain competitive solutions on large WFLO instances 
 is crucial when solving
@@
 wind farm layout optimization. 
-The framework is embodied in two models: constrained (WFL-QCOP) and unconstrained binary
- (WFL-QUBO). The former can be used to solve WFLO using advanced optimization software,
+The framework is embodied in two models: constrained (\qcls) and unconstrained binary
+ (\quls). The former can be used to solve WFLO using advanced optimization software,
 while the latter can be mounted onto quantum-inspired machines that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
@@
 our empirical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
-coupled with our WFL-QCOP and WFL-QUBO models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
+coupled with our \qcls{} and \quls{} models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
 We also show that when an advanced hardware is used with the QUBO representation, large WFLO instances can be solved quickly with no significant degradation in the total energy compared to exact methods.
@@
 \bibliography{Biblio}
+
+\appendix
+
+\section{Speedup Results}
+\label{app1}
+
+\begin{table}[t!]
+	\small
+	\begin{tabular}{| c | c | c | c | c | c | }
+		\toprule
+		Wind Directions  & n  & m  & \qcls &  \ilpls & \qulsgrb  \\
+		\toprule
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}  & 20 & 1       & 1 & 1  \\
+		& & 30   & N/A & N/A & 1     \\
+		& & 40 & 1 & 1 & N/A                \\
+		\cline{2-6}
+		&\multirow{3}{*}{20}   & 20  & 1  & 1  & 259         \\
+		&&30   & 3 & 2  & 9    \\
+		&&40   & 32 & 22  & 602        \\
+		\hline
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20 & 1       & 1661 & 1  \\
+		&& 30  & 3600 & 3600 & 3600                     \\
+		&&40   & 1 & 3600 & 1          \\
+		\cline{2-6}
+		&  \multirow{3}{*}{20}   & 20   & 3600  & 538 & 944            \\
+		&&30   & 22 & 3600 & 3304                     \\
+		&&40   & 60 & 3600 & 1717 \\
+		\bottomrule                   
+	\end{tabular}
+
+	\vspace{0.5em}
+	\caption{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 1 second. The value 3600 implies a lower bound on the speedup factor.}\label{tab:results3}
+\end{table}
+
+\begin{table}[t!]
+	\small
+	\begin{tabular}{| c | c | c | c | c | c | }
+		\toprule
+		Wind Directions  & n  & m  & \qcls &  \ilpls & \qulsgrb  \\
+		\toprule
+		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}  & 20 & 0.1       & 0.1 & 0.1  \\
+		& & 30   & N/A & N/A & 0.1     \\
+		& & 40 & 0.1 & 0.1 & N/A                \\
+		\cline{2-6}
+		&\multirow{3}{*}{20}   & 20  & 1  & 1  & 25.9         \\
+		&&30   & 0.3 & 0.2  & 0.9    \\
+		&&40   & 6.1 & 2.2  & 360        \\
+		\hline
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20 & 0.1       & 360 & 0.2  \\
+		&& 30  & 360 & 360 & 360                     \\
+		&&40   & 0.1 & 360 & 0.1          \\
+		\cline{2-6}
+		&  \multirow{3}{*}{20}   & 20   & 360  & 360 & 360            \\
+		&&30   &  360  & 360 & 360                    \\
+		&&40   &  8  & 360 & 360 \\
+
+		\bottomrule                   
+	\end{tabular}
+
+	\vspace{0.5em}
+	\caption{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 10 second. The value 360 implies a lower bound on the speedup factor.}\label{tab:results4}
+\end{table}
+Table~\ref{tab:results3} presents the speedup results across all problem instances
+for every LS-based method (computed using the SS energy term). The speedup is calculated according to Eq.~(\ref{eq:speedup}). Table~\ref{tab:results4} show similar results for the speedup factor
+compared to DA at 10 seconds. Note that N/A values
+indicate that the solver has gotten an LS solution within 3600 seconds that turned out to be worse
+than a DA solution after 1 (or 10 seconds). The average lower bound on the speedup at $1$ seconds
+is 665,	1511, and 949 for \qclsgrb, \qulsgrb, and \ilplsgrb, respectively. For the $10$ seconds
+results, the average
+speedup factor is at least 99, 196, and	166 for the respective methods.
+For instances where speedup factor in Table~\ref{tab:results4} is 10 times less than in Table~\ref{tab:results3},
+the DA has found the 10 seconds solution after 1 second.
 

### 2020-05-21 paper.tex
Figures and Fujitsu.

@@
 Our main results are: \begin{itemize}
-	\item \quls{} when solved using the digital annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
+	\item \quls{} when solved using the Digital Annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
 		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  

### 2020-05-27 paper.tex
Minor revision

@@
 we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
-found until that point in time. The exact methods were all solved using Gurobi's quadratic programming solver on a 
-Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
+found until that point in time. The exact methods were solved using Gurobi's integer programming and quadratic programming solvers on a 
+Ubuntu PC with Intel(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
 

### 2020-05-28 paper.tex
Eldan's comments.

@@
 similar quality to software-based approaches but with two orders of
-magnitude lower run-time.
+magnitude lower \todo{Eldan: shorter instead of lower?} run-time.
 %The ability to quickly obtain competitive
@@
 objectives (including turbine installation and maintenance
-costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we focus on the
-novel quadratic approach for WFLO, and hence consider the most
+costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we 
+choose to focus on a
+novel approach for solving WFLO, and hence consider the most
 abstract wind farm model. Our model can be extended, in the
-future, to include additional problem characteristics.
+future, to include additional problem characteristics as mentioned above.
 
@@
 the wind's free speed and direction.  The wake effect can be pre-computed
-for each pair of cells since we know: the distance ($x$), the turbine
-specifications and terrain conditions (that jointly dictate $r$), and
+for each pair of cells since we know the distance ($x$), the turbine
+specifications and terrain conditions (that jointly dictate $r$), and the
 probabilistic behavior of wind regimes~\cite{Zhang2014}.
@@
 regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
-Rather, $u_{id}$, can be computed using the following equation~\cite{Zhang2014}:
+Rather, $u_{id}$ can be computed using the following equation~\cite{Zhang2014}:
 \begin{equation}
@@
 compared to existing approximate and exact techniques.
-Another non-linear optimization approach \cite{ulku2019new}, relaxed the binary placement variables to continuous values in $[0,1]$, 
+Another non-linear optimization approach \cite{ulku2019new} relaxed the binary placement variables to continuous values in $[0,1]$
 and approximated the constraints such that the resulting variables yield an integer placement.
@@
 A benchmark study comparing state-of-the-art exact approaches \cite{Zhang2014} to 
-approximate methods and found that the two perform in a comparable fashion, without large differences between
+approximate methods found that the two perform in a comparable fashion, without large differences between
 the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
@@
 The former represents WFLO characteristics (proximity and number of turbines)
-as hard constraints, while the latter considers these characteristics as soft constraints that are penalized as part of the objective function.
+as hard constraints, while the latter considers these characteristics as soft constraints such that the objective function is penalized in case these
+constraints are violated.
 
@@
 constraints we write
-the following quadratic optimization problem (QCOP):
+the following quadratic constrained optimization problem (QCOP):
 \begin{eqnarray} \label{QCOP}
@@
 \qcls{} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
-Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
+Linear versions of the QCOP (\qcls) have been previously proposed and solved using the corresponding exact solvers~\cite{Zhang2014,donovan2005wind}.
 %\todo{JCB: We need to formally present the ILP model in the experiment section.
@@
 Let $\lambda = (\lambda_1,\lambda_2)$ 
-be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
+be a vector of constraint weights with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
 \begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
@@
 penalizing solutions that violate the exactly $m$ turbines constraint and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
-solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
+solutions that violate the proximity constraints. The two penalty terms $\lambda_1$ and $\lambda_2$ 
 must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as \quls.
@@
 We prove by showing that a special case of the \qcls{} \eqref{QCOP}-\eqref{Vars}  is
-$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
+$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse-grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
 are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
@@
 Our main results are: \begin{itemize}
-	\item \quls{} when solved using the Digital Annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
+	\item \quls{} when solved using the Digital Annealer (DA) achieves a solution quality as good or better (within time limit) than existing state-of-the-art exact methods on 10 of 12 instances.
 		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  
@@
   RE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.
-\end{equation} Note that the expression is always positive, since we are solving a maximization problem.
+\end{equation} Note that the expression is always non-negative, since we are solving a maximization problem.
 We can now average over the different instances at time $t$.
@@
 comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
-the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
+the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 5 out of 6 instances. For easier instances, 
 the linear approach (\ilplsgrb) finds better solutions.
 %\todo{@Chris, Jason: do we have a good explanation for this?}. NO
-Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds.
+Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds for exact methods.
 For the exact technique, we can see that the improvement between 100 and 3600 seconds of run-time is marginal.
@@
 while \qclsgrb{} finds the best solutions in 10 out of the 12 scenarios. 
-Our quadratic approaches dominate all six hard scenarios (12 wind directions).
+Our quadratic approaches dominate all six hard scenarios (36 wind directions).
 %\todo{@Chris: I am not sure that the first table contributes anything beyond the second table.
@@
 We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
-presents the MRE after $50$ seconds, while Figure~\ref{fig:results2} corresponds
+presents the MRE over the first $50$ seconds, while Figure~\ref{fig:results2} corresponds
 to the full run (after $3600$ seconds).
@@
 
-Our proposed quadratic models, \qcls{} and \quls{},
+\todo{Eldan regarding the first sentence in what follows: I feel like this is not well defined?
+MRE is a single metric.
+Do you mean that our approahces have the best MRE both after a short time period (of say 10 seconds) but also at longer time limits (including 3600)?} Our proposed quadratic models, \qcls{} and \quls{},
 outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
@@
 The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
-Per instance, the minimum speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
-The maximum speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
+Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
+The \emph{maximum} speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
 \qclsgrb{} was not able to find a comparable solution to \qulsda{} within the $3600$ time limit. As a consequence, 
@@
 When compared
-to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the speedup factor compared to \ilplsgrb{} 
+to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the mean speedup factor compared to \ilplsgrb{} 
 is $1385$ ($180$) and for the \qulsgrb{} it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
-For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in the Section~\ref{app1}.
+For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in~\ref{app1}.
 %speed-up factor, which we did not compute due to the time limit of $3600$. 
@@
 
-Our result show empirically that the quadratic programming models that we
+Our results show empirically that the quadratic programming models that we
 introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
@@
 \qulsda{} yields the best results in 7 out of 12
-instances, shows competitive performance on the other instances, and outperforms all methods in terms of runtime (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).
+instances, shows competitive performance on the other instances, and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).
 
@@
 and solve them 
-using our quadratic programming approach.
+using our quadratic programming approaches.
 

### 2020-05-28 paper.tex
Edited to end of Sect 2.

@@
 that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
-An alternative formulation of WFLO relies on a less accurate wake model, which uses linear superpositioning
+An alternative formulation of WFLO relies on a less accurate wake model, using linear superpositioning
 to capture multiple wakes~\cite{donovan2005wind}.
@@
 (QUBO) representation of the WFLO problem.  Such a formulation enables the use of
-nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics} here.
+nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics}.
 %, a hardware-based
@@
 The main
-contributions of this paper are:
+contributions of this paper are as follows:
 \begin{enumerate} 
@@
 similar quality to software-based approaches but with two orders of
-magnitude lower \todo{Eldan: shorter instead of lower?} run-time.
+magnitude shorter run-time.
 %The ability to quickly obtain competitive
@@
 novel approach for solving WFLO, and hence consider the most
-abstract wind farm model. Our model can be extended, in the
-future, to include additional problem characteristics as mentioned above.
+abstract wind farm model. Extension of our framework to include such additional problem characteristics is left as future research.
+%Our model can be extended, in the
+%future, to include additional problem characteristics as mentioned above.
 
@@
 Turbines must be placed at least \emph{five rotor diameters}
-apart. Depending on the cell size, we shall formulate the
+apart. Depending on the cell size, we formulate the
 corresponding constraint as part of the optimization model. For
@@
 around the turbine represents the forbidden cells.
+Conversely, if $c = 200$ and turbines are placed in the center of the cell, no proximity constraints are required as the required separation is enforced by the granularity of the spatial discretization.
 
@@
 
-Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
-regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
+Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.
+%\footnote{Turbines can adjust their orientation according to the current wind regime.}
+  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
 Rather, $u_{id}$ can be computed using the following equation~\cite{Zhang2014}:
@@
 a neighborhood search method that allows visits to infeasible solutions. 
-This creates a well-connected solution space that leads the algorithm to escape 
+SA creates a well-connected solution space that allows the algorithm to escape 
 local optima. The SA method was shown to be superior to the approach of Ozturk \& Norman~\cite{ozturk2004heuristic} 

### 2020-05-28 paper.tex
Finished edits

@@
 $\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse-grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
-are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
-Since a special case of \qcls{} is  $\mathcal{NP}$-hard it implies that \qcls{} is at least $\mathcal{NP}$-hard. On the other hand,
+are dropped and we get the heaviest k-subgraph problem, which is a known $\mathcal{NP}$-hard problem~\cite{billionnet2005different}.
+Since a special case of \qcls{} is  $\mathcal{NP}$-hard,  \qcls{} is at least $\mathcal{NP}$-hard. On the other hand,
 \qcls{} is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that \qcls{} is $\mathcal{NP}$-hard.
@@
 
-Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements delivering orders of magnitude improved performance \cite{Bixby07a}.
-More recently the development of such solvers has focused on improved performance for quadratic problems such as \qcls{} \cite{Furini19a}.
+Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements have delivered orders of magnitude improved performance \cite{Bixby07a}.
+More recently the development of such solvers has focused on representing and solving quadratic problems such as \qcls{}~\cite{Furini19a}.
 %By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.
@@
 The algorithm used by the DA is based on simulated annealing \cite{kirkpatrick1983optimization}, however it takes advantage of the massive parallelization provided by the custom CMOS hardware and uses a dynamic offset mechanism to escape local minima \cite{aramon2019physics}.
-The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising}, communication \cite{naghsh2019digitally,rahman2019ising}, etc.
+The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising} and communication \cite{naghsh2019digitally,rahman2019ising}.
 
@@
 Our main results are: \begin{itemize}
-	\item \quls{} when solved using the Digital Annealer (DA) achieves a solution quality as good or better (within time limit) than existing state-of-the-art exact methods on 10 of 12 instances.
-		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  
-	\item Given sufficient runtime, solving the \qcls{} model yields strong results on most of the instances.
-Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios.
+	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than existing state-of-the-art exact methods on 10 of 12 instances.
+		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
+	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances.
+Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard ones.
 %  \todo{Arik: Is this correct?
@@
 We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve \qcls{}, \ilpls{} and \qcss. 
-We shall denote the problems using their names and the solvers used as follows: \qclsgrb{} (GRB stands for Gurobi), \ilplsgrb{}, and \qcssgrb{}.
+We shall denote the problems using their names and the solvers used as follows: \qclsgrb{} (GRB for Gurobi), \ilplsgrb{}, and \qcssgrb{}.
 In addition, we 
-used both Fujitsu's Digital Annealer and Gurobi to solve \quls: we shall refer to the former as \qulsda{}, and to the latter
+used both Fujitsu's Digital Annealer and Gurobi to solve \quls: we shall refer to the former as \qulsda{} and to the latter
 as \qulsgrb.
@@
 found until that point in time. The exact methods were solved using Gurobi's integer programming and quadratic programming solvers on a 
-Ubuntu PC with Intel(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.
+Ubuntu PC with Intel(R) Core(TM) i7-9700 8 core CPU@3.00 GHz with 12 MB cache and 8 GB RAM.
 
@@
 performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~\eqref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
-is set to the highest value that the Digital Annealer allows.\footnote{The DA has an upper limit on the size of objective function coefficients.}
+is the highest value that the Digital Annealer allows.\footnote{The DA has an upper limit on the size of objective function coefficients.}
 
@@
 %\footnote{The total mean energy is presented in kW.}.
-The only method that directly optimizes SS is \qcssgrb{}. Hence, for the other methods, we converted solutions
-LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}.
+The only method that directly optimizes SS is \qcssgrb{}. Hence, for the other methods, we converted
+LS-based solutions into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}.
 Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
@@
 Let $\mathcal{A}$ be the set of solution approaches, $$\mathcal{A}= \mbox{
-	\{\emph{\qulsgrb}}, \mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, \mbox{\emph{\ilplsgrb}}, \mbox{\emph{\qcssgrb}}\}.$$ Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$.
+	\{\emph{\qulsgrb}}, \mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, \mbox{\emph{\ilplsgrb}}, \mbox{\emph{\qcssgrb}}\}.$$ Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ by approach $a$ for instance $w$.
 Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
@@
 We observe that the DA-based solution provides
-comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
+comparable results to state-of-the-art exact methods. For the harder instances (with $36$ wind directions),
 the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 5 out of 6 instances. For easier instances, 
@@
 
-\todo{Eldan regarding the first sentence in what follows: I feel like this is not well defined?
+%\todo{Eldan regarding the first sentence in what follows: I feel like this is not well defined?
 MRE is a single metric.
-Do you mean that our approahces have the best MRE both after a short time period (of say 10 seconds) but also at longer time limits (including 3600)?} Our proposed quadratic models, \qcls{} and \quls{},
-outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
+Do you mean that our approaches have the best MRE both after a short time period (of say 10 seconds) but also at longer time limits (including 3600)?}
+Our proposed quadratic models, \qcls{} and \quls{},
+outperform the rest of the techniques across the entire runtime time horizon that was tested.
 Specifically, for most instances, \qclsgrb{} gives the best solution at 3600 seconds, while the DA-based approach
@@
 \qclsgrb{} was not able to find a comparable solution to \qulsda{} within the $3600$ time limit. As a consequence, 
-the \qulsda{} speed-up is actually even higher than computed above. \end{sloppypar}
+the mean \qulsda{} speed-up is actually even higher than computed above. \end{sloppypar}
 %\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}
@@
 The framework is embodied in two models: constrained (\qcls) and unconstrained binary
- (\quls). The former can be used to solve WFLO using advanced optimization software,
-while the latter can be mounted onto quantum-inspired machines that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
+ (\quls) quadratic programs. The former can be used to solve WFLO using advanced optimization software,
+while the latter can be mounted onto specialized hardware that can quickly (approximately) solve WFLO. We used the constrained quadratic formulation to prove 
 that the computational complexity of WFLO is $\mathcal{NP}$-hard. Nonetheless,
@@
 Table~\ref{tab:results3} presents the speedup results across all problem instances
-for every LS-based method (computed using the SS energy term). The speedup is calculated according to Eq.~(\ref{eq:speedup}). Table~\ref{tab:results4} show similar results for the speedup factor
+for every LS-based method (computed using the SS energy term). The speedup is calculated according to Eq.~(\ref{eq:speedup}). Table~\ref{tab:results4} shows similar results for the speedup factor
 compared to DA at 10 seconds. Note that N/A values
-indicate that the solver has gotten an LS solution within 3600 seconds that turned out to be worse
-than a DA solution after 1 (or 10 seconds). The average lower bound on the speedup at $1$ seconds
-is 665,	1511, and 949 for \qclsgrb, \qulsgrb, and \ilplsgrb, respectively. For the $10$ seconds
+indicate that the solver has achieved an LS solution within 3600 seconds that turned out to be worse
+than a DA solution after 1 (or 10 seconds). \todo{JCB: What is the difference between ``N/A'' and 3600/360?} The average lower bound on the speedup at $1$ seconds
+is 665,	1511, and 949 for \qclsgrb, \qulsgrb, and \ilplsgrb, respectively. For the $10$ second
 results, the average
 speedup factor is at least 99, 196, and	166 for the respective methods.
-For instances where speedup factor in Table~\ref{tab:results4} is 10 times less than in Table~\ref{tab:results3},
-the DA has found the 10 seconds solution after 1 second.
+For instances where the speedup factor in Table~\ref{tab:results4} is 10 times less than in Table~\ref{tab:results3},
+the DA found the 10-second solution after 1 second.
 

### 2020-05-28 paper.tex
Appendix.

@@
 for every LS-based method (computed using the SS energy term). The speedup is calculated according to Eq.~(\ref{eq:speedup}). Table~\ref{tab:results4} shows similar results for the speedup factor
-compared to DA at 10 seconds. Note that N/A values
-indicate that the solver has achieved an LS solution within 3600 seconds that turned out to be worse
-than a DA solution after 1 (or 10 seconds). \todo{JCB: What is the difference between ``N/A'' and 3600/360?} The average lower bound on the speedup at $1$ seconds
+compared to DA at 10 seconds. The average lower bound on the speedup at $1$ seconds
 is 665,	1511, and 949 for \qclsgrb, \qulsgrb, and \ilplsgrb, respectively. For the $10$ second
@@
 the DA found the 10-second solution after 1 second.
+
+Note that the ``N/A'' values in the table correspond to a particular situation arising from our use of two different energy functions, $E_SS$ and $E_LS$.
+For some problem instances an exact solver for \qcls{} finds and proves an optimal solution for the $E_LS$ objective.
+However, when that solution is evaluated using the $E_SS$ energy function, it has a lower energy than the solution found by the DA.
+This phenomenon is correct because the mapping from $E_LS$ to $E_SS$ does not necessarily preserve the ordering of solution quality.
+We do not include these instances in our speedup computations.
 

### 2020-05-28 paper.tex
appendix again.

@@
 
-Note that the ``N/A'' values in the table correspond to a particular situation arising from our use of two different energy functions, $E_SS$ and $E_LS$.
-For some problem instances an exact solver for \qcls{} finds and proves an optimal solution for the $E_LS$ objective.
+Note that the ``N/A'' values in the table correspond to a particular situation arising from our use of two different energy functions, $E_{SS}$ and $E_{LS}$.
+For some problem instances an exact solver for \qcls{} finds and proves an optimal solution for the $E_{LS}$ objective.
 However, when that solution is evaluated using the $E_SS$ energy function, it has a lower energy than the solution found by the DA.
-This phenomenon is correct because the mapping from $E_LS$ to $E_SS$ does not necessarily preserve the ordering of solution quality.
+This phenomenon is correct because the mapping from $E_{LS}$ to $E_{SS}$ does not necessarily preserve the ordering of solution quality.
 We do not include these instances in our speedup computations.

### 2020-05-28 paper.tex
appendix yet again.
Minor edits on Appendix

@@
 
-Note that the ``N/A'' values in the table correspond to a particular situation arising from our use of two different energy functions, $E_{SS}$ and $E_{LS}$.
-For some problem instances an exact solver for \qcls{} finds and proves an optimal solution for the $E_{LS}$ objective.
-However, when that solution is evaluated using the $E_SS$ energy function, it has a lower energy than the solution found by the DA.
+Note that the ``N/A'' values in the table arise from our use of two different energy functions, $E_{SS}$ and $E_{LS}$.
+For some problem instances, an exact solver for \qcls{} finds and proves an optimal solution for the $E_{LS}$ objective but, when that solution is evaluated using the $E_{SS}$ energy function, it has a lower energy than the solution found by the DA.
 This phenomenon is correct because the mapping from $E_{LS}$ to $E_{SS}$ does not necessarily preserve the ordering of solution quality.

### 2020-07-17 paper.tex
efficiency issues added.

@@
 %While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
-to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization models.
-Recent advances in both optimization hardware and software have targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
+to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches.
+However, state-of-the-art solutions often trade-off between efficiency (i.e., providing quick solutions to WFLO) and accuracy (i.e., providing an optimal energy solution).
+In this work, we aim to find a balancing approach that would enable to obtain WFLO solutions quickly and, solve the problem to optimality if accuracy is preferred.
+To this end, we exploit recent advances in both optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
 %It is a key problem 
@@
 that
-improve the state of the art.
+improve the state of the art and strike a balance between efficiency and optimality.
 In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
@@
 
+In recent years, there has been a growing need to efficiently solve WFLO~\cite{wilson2013learning,wilson2014continuous},
+while maintaining competitive (albeit not always optimal) solutions. 
+The focus on quickly solving the problem was mainly motivated by the fact
+that modern farms cover large areas and include hundreds, and
+sometimes even thousands, of turbines. The layout design
+process is iterative, computationally expensive, and controlled
+by various stakeholders. For each iteration, designers must either
+alter an incremental layout or propose a new design
+which they have generated by incorporating new constraints and parameter values.
+Hence,
+one wishes
+to solve WFLO interactively at design time, while changing 
+many of the input parameter values
+and performing sensitivity analysis on those values~\cite{wilson2018evolutionary}.
+
 In this work, we introduce two novel quadratic WFLO models. The first
@@
 nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics}.
+By employing the two types of advances (software and hardware) for solving
+quadratic problems, we enjoy the best of both worlds: we are able to provide fast as well as high performing solutions (in terms of the total energy).
+Therefore, our approach achieves the desired balance between efficiency and effectiveness of the resulting design,
 %, a hardware-based
@@
 The ability of \qulsda{} to quickly obtain competitive solutions on large WFLO instances 
-is crucial when solving
-multiple wind farm layout problems.
+is crucial when iteratively solving
+multiple wind farm layout problems at design time.
 Moreover, it allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that come
@@
 using our quadratic programming approaches.
-
-\bibliographystyle{elsarticle-num-names}
-%\bibliographystyle{plainnat}
-\bibliography{Biblio}
 
@@
 
+\bibliographystyle{elsarticle-num-names}
+%\bibliographystyle{plainnat}
+\bibliography{Biblio}
+
 \end{document}

### 2020-07-17 paper.tex
literature review and minor modifications

@@
 to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches.
-However, state-of-the-art solutions often trade-off between efficiency (i.e., providing quick solutions to WFLO) and accuracy (i.e., providing an optimal energy solution).
-In this work, we aim to find a balancing approach that would enable to obtain WFLO solutions quickly and, solve the problem to optimality if accuracy is preferred.
+However, state-of-the-art solutions often trade-off between efficiency (providing quick designs) and accuracy (guaranteeing optimal energy solutions).
+In this work, we aim to find a balancing approach that would enable to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when accuracy is the goal.
 To this end, we exploit recent advances in both optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
@@
 %solution (Gurobi), 
-In this paper, we 
+We 
 introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
@@
 
-In recent years, there has been a growing need to efficiently solve WFLO~\cite{wilson2013learning,wilson2014continuous},
+In recent years, there has been a growing need to efficiently solve WFLO~
+\cite{wilson2014continuous},
 while maintaining competitive (albeit not always optimal) solutions. 
@@
 that modern farms cover large areas and include hundreds, and
-sometimes even thousands, of turbines. The layout design
+sometimes even thousands, of turbines~\cite{wilson2013learning}. The layout design
 process is iterative, computationally expensive, and controlled
@@
 many of the input parameter values
-and performing sensitivity analysis on those values~\cite{wilson2018evolutionary}.
-
-In this work, we introduce two novel quadratic WFLO models. The first
+and performing a sensitivity analysis to test the impact of those changes~\cite{wilson2018evolutionary}.
+The efficient approaches proposed in~\cite{wilson2013learning,wilson2014continuous} do not
+take into account the physical model of the wind farm, and hence
+do not aim at optimizing the actual energy objective.
+
+In this work, we introduce two novel quadratic WFLO models with the goal 
+of combining efficiency, physical farm representation, and optimality. The first
 model represents WFLO as a quadratic constrained optimization
@@
 By employing the two types of advances (software and hardware) for solving
-quadratic problems, we enjoy the best of both worlds: we are able to provide fast as well as high performing solutions (in terms of the total energy).
-Therefore, our approach achieves the desired balance between efficiency and effectiveness of the resulting design,
+quadratic problems, we enjoy the best of both worlds: 
+we are able to provide fast and high performing solutions (in terms of the total energy). Moreover, our approach is based
+on a physical model of the wind farm, taking into account
+exact calculations of the total energy and the wake effects. 
+Therefore, our approach achieves the desired balance between efficiency and effectiveness of the resulting design, in a model-based fashion.
 %, a hardware-based
@@
 The main
-contributions of this paper are as follows:
+contributions of this paper are as follows: \todo{Chris, perhaps we can strengthen the contribution list}
 \begin{enumerate} 
@@
 
+Recently, several efficient WFLO solutions based on developmental models were introduced in~\cite{wilson2013learning,wilson2014continuous}.
+These works do not represent the actual underlying physical model of the system, and while providing fast solutions to WFLO,
+they do not attempt to optimize on the actual objective function (that is, the total energy that a farm produced) nor capture the complexities of the underlying problem.
+For example, the approach in~\cite{wilson2013learning} models turbine interference using a parametric probabilistic model (assuming a Weibull distribution).
+In our quadratic models, we explicitly represent the energy term, and the physical dependencies between the turbines (i.e., wakes and proximity constraints).
+Furthermore, our models are
+optimizing on an empirically established energy objective, namely the linear superposition expression in Eq.~(\ref{eq:ls}).
+
 %\paragraph{Quadratic Programming as a Unified Framework for Exact and Approximate WFLO}

### 2020-07-17 paper.tex
Editted Abstract.

@@
 %role as a source of renewable energy and 
-  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and a model of the wind patterns, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference among turbines.
+  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines.
 Given the pairwise wake interactions, it is natural
@@
 However, state-of-the-art solutions often trade-off between efficiency (providing quick designs) and accuracy (guaranteeing optimal energy solutions).
-In this work, we aim to find a balancing approach that would enable to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when accuracy is the goal.
+In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when accuracy is the goal.
 To this end, we exploit recent advances in both optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
@@
 %Our results
-show that both approaches yield fast, high-quality solutions 
+show that our quadratic framework achieves fast, high-quality solutions 
 that
 improve the state of the art and strike a balance between efficiency and optimality.
-In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
+In particular, the QUBO model delivers high quality solutions in a few seconds while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
 \end{abstract}

### 2020-07-17 paper.tex
Editted Section 1

@@
 
-In recent years, there has been a growing need to efficiently solve WFLO~
-\cite{wilson2014continuous},
-while maintaining competitive (albeit not always optimal) solutions. 
-The focus on quickly solving the problem was mainly motivated by the fact
+In recent years, there has been a growing need to quickly find high quality (albeit not always optimal) WFLO solutions~\cite{wilson2014continuous}.
+The focus on quickly solving the problem was primarily motivated by the fact
 that modern farms cover large areas and include hundreds, and
@@
 and performing a sensitivity analysis to test the impact of those changes~\cite{wilson2018evolutionary}.
-The efficient approaches proposed in~\cite{wilson2013learning,wilson2014continuous} do not
-take into account the physical model of the wind farm, and hence
-do not aim at optimizing the actual energy objective.
+
+%The efficient approaches proposed in~\cite{wilson2013learning,wilson2014continuous} do not
+%take into account the physical model of the wind farm, and hence
+%do not aim at optimizing the actual energy objective.
 
@@
 model represents WFLO as a quadratic constrained optimization
-problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical solvers
-that have been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
+problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical software
+that has been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
 second model is a quadratic unconstrained binary optimization
@@
 By employing the two types of advances (software and hardware) for solving
-quadratic problems, we enjoy the best of both worlds: 
+quadratic problems, 
 we are able to provide fast and high performing solutions (in terms of the total energy). Moreover, our approach is based
@@
 %, often outperforming existing solutions when provided sufficient run-time.
-Furthermore, we observe that the DA often provides solutions of
+Furthermore, we observe that the DA finds solutions of
 similar quality to software-based approaches but with two orders of
@@
 %of the proposed methods.
+In Section~\ref{sec:discussion}, we discuss the results and their practical implications.
 Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.
@@
 \section{Discussion and Practical Implications}
+\label{sec:discussion}
 

### 2020-07-17 paper.tex
Editted Section 3

@@
 an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
-$n=20$, the farm area will be $2km\times 2km$.  We
+$n=20$, the farm area will be $2 \mbox{ km} \times 2 \mbox{ km}$.  We
 assume that the terrain is flat and that we have no constraints
@@
 thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
-Furthermore, there is no natural way of introducing constraints into genetic methods (see e.g., \cite{sorkhabi2018constrained}), which
-leads to a potentially costly feasibility check for every solution found.
+Furthermore, it is awkward to introduce constraints into genetic methods (see e.g., \cite{sorkhabi2018constrained}), leading to potentially costly feasibility checks and problem-specific repair operations for every solution found.
 
@@
 
-Recently, several efficient WFLO solutions based on developmental models were introduced in~\cite{wilson2013learning,wilson2014continuous}.
+Recently, several efficient WFLO solutions based on developmental models were introduced~\cite{wilson2013learning,wilson2014continuous}.
 These works do not represent the actual underlying physical model of the system, and while providing fast solutions to WFLO,
 they do not attempt to optimize on the actual objective function (that is, the total energy that a farm produced) nor capture the complexities of the underlying problem.
-For example, the approach in~\cite{wilson2013learning} models turbine interference using a parametric probabilistic model (assuming a Weibull distribution).
+For example, one approach~\cite{wilson2013learning} models turbine interference using a parametric probabilistic model (assuming a Weibull distribution).
 In our quadratic models, we explicitly represent the energy term, and the physical dependencies between the turbines (i.e., wakes and proximity constraints).
-Furthermore, our models are
-optimizing on an empirically established energy objective, namely the linear superposition expression in Eq.~(\ref{eq:ls}).
+Furthermore, our models
+optimize an empirically established energy objective, the linear superposition expression in Eq.~(\ref{eq:ls}).
 
@@
 the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
-ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}
+ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2 \mbox{km} \times 2 \mbox{km}$.\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}
 

### 2020-07-17 paper.tex
Edits through the rest of the document.

@@
 %optimization hardware such as the DA. 
-A thorough empirical 
+A thorough numerical 
 evaluation using a commercial solver and specialized QUBO hardware
@@
 optimization solutions to the layout problem.
-The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO} and Section~\ref{sec:eval} demonstrates the value of our approach via a thorough empirical evaluation.
+The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO} and Section~\ref{sec:eval} demonstrates the value of our approach via a thorough  numerical evaluation.
 %of the proposed methods.
@@
 In this section, we 
-present an empirical evaluation of the two quadratic models based on 
+present an numerical evaluation of the two quadratic models based on 
 12 standard WFLO instances.
@@
 
-Our results show empirically that the quadratic programming models that we
+Our results show numerically that the quadratic programming models that we
 introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
@@
 that the computational complexity of WFLO is $\mathcal{NP}$-hard. Nonetheless,
-our empirical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
+our numerical evaluation shows that when exact solvers like Gurobi~\cite{gurobi} are
 coupled with our \qcls{} and \quls{} models they are able to solve small WFLO instances to optimality in less than 3600 seconds.

### 2020-07-17 paper.tex
Extended the contributions

@@
 The main
-contributions of this paper are as follows: \todo{Chris, perhaps we can strengthen the contribution list}
+contributions of this paper are as follows:
 \begin{enumerate} 
@@
 \item We prove that the QCOP variation of WFLO is $\mathcal{NP}$-hard.
-\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
+
+\item We show through numerical experiments that, when solved with optimization software (Gurobi), the QCOP model results in state-of-the-art performance in terms of finding solutions with guaranteed bounds on solution quality, including finding and proving optimal solutions for a subset of the tested problem instances.
+
+\item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), the QUBO model finds solutions competitive with the state of the art in one second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.
+Such a short runtime enables an interactive wind farm design cycle.
+
+  %\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
 \end{enumerate}

### 2021-09-29 paper.tex
Updated abstract

@@
 %role as a source of renewable energy and 
-  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines.
+  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site and information about the wind patterns, the problem is to decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference between turbines.
 Given the pairwise wake interactions, it is natural
@@
 to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization approaches.
-However, state-of-the-art solutions often trade-off between efficiency (providing quick designs) and accuracy (guaranteeing optimal energy solutions).
-In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when accuracy is the goal.
-To this end, we exploit recent advances in both optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
+However, state-of-the-art solutions often trade-off between speed in producing designs and quality in terms of finding and proving optimal solutions.
+In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when quality is more important.
+To this end, we exploit recent advances in optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
 %It is a key problem 
@@
 that
-improve the state of the art and strike a balance between efficiency and optimality.
+improve the state of the art and strike a balance between speed and quality.
 In particular, the QUBO model delivers high quality solutions in a few seconds while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.

### 2021-10-16 Jason.tex
Update the paper (see purple font) according to the latest results with DA3 and Gurobi 9.1.1. Create a new branch.

--- 2021-09-29 paper.tex
@@
 %\address[3]{}
-\address[add1]{Faculty of Information, University of Toronto, 140 St. George St., Toronto, Canada}
+\address[add1]{Rotman School of Management, University of Toronto, 105 St. George St., Toronto, Canada}
 \cortext[cor1]{Corresponding author}
@@
 
-In this work, we use the first generation of the Fujitsu Digital Annealer (DA) \cite{aramon2019physics}, a specialized CMOS hardware capable of solving QUBO problems with up to 1024 binary variables.
+\textcolor{purple}{In this work, we use the Fujitsu Digital Annealer (DA), a recent computer architecture designed for solving QUBO problems \cite{matsubara2020digital}.
+The third generation DA (DA3), a hybrid system of hardware and software, is capable of representing QUBOs with up to 100000 variables.
+For our DA environment, the coefficients for the quadratic terms range from $-2^{62}$ to $2^{62}$ and those for the linear terms range from $-2^{73}$ to $2^{73}$ \cite{DA3}.
 The algorithm used by the DA is based on simulated annealing \cite{kirkpatrick1983optimization}, however it takes advantage of the massive parallelization provided by the custom CMOS hardware and uses a dynamic offset mechanism to escape local minima \cite{aramon2019physics}.
-The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising} and communication \cite{naghsh2019digitally,rahman2019ising}.
+In addition, DA supports a dedicated bit flip mechanism, over a subset of variables belonging to one-hot equivalent constraints when using DA3.
+The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising} and communication \cite{naghsh2019digitally,rahman2019ising}.}
 
@@
 Our main results are: \begin{itemize}
-	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than existing state-of-the-art exact methods on 10 of 12 instances.
-		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
+	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than existing state-of-the-art exact methods on \textcolor{purple}{8 of 12 instances}.
+	\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least \textcolor{purple}{one orders of magnitude} shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
 	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances.
-Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard ones.
-%  \todo{Arik: Is this correct?
+Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on \textcolor{purple}{6 of 12 benchmark problems and for 3 of the 6 hard ones.} %  \todo{Arik: Is this correct?
 Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12.
@@
 reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
-We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve \qcls{}, \ilpls{} and \qcss. 
+We used state-of-the-art commercial software, \textcolor{purple}{Gurobi v9.1.1~\cite{gurobi}}, to solve \qcls{}, \ilpls{} and \qcss. 
 We shall denote the problems using their names and the solvers used as follows: \qclsgrb{} (GRB for Gurobi), \ilplsgrb{}, and \qcssgrb{}.
@@
 we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
-found until that point in time. The exact methods were solved using Gurobi's integer programming and quadratic programming solvers on a 
-Ubuntu PC with Intel(R) Core(TM) i7-9700 8 core CPU@3.00 GHz with 12 MB cache and 8 GB RAM.
+found until that point in time.
+The exact methods were solved using Gurobi's integer programming and quadratic programming solvers on a \textcolor{purple}{Window PC with Intel(R) Core(TM) i7-8700K CPU @3.20GHz with 16 GB RAM.}
 
 For the DA, we used a single configuration that was found to work best across the different instances. 
-The DA was run for increasingly large time limits ranging from $0.7$ seconds to $65$ seconds where we no longer observe an improvement in 
-performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~\eqref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and sum of the two penalty terms, namely $\lambda_1+\lambda_2$, 
-is the highest value that the Digital Annealer allows.\footnote{The DA has an upper limit on the size of objective function coefficients.}
+The DA was run for increasingly large time limits \textcolor{purple}{ranging from $1$ second to $10$ seconds} where we no longer observe an improvement in 
+performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~\eqref{QUBO}) are selected such that $\lambda_1= \lambda_2$ \textcolor{purple}{and their magnitudes are tuned by DA3 automatically.}
 
@@
 		\toprule
-		Wind Directions  & n  & m  & \qulsda{}  & \qulsgrb  & \qcls & \qcss  & \ilpls  \\
+		Wind Directions  & n  & m  & \qulsda{}  & \qulsgrb  & \qclsgrb & \qcssgrb  & \ilplsgrb  \\
 		\toprule
 		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
-		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.10  & 15739.57     \\
-		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
+		& & 30   & \textbf{15742.93} & \textbf{15742.93}  & 15740.69  & 15741.81  & 15741.81     \\
+		& & 40 & \textbf{19265.21} & 18954.28 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
 		\cline{2-8}
 		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
-		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
-		&&40   & 21874.04 & 21643.56  & 21905.50 & \textbf{21977.17} & \textbf{21977.17}        \\
+		&&30   & 16399.56 & 16602.85  & 16761.18  & \textbf{16774.37} & 16760.91                  \\
+		&&40   & \textbf{21973.80} & 21854.57  & 21907.27 & 21918.85 & 21894.62       \\
 		\hline
-		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
-		&& 30  & \textbf{27450.99} & 27418.74 & 27442.90 & 27387.06 & 27247.50                     \\
-		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19175.88 & 19186.58 \\
+		&& 30  & 27440.08 & 27442.90 & \textbf{27443.34} & 27393.49 & 27406.48                     \\
+		&& 40   & \textbf{35409.58} & 34938.65 & 34888.37 & 34824.24  & 34867.31          \\
 		\cline{2-8}
-		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
-		&&30   & \textbf{27975.55} & 27443.80  & 27752.00 & 27492.11  & 26367.20                      \\
-		&&40   & 35604.45 & 35127.32 & \textbf{35665.60}   & 35065.85 & 33451.94 \\
+		&  \multirow{3}{*}{20}   & 20   & \textbf{19437.49}  & 19310.96 & 19391.76 & 19333.97  & 19331.84            \\
+		&&30   & \textbf{27939.08} & 27689.15  & 27885.49 & 27747.88  & 27859.50                      \\
+		&&40   & 35547.27 & 35352.31 & \textbf{35623.11}   & 35229.93 & 35521.68 \\
 		\bottomrule                   
@@
 		\toprule
-		Wind Directions  & n  & m  & \qulsda  & \qulsgrb  & \qcls & \qcss  & \ilpls  \\
+		Wind Directions  & n  & m  & \qulsda{}  & \qulsgrb  & \qclsgrb & \qcssgrb  & \ilplsgrb  \\
 		\toprule
 		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
-		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
-		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
+		& & 30   & \textbf{15742.93} & \textbf{15742.93}  & 15740.69  & 15741.81  & 15741.81     \\
+		& & 40 & \textbf{19265.21} & 18954.28 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
 		\cline{2-8}
 		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
-		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.20 }                 \\
-		&&40   & 21874.04 & 21832.50  & 21963.81 & \textbf{21977.17} & \textbf{21977.17}        \\
+		&&30   & 16399.56 & 16769.09  & 16761.18  & \textbf{16774.37} & 16770.17                  \\
+		&&40   & \textbf{21973.80} & 21891.25  & 21919.68 & 21918.85 & 21973.69       \\
 		\hline
-		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19213.96 \\
-		&& 30  & \textbf{27450.99} & 27442.90 & 27442.90 & 27387.06 & 27389.21                     \\
-		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34856.45          \\
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19175.88 & 19186.58 \\
+		&& 30  & 27440.08 & 27442.90 & \textbf{27443.34} & 27393.49 & 27422.53                     \\
+		&& 40   & \textbf{35409.58} & 34938.65 & 34938.65 & 34824.24  & 34867.31          \\
 		\cline{2-8}
-		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19394.58 & 19336.33 & 19354.23  & 19368.86            \\
-		&&30   & \textbf{27975.55} & 27785.74  & 27752.00 & 27492.11  & 27633.20                      \\
-		&&40   & 35604.45 & 35480.90 & \textbf{35665.60}   & 35065.85 & 35210.78 \\
+		&  \multirow{3}{*}{20}   & 20   & 19437.49  & \textbf{19437.52} & 19411.21 & 19388.72  & 19337.11            \\
+		&&30   & \textbf{27939.08} & 27812.30  & 27885.49 & 27866.93  & 27869.54                      \\
+		&&40   & 35547.27 & 35571.63 & \textbf{35623.11}   & 35264.47 & 35539.74 \\
 		\bottomrule                   
@@
 comparable results to state-of-the-art exact methods. For the harder instances (with $36$ wind directions),
-the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in 5 out of 6 instances. For easier instances, 
-the linear approach (\ilplsgrb) finds better solutions.
+the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in \textcolor{purple}{all 6 instances.
+For easier instances,
+the DA-based approach (\qulsda{}) is also dominating as it finds the best solutions in 5 out of 6 solutions.}
 %\todo{@Chris, Jason: do we have a good explanation for this?}. NO
@@
 We also observe that the DA at 10 seconds is, therefore, still competitive with the other approaches,
-while \qclsgrb{} finds the best solutions in 10 out of the 12 scenarios. 
+while \qclsgrb{} finds the best solutions in \textcolor{purple}{6} out of the 12 scenarios. 
 Our quadratic approaches dominate all six hard scenarios (36 wind directions).
@@
 
-We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
-presents the MRE over the first $50$ seconds, while Figure~\ref{fig:results2} corresponds
+We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  \textcolor{purple}{Figure~\ref{fig:results1} 
+presents the MRE over the first $100$ seconds,} while Figure~\ref{fig:results2} corresponds
 to the full run (after $3600$ seconds).
@@
 outperform the rest of the techniques across the entire runtime time horizon that was tested.
-Specifically, for most instances, \qclsgrb{} gives the best solution at 3600 seconds, while the DA-based approach
-yields competitive solutions after less than 10 seconds. 
+\textcolor{purple}{Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, while the DA-based approach \qulsda{}
+yields competitive solutions after less than 10 seconds. However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
 Furthermore, 
-\qclsgrb{} converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method.
-
-The results also show that on average 
-the \ilplsgrb{} method starts with poor solutions, yet ends up improving over \qcssgrb{} and \qulsgrb{} after around 1250 seconds.
+\qclsgrb{} converges to the best solutions faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method when the runtime is relatively small.}
+
+\textcolor{purple}{The results also show that on average 
+the \ilplsgrb{} method starts with poor solutions, and ends up improving over \qcssgrb{} after around 3 seconds.}
 
@@
 %and averaged the speed-up factors across the instances. 
-The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. 
+The average speed-up factors were found to be \textcolor{purple}{$666$ and $100$ for $1$ and $10$ seconds, respectively.} \textcolor{red}{(Arik, please change the speed-up ratios, the rest of the paragraph, and next paragraph.)}
 Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
@@
 
-	\includegraphics[width=\textwidth]{energy_50.png}  
-	\caption{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 50 seconds).}%
+	\includegraphics[width=\textwidth]{Figures/MRE_100s_DA3.png}  
+	\caption{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 100 seconds).}%
 
@@
 
-	\includegraphics[width=\textwidth]{energy_3600.png}   \\%\\
+	\includegraphics[width=\textwidth]{Figures/MRE_3600s_DA3.png}   \\%\\
 	\caption{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 3600 seconds).}%
@@
 
-Our results show numerically that the quadratic programming models that we
+\textcolor{purple}{Our results show numerically that the quadratic programming models that we
 introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
-First, the two methods are powerful approaches to solving WFLO. Specifically, \qcls{} yields the best performance (after 3600 seconds) for most problem instances
-and especially,
-for the hard instances with 36 wind directions. In addition, the runtime performance of \qcls{} is second
-only to the Digital Annealer.
+First, the two methods are powerful approaches to solving WFLO.
+Specifically, \qcls{} yields the best performance (after 3600 seconds).
 Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the \qulsda{} approach.
-\qulsda{} yields the best results in 7 out of 12
-instances, shows competitive performance on the other instances, and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second).
+\qulsda{} yields the best results in 8 out of 12 problem instances (both easy instances with 1 wind direction and hard instances with 36 wind directions).
+In addition, the runtime performance of \qulsda{} is second
+only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes} (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second) \textcolor{red}{(Arik, please change the speed-up factors.)}.
 

### 2021-10-26 Jason.tex
Update README.
Edits to get it ready for the next submission up to Sect 3.5.1,

--- 2021-09-29 paper.tex
@@
 
+%\address[1]{}
+
+%\address[2]{40 St. George St., Toronto, Canada}
+\author[add2]{Jiachen Zhang}
+\ead{jasonzjc@mie.utoronto.ca}
+
 \author[add2]{Eldan Cohen}
@@
 
-%\address[1]{}
-
-%\address[2]{40 St. George St., Toronto, Canada}
-\author[add2]{Jiachen Zhang}
-\ead{jasonzjc@mie.utoronto.ca}
-
 \author[add2]{J. Christopher Beck\corref{cor1}}
@@
 However, state-of-the-art solutions often trade-off between speed in producing designs and quality in terms of finding and proving optimal solutions.
-In this work, we aim to find a balancing approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when quality is more important.
+In this work, we aim to find a balanced approach to obtain WFLO solutions quickly for interactive design \emph{and} solve the problem to optimality when quality is more important.
 To this end, we exploit recent advances in optimization hardware and software that target quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving quadratic unconstrained binary optimization (QUBO) problems.
@@
 \section{Introduction}
-% Todo
+
 The goal in the wind farm layout optimization (WFLO)
@@
 quadratic problems, 
-we are able to provide fast and high performing solutions (in terms of the total energy). Moreover, our approach is based
+we are able to quickly find high quality solutions in terms of the total energy. Moreover, our approach is based
 on a physical model of the wind farm, taking into account
 exact calculations of the total energy and the wake effects. 
-Therefore, our approach achieves the desired balance between efficiency and effectiveness of the resulting design, in a model-based fashion.
+These two models together achieve a balance between efficiency and effectiveness of the resulting design, in a model-based fashion.
 %, a hardware-based
@@
 %, often outperforming existing solutions when provided sufficient run-time.
-Furthermore, we observe that the DA finds solutions of
-similar quality to software-based approaches but with two orders of
-magnitude shorter run-time.
+
+%Furthermore, we observe that the DA finds solutions of
+%similar quality to software-based approaches but with two orders of
+%magnitude shorter run-time.
+
 %The ability to quickly obtain competitive
@@
 based on the common
-notation and description presented in Zhang et al.\ \cite{Zhang2014}.  Subsequently,
-we provide a literature review of existing mathematical optimization
-methods for solving WFLO, focusing on their similarities and
+notation and description of Zhang et al.\ \cite{Zhang2014}.  Subsequently,
+we review the literature on existing mathematical optimization
+methods for solving WFLO, focusing on similarities and
 differences with our approach.
@@
 choose to focus on a
-novel approach for solving WFLO, and hence consider the most
+novel approach for solving WFLO, and hence consider a more
 abstract wind farm model. Extension of our framework to include such additional problem characteristics is left as future research.
@@
 in advance and denoted by $m$. To place these $m$ turbines, 
-we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
-In what follows, we explain the two effects using a toy example of placing a single turbine on an 
+we must consider two types of interference effect: \emph{proximity constraints} and \emph{wakes}.
+In what follows, we explain the two effects using a toy example with an 
 $8\times8$ grid (see Figure~\ref{fig:field_model}).
@@
 \paragraph{Wakes}
-Turbines influence each other through \emph{wake} interference effects~\cite{jensen1983note,shakoor2016wake}.
+Turbines influence each other through \emph{wake} interference effects that reduce the effective power produced by a turbine due to upstream turbines that change the airflow dynamics~\cite{jensen1983note,shakoor2016wake}.
 %that are
 %referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
-Wakes cause reduction of the effective power produced by a turbine due
-to upstream turbines that change the airflow dynamics. In
+ In
 Figure~\ref{fig:field_model}, the wind is assumed to be
@@
 function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
-a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
+one regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-easterly wind blowing at a free velocity of $12\mbox{ km/h}$. The
 probability of this regime is $0.008$.
@@
 These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
-the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
+the best solution one can obtain for the given model of the problem. Donovan was the first to solve WFLO using a
 MILP approach based on the LS energy expression~\cite{donovan2005wind}.
@@
 In fact, Section~\ref{sec:QUBO4WFLO} shows formally that WFLO is computationally hard, a result
-that we have not yet encountered in the WFLO literature. Therefore,
+that is not surprising but that has apparently not explicitly appeared in the WFLO literature. Therefore,
 when one considers solving multiple wind farm design problems, if a fast solution is of essence,
@@
 \paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
-quick and, hopefully, sufficiently accurate solutions. Several works have solved 
+quick and, hopefully, sufficiently high-quality solutions. Several works have solved 
 WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
@@
 These methods do not guarantee termination 
-(i.e., they may run forever without achieving a quality criteria), and even when they do terminate (in practical cases) they do
+(i.e., they may run forever without achieving a quality criteria), and even when they do terminate they do
 not guarantee optimal solutions. Hence, one must often
 introduce non-quality related stopping criteria, 
-thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
+compromising the quality of the obtained solution~\cite{davis1991handbook}.
 Furthermore, it is awkward to introduce constraints into genetic methods (see e.g., \cite{sorkhabi2018constrained}), leading to potentially costly feasibility checks and problem-specific repair operations for every solution found.
@@
 was made using a local search approach~\cite{ozturk2004heuristic}. 
-The solution does not guarantee optimality, 
+The approach does not guarantee optimality, 
 yet it circumvents the termination and feasibility checking issues of the evolutionary
-methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
+methods. The local search attempts to apply either move, remove or add actions given an incumbent 
 feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
@@
 our methodology is generic since it is based on a declarative problem model and
-on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.
+on the ability of specialized hardware to solve %(non-ad-hoc)
+quadratic problems.
 
@@
 in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
-These results lead to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
+These results led to the conclusion that both types of methods (exact and approximate) should be considered. Therefore, we introduce a unified paradigm, namely quadratic programming, that enables
 both exact and approximate solutions to WFLO. Further, formulating WFLO using quadratic programming enables us to employ 
@@
 
-We start by presenting the inputs and decision variables into our optimization approaches.
+We start by presenting the inputs and decision variables of our optimization approaches.
 Subsequently, we introduce the two quadratic formulations of WFLO and 
@@
 In operations research, the generic exact approach for solving IQCOPs is the branch-and-cut based mixed integer programming (MIP) \cite{bonami2008algorithmic}.
-A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function.
+
+%A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function.
 This relaxation is used to generate cuts, implied constraints that tighten the relaxation, to identify sub-trees that can be discarded without search because they are guaranteed not to contain a solution better than the current incumbent, and to drive the heuristic branching that systematically searches for a solution.
-Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.
-
-Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements have delivered orders of magnitude improved performance \cite{Bixby07a}.
+%Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.
+
+Commercial MIP solvers have seen extensive development over the past 20 years with hardware-independent algorithmic improvements have delivered orders of magnitude improved performance \cite{Bixby07a}.
 More recently the development of such solvers has focused on representing and solving quadratic problems such as \qcls{}~\cite{Furini19a}.

### 2021-10-26 Jason.tex
Small update

@@
 Recent years have seen significant advancement in the use of specialized hardware platforms, such as adiabatic and gate-based quantum computers, Digital/CMOS annealers, and neuromorphic computers \cite{coffrin2019evaluating}, to solve hard combinatorial optimization problems.
-Interestingly, a large number of these platforms support the optimization of quadratic functions over binary variables by using Ising models \cite{johnson2011quantum}, an equivalent representation to QUBO \cite{bian2010ising}, as their mathematical abstraction.
+A large number of these platforms support the optimization of quadratic functions over binary variables by using Ising models \cite{johnson2011quantum}, an equivalent representation to QUBO \cite{bian2010ising}, as their mathematical abstraction.
 

### 2021-10-26 Jason_v2.tex
Jason_v2.tex combines the updates by Jason and the recent updates by Chris.

--- 2021-10-16 Jason.tex
@@
 %\ead[url]{homepageldan}
-
-%\address[1]{}
-
-%\address[2]{40 St. George St., Toronto, Canada}
-\author[add2]{Jiachen Zhang}
-\ead{jasonzjc@mie.utoronto.ca}
 

### 2021-10-26 Jason_v2.tex

@@
 \item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), the QUBO model finds solutions competitive with the state of the art in one second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.
-Such a short runtime enables an interactive wind farm design cycle.
+\todo{Jason: Does the previous sentence need to be updated given the new results?} Such a short runtime enables an interactive wind farm design cycle.
 
@@
 
-Recent years have seen significant advancement in the use of specialized hardware platforms, such as adiabatic and gate-based quantum computers, Digital/CMOS annealers, and neuromorphic computers \cite{coffrin2019evaluating}, to solve hard combinatorial optimization problems.
+Recent years have seen significant advancement in the use of specialized hardware platforms, such as adiabatic and gate-based quantum computers, digital/CMOS annealers, and neuromorphic computers \cite{coffrin2019evaluating}, to solve hard combinatorial optimization problems.
 A large number of these platforms support the optimization of quadratic functions over binary variables by using Ising models \cite{johnson2011quantum}, an equivalent representation to QUBO \cite{bian2010ising}, as their mathematical abstraction.
 
-\textcolor{purple}{In this work, we use the Fujitsu Digital Annealer (DA), a recent computer architecture designed for solving QUBO problems \cite{matsubara2020digital}.
+In this work, we use the Fujitsu Digital Annealer (DA), a recent computer architecture designed for solving QUBO problems \cite{matsubara2020digital}.
 The third generation DA (DA3), a hybrid system of hardware and software, is capable of representing QUBOs with up to 100000 variables.
@@
 The algorithm used by the DA is based on simulated annealing \cite{kirkpatrick1983optimization}, however it takes advantage of the massive parallelization provided by the custom CMOS hardware and uses a dynamic offset mechanism to escape local minima \cite{aramon2019physics}.
-In addition, DA supports a dedicated bit flip mechanism, over a subset of variables belonging to one-hot equivalent constraints when using DA3.
-The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising} and communication \cite{naghsh2019digitally,rahman2019ising}.}
+In addition, DA3 supports a dedicated bit flip mechanism, over a subset of variables belonging to one-hot equivalent constraints.
+The Fujitsu Digital Annealer has been previously applied to variety of problems in different research areas such as machine learning \cite{cohen2020ising,salehinejad2019ising} and communication \cite{naghsh2019digitally,rahman2019ising}.
 
@@
 standard exact optimization solvers that run until obtaining an optimal solution, the DA runs until reaching a time limit and returns the best found solution.
-In our experiments, we analyze the quality of obtained solutions for different time limits.
+In our experiments, we analyze the quality of solutions for different time limits.
 
@@
 Our main results are: \begin{itemize}
-	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than existing state-of-the-art exact methods on \textcolor{purple}{8 of 12 instances}.
-	\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least \textcolor{purple}{one orders of magnitude} shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
+	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than all tested state-of-the-art exact methods on 8 of 12 instances.
+	\item This performance is achieved for the \quls{} model with a runtime that is at least one order of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
+\todo{Jason: This claim (``one order of magnitude'') need to be made consistent with the claims and results later in the papaer.}
 	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances.
-Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on \textcolor{purple}{6 of 12 benchmark problems and for 3 of the 6 hard ones.} %  \todo{Arik: Is this correct?
+Compared to all tested previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 6 of 12 benchmark problems and for 3 of the 6 hard ones.
+%  \todo{Arik: Is this correct?
 Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12.
@@
 
-We start by outlining the experimental design of our evaluation, followed by an overview of our main results,
+We start by outlining the experimental design of our evaluation, followed by our main results,
 and conclude the section with a discussion of the results and their practical implications.
@@
 from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement} that we denote with the set $\mathcal{W}$. 
-The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
-2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
-and 3 variations for number of turbines ($m \in \{20, 30, 40\}$). 
+The 12 benchmarks result from varying two wind settings (1 and 36 wind directions),
+two grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
+and three turbine cardinalities ($m \in \{20, 30, 40\}$). 
 For the $36$ wind directions setting, the velocities come from the probability distribution presented in Figure~\ref{fig:prob_wind}.
@@
 the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
-ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2 \mbox{km} \times 2 \mbox{km}$.\footnote{The implementation of constructing the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}
+ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2 \mbox{km} \times 2 \mbox{km}$.\footnote{The implementation of construction of the 12 standard WFLO instances can be found in~\url{https://bit.ly/2YIjTl1}.}
 
@@
 reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
-We used state-of-the-art commercial software, \textcolor{purple}{Gurobi v9.1.1~\cite{gurobi}}, to solve \qcls{}, \ilpls{} and \qcss. 
+We used state-of-the-art commercial software, Gurobi v9.1.1~\cite{gurobi}, to solve \qcls{}, \ilpls{} and \qcss. 
 We shall denote the problems using their names and the solvers used as follows: \qclsgrb{} (GRB for Gurobi), \ilplsgrb{}, and \qcssgrb{}.
@@
 found until that point in time.
-The exact methods were solved using Gurobi's integer programming and quadratic programming solvers on a \textcolor{purple}{Window PC with Intel(R) Core(TM) i7-8700K CPU @3.20GHz with 16 GB RAM.}
+The exact methods were solved using Gurobi's integer programming and quadratic programming solvers on a Windows PC with Intel(R) Core(TM) i7-8700K CPU @3.20GHz with 16 GB RAM.
 
 For the DA, we used a single configuration that was found to work best across the different instances. 
-The DA was run for increasingly large time limits \textcolor{purple}{ranging from $1$ second to $10$ seconds} where we no longer observe an improvement in 
-performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~\eqref{QUBO}) are selected such that $\lambda_1= \lambda_2$ \textcolor{purple}{and their magnitudes are tuned by DA3 automatically.}
+The DA was run for increasingly large time limits ranging from $1$ second to $10$ seconds, at which point we no longer observe an improvement in 
+performance. The penalty terms $\lambda_1$ and $\lambda_2$ (see~Eq.~\eqref{QUBO}) are selected such that $\lambda_1= \lambda_2$ and their magnitudes are tuned by DA3 automatically.
 
@@
 is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
-For example, $$S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, 10, w_1)$$ is the ratio between the time that it takes \qclsgrb{} to 
+For example, $$S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, 10, w_1)$$ is the ratio of the time that it takes \qclsgrb{} to 
 reach the same quality of solution as \qulsda{} achieves after $10$ seconds on problem $w_1$. Since
@@
 The runtime for the DA was always less than $10$ seconds.
+
+\todo{Jason: Fix tables to fit on width of page.}
 
@@
 comparable results to state-of-the-art exact methods. For the harder instances (with $36$ wind directions),
-the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in \textcolor{purple}{all 6 instances.
-For easier instances,
-the DA-based approach (\qulsda{}) is also dominating as it finds the best solutions in 5 out of 6 solutions.}
+the three quadratic LS-based approaches (\qulsda{}, \qulsgrb{}, and \qclsgrb{}) dominate the others by finding the best (or equal) solutions in all 6 instances. For easier instances, 
+the DA-based approach (\qulsda{}) also dominates as it finds the best solutions in 5 out of 6 instances.
 %\todo{@Chris, Jason: do we have a good explanation for this?}. NO
@@
 We also observe that the DA at 10 seconds is, therefore, still competitive with the other approaches,
-while \qclsgrb{} finds the best solutions in \textcolor{purple}{6} out of the 12 scenarios. 
+while \qclsgrb{} finds the best solutions in 6 out of the 12 scenarios. 
 Our quadratic approaches dominate all six hard scenarios (36 wind directions).
@@
 
-We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  \textcolor{purple}{Figure~\ref{fig:results1} 
-presents the MRE over the first $100$ seconds,} while Figure~\ref{fig:results2} corresponds
+We now turn to the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
+presents the MRE over the first $100$ seconds, while Figure~\ref{fig:results2} corresponds
 to the full run (after $3600$ seconds).
@@
 outperform the rest of the techniques across the entire runtime time horizon that was tested.
-\textcolor{purple}{Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, while the DA-based approach \qulsda{}
-yields competitive solutions after less than 10 seconds. However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
+Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, while the DA-based approach \qulsda{}
+yields the best solutions after 10 seconds. However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
 Furthermore, 
-\qclsgrb{} converges to the best solutions faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method when the runtime is relatively small.}
-
-\textcolor{purple}{The results also show that on average 
-the \ilplsgrb{} method starts with poor solutions, and ends up improving over \qcssgrb{} after around 3 seconds.}
+\qclsgrb{} converges to the best solutions faster than all other exact methods. In fact, it is inferior only to the approximate \qulsda{} method when the runtime is relatively small.
+
+The results also show that on average 
+the \ilplsgrb{} method starts with poor solutions, and ends up improving over \qcssgrb{} after around 3 seconds.
 
@@
 %and averaged the speed-up factors across the instances. 
-The average speed-up factors were found to be \textcolor{purple}{$666$ and $100$ for $1$ and $10$ seconds, respectively.} \textcolor{red}{(Arik, please change the speed-up ratios, the rest of the paragraph, and next paragraph.)}
+The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. \todo{(Arik, please change the speed-up ratios, the rest of the paragraph, and next paragraph.)}
 Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
@@
 
-\textcolor{purple}{Our results show numerically that the quadratic programming models that we
+Our results show numerically that the quadratic programming models that we
 introduced in Section~\ref{sec:QUBO4WFLO} (\qcls{} and \quls{}) are useful in two ways.
@@
 In addition, the runtime performance of \qulsda{} is second
-only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes} (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second) \textcolor{red}{(Arik, please change the speed-up factors.)}.
+only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second) \todo{(Arik, please change the speed-up factors.)}.
 
@@
 multiple wind farm layout problems at design time.
-Moreover, it allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that come
+Moreover, it allows the flexibility to tune the various WFLO parameters (including turbine specs, number of turbines, and farm sizes) without incurring a steep computational cost that comes
 with (re-)solving the problem using exact methods. The quality loss due to the use of an approximate method 
 that exploits advanced technology is negligible. This balance between
-runtime and total energy, points at the potential contribution of the proposed quadratic framework for solving WFLO.
+runtime and total energy, highlights the contribution of the proposed quadratic framework for solving WFLO.
 
@@
 
-In future work, we aim at extending our quadratic models to include additional 
-WFLO aspects, such as noise, non-flat terrain, and multiple objectives. 
-Moreover, we wish to explore real-world wind farm design problems 
-that may be larger and more complex than the 12 standard instances 
-and solve them 
-using our quadratic programming approaches.
+%In future work, we aim at extending our quadratic models to include additional 
+%WFLO aspects, such as noise, non-flat terrain, and multiple objectives. 
+%Moreover, we wish to explore real-world wind farm design problems 
+%that may be larger and more complex than the 12 standard instances 
+%and solve them 
+%using our quadratic programming approaches.
 
@@
 \label{app1}
+
+\todo{Arik \& Jason: do these need to be updated?}
 
@@
 Table~\ref{tab:results3} presents the speedup results across all problem instances
-for every LS-based method (computed using the SS energy term). The speedup is calculated according to Eq.~(\ref{eq:speedup}). Table~\ref{tab:results4} shows similar results for the speedup factor
+for every LS-based method (computed using the SS energy formulation).
+The speedup is calculated according to Eq.~(\ref{eq:speedup}).
+Table~\ref{tab:results4} shows similar results for the speedup factor
 compared to DA at 10 seconds. The average lower bound on the speedup at $1$ seconds

### 2021-11-04 Jason_v2.tex
Change a graph and speedup ratios based on new results.

@@
 
-\item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), the QUBO model finds solutions competitive with the state of the art in one second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.
-\todo{Jason: Does the previous sentence need to be updated given the new results?} Such a short runtime enables an interactive wind farm design cycle.
+\item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), \textcolor{purple}{the QUBO model finds solutions competitive with the state of the art in 1 second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.} Such a short runtime enables an interactive wind farm design cycle.
 
@@
 	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than all tested state-of-the-art exact methods on 8 of 12 instances.
-	\item This performance is achieved for the \quls{} model with a runtime that is at least one order of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
-\todo{Jason: This claim (``one order of magnitude'') need to be made consistent with the claims and results later in the papaer.}
+	\item \textcolor{purple}{This performance is achieved for the \quls{} model with a runtime that is two to three orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.}
 	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances.
@@
 
-\todo{Jason: Fix tables to fit on width of page.}
+% \todo{Jason: Fix tables to fit on width of page.}
 
@@
 	\small
+	\centering
+	\scalebox{0.76}{
 	\begin{tabular}{| c | c | c | c | c | c | c | c |}
@@
 	\end{tabular}
-
+	}
 	\vspace{0.5em}
@@
 outperform the rest of the techniques across the entire runtime time horizon that was tested.
-Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, while the DA-based approach \qulsda{}
-yields the best solutions after 10 seconds. However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
+Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, \textcolor{purple}{while the DA-based approach \qulsda{}
+yields the best solutions after 6 seconds.} However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
 Furthermore, 
@@
 
-The results also show that on average 
-the \ilplsgrb{} method starts with poor solutions, and ends up improving over \qcssgrb{} after around 3 seconds.
+\textcolor{purple}{The results also show that on average 
+the \ilplsgrb{} method starts with poor solutions, and ends up improving over \qcssgrb{} after around 2 seconds.}
 
@@
 %and averaged the speed-up factors across the instances. 
-The average speed-up factors were found to be $666$ and $100$ for $1$ and $10$ seconds, respectively. \todo{(Arik, please change the speed-up ratios, the rest of the paragraph, and next paragraph.)}
-Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
+\textcolor{purple}{The average speed-up factors were found to be $988$ and $131$ for $1$ and $10$ seconds, respectively.} Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
 The \emph{maximum} speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
@@
 When compared
-to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the mean speedup factor compared to \ilplsgrb{} 
-is $1385$ ($180$) and for the \qulsgrb{} it is $900$ ($152$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
+to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  \textcolor{purple}{At $1$ ($10$) seconds, the mean speedup factor compared to \ilplsgrb{} 
+is $1430$ ($197$) and for the \qulsgrb{} it is $1013$ ($134$).} We compare against the LS based methods as they directly optimize on the same objective as the DA.  
 For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in~\ref{app1}.
@@
 
-	\includegraphics[width=\textwidth]{Figures/MRE_100s_DA3.png}  
-	\caption{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 100 seconds).}%
+	\includegraphics[width=\textwidth]{Figures/MRE_100s_DA3_NEW.png}  
+	\caption{\textcolor{purple}{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 100 seconds).}}%
 
@@
 In addition, the runtime performance of \qulsda{} is second
-only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $1$ second) \todo{(Arik, please change the speed-up factors.)}.
+only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $10$ seconds).
 
@@
 
-\todo{Arik \& Jason: do these need to be updated?}
-
 \begin{table}[t!]
 	\small
+	\centering
 	\begin{tabular}{| c | c | c | c | c | c | }
@@
 		\cline{2-6}
-		&\multirow{3}{*}{20}   & 20  & 1  & 1  & 259         \\
-		&&30   & 3 & 2  & 9    \\
-		&&40   & 32 & 22  & 602        \\
+		&\multirow{3}{*}{20}   & 20  & 1  & 1  & 30         \\
+		&&30   & 1 & 1  & 2    \\
+		&&40   & 3600 & 143 & 3600        \\
 		\hline
-		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20 & 1       & 1661 & 1  \\
-		&& 30  & 3600 & 3600 & 3600                     \\
-		&&40   & 1 & 3600 & 1          \\
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20 & 1       & 3600 & 1  \\
+		&& 30  & 1 & 1184 & 1                     \\
+		&&40   & 1 & 1 & 1          \\
 		\cline{2-6}
-		&  \multirow{3}{*}{20}   & 20   & 3600  & 538 & 944            \\
-		&&30   & 22 & 3600 & 3304                     \\
-		&&40   & 60 & 3600 & 1717 \\
+		&  \multirow{3}{*}{20}   & 20   & 3600  & 3600 & 2224            \\
+		&&30   & 3600 & 3600 & 3600                     \\
+		&&40   & 66 & 3600 & 1683 \\
 		\bottomrule                   
@@
 	\vspace{0.5em}
-	\caption{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 1 second. The value 3600 implies a lower bound on the speedup factor.}\label{tab:results3}
+	\caption{\textcolor{purple}{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 1 second. The value 3600 implies a lower bound on the speedup factor.}}\label{tab:results3}
 \end{table}
@@
 	\small
+	\centering
 	\begin{tabular}{| c | c | c | c | c | c | }
@@
 		\cline{2-6}
-		&\multirow{3}{*}{20}   & 20  & 1  & 1  & 25.9         \\
-		&&30   & 0.3 & 0.2  & 0.9    \\
-		&&40   & 6.1 & 2.2  & 360        \\
+		&\multirow{3}{*}{20}   & 20  & 1  & 1  & 3         \\
+		&&30   & 0.1 & 0.1  & 0.2    \\
+		&&40   & 360 & 14.3  & 360        \\
 		\hline
-		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20 & 0.1       & 360 & 0.2  \\
+		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20 & 0.1       & 360 & 0.1  \\
+		&& 30  & 0.1 & 360 & 0.2                     \\
+		&& 40   & 360 & 360 & 360          \\
+		\cline{2-6}
+		&  \multirow{3}{*}{20}   & 20   & 360  & 360 & 222.4            \\
 		&& 30   &  360  & 360 & 360                   \\
-		&&40   & 0.1 & 360 & 0.1          \\
-		\cline{2-6}
-		&  \multirow{3}{*}{20}   & 20   & 360  & 360 & 360            \\
-		&&30   &  360  & 360 & 360                    \\
-		&&40   &  8  & 360 & 360 \\
+		&& 40   &  6.6  & 360 & 168.3 \\
 
@@
 	\vspace{0.5em}
-	\caption{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 10 second. The value 360 implies a lower bound on the speedup factor.}\label{tab:results4}
+	\caption{\textcolor{purple}{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 10 second. The value 360 implies a lower bound on the speedup factor.}}\label{tab:results4}
 \end{table}
@@
 Table~\ref{tab:results4} shows similar results for the speedup factor
-compared to DA at 10 seconds. The average lower bound on the speedup at $1$ seconds
-is 665,	1511, and 949 for \qclsgrb, \qulsgrb, and \ilplsgrb, respectively. For the $10$ second
+compared to DA at 10 seconds. \textcolor{purple}{The average lower bound on the speedup at $1$ seconds
+is 988, 1430, and 1013 for \qclsgrb, \ilplsgrb, and \qulsgrb,  respectively. For the $10$ second
 results, the average
-speedup factor is at least 99, 196, and	166 for the respective methods.
+speedup factor is at least 131, 197, and 134 for the respective methods.}
 For instances where the speedup factor in Table~\ref{tab:results4} is 10 times less than in Table~\ref{tab:results3},

### 2021-11-05 main.tex
Editted to end of Sect 2.

--- 2021-11-04 Jason_v2.tex
@@
 this wake modeling yields intractable optimization problems
-that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
+that have been approximated using either quadratic or linear functions~\cite{turner2014new}.
 An alternative formulation of WFLO relies on a less accurate wake model, using linear superpositioning
@@
 
-\item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), \textcolor{purple}{the QUBO model finds solutions competitive with the state of the art in 1 second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.} Such a short runtime enables an interactive wind farm design cycle.
+\item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), the QUBO model finds solutions competitive with the state of the art in one second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.
+Such a short runtime enables an interactive wind farm design cycle.
 
@@
 
-We assume that the number of identical turbines to be placed is known 
+As is common, we assume that the number of identical turbines to be placed is known 
 in advance and denoted by $m$. To place these $m$ turbines, 
@@
 CP is an optimization technique that has grown from the AI literature that does not make the assumptions on the functional form of the constraints or objective \cite{rossi2006handbook}.
-The authors compared existing MILP and QP models to their approach for directly optimizing the SS objective and showed that, while promising for smaller problems, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.
+The authors compared existing MILP and QP models to their approach for directly optimizing the SS objective and showed that, while promising for smaller problems, applying CP was computationally intractable for the larger standard WFLO benchmarks.
 
@@
 yet it circumvents the termination and feasibility checking issues of the evolutionary
-methods. The local search attempts to apply either move, remove or add actions given an incumbent 
-feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
+methods. The local search attempts to apply either move, remove, or add actions given an incumbent 
+feasible solution. When a turbine is moved, removed, or added, the new solution is evaluated. The procedure terminates 
 when a stopping criteria is reached.
@@
 To overcome this limitation, Rivas et al.\ \cite{rivas2009solving} used simulated annealing (SA), 
-a neighborhood search method that allows visits to infeasible solutions. 
-SA creates a well-connected solution space that allows the algorithm to escape 
+a neighborhood search method that creates a connected solution space that allows the algorithm to escape 
 local optima. The SA method was shown to be superior to the approach of Ozturk \& Norman~\cite{ozturk2004heuristic} 

### 2021-11-05 main.tex
Finished edit

@@
 
-Commercial MIP solvers have seen extensive development over the past 20 years with hardware-independent algorithmic improvements have delivered orders of magnitude improved performance \cite{Bixby07a}.
+Commercial MIP solvers have seen extensive development over the past 20 years with hardware-independent algorithmic improvements delivering orders of magnitude improved performance \cite{Bixby07a}.
 More recently the development of such solvers has focused on representing and solving quadratic problems such as \qcls{}~\cite{Furini19a}.
@@
 
-Unlike %exact techninques
+Unlike %exact techniques
 standard exact optimization solvers that run until obtaining an optimal solution, the DA runs until reaching a time limit and returns the best found solution.
@@
 	\item For given time limit, the Digital Annealer (DA) achieves a solution quality as good or as better than all tested state-of-the-art exact methods on 8 of 12 instances.
-	\item \textcolor{purple}{This performance is achieved for the \quls{} model with a runtime that is two to three orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.}
+	\item This performance is achieved for the \quls{} model with a runtime that is two to three orders of magnitude shorter than \qcls{} on state-of-the-art commercial solvers for hard WFLO instances.
 	\item Given sufficient runtime, solving the \qcls{} model with commercial software yields strong results on most of the instances.
@@
 outperform the rest of the techniques across the entire runtime time horizon that was tested.
-Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, \textcolor{purple}{while the DA-based approach \qulsda{}
-yields the best solutions after 6 seconds.} However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
+Specifically, for most instances, \qclsgrb{} gives the best solutions at 3600 seconds, while the DA-based approach \qulsda{}
+yields the best solutions after 6 seconds. However, the good performance of \quls{} only occurs when the solver is the DA. In fact, \qulsgrb{} trails all the other methods.
 Furthermore, 
@@
 
-\textcolor{purple}{The results also show that on average 
-the \ilplsgrb{} method starts with poor solutions, and ends up improving over \qcssgrb{} after around 2 seconds.}
+The results also show that on average 
+the \ilplsgrb{} method starts with poor solutions but improves over \qcssgrb{} after around 2 seconds.
 
@@
 %and averaged the speed-up factors across the instances. 
-\textcolor{purple}{The average speed-up factors were found to be $988$ and $131$ for $1$ and $10$ seconds, respectively.} Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
+The average speed-up factors were found to be $988$ and $131$ for $1$ and $10$ seconds, respectively.
+Per instance, the \emph{minimum} speedup was $1$ for $1$ second and $0.1$ for $10$ seconds: for some instances \qclsgrb{} achieved the same quality as \qulsda{} in less time.
 The \emph{maximum} speedup was (at least) $3600$ for $1$ second and (at least) $360$ for $10$ seconds: for those instances
 \qclsgrb{} was not able to find a comparable solution to \qulsda{} within the $3600$ time limit. As a consequence, 
-the mean \qulsda{} speed-up is actually even higher than computed above. \end{sloppypar}
+the mean \qulsda{} speed-up is actually higher than computed above. \end{sloppypar}
 %\footnote{The speed-up factors  were rounded to have 0 values after the floating point.}
@@
 When compared
-to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  \textcolor{purple}{At $1$ ($10$) seconds, the mean speedup factor compared to \ilplsgrb{} 
-is $1430$ ($197$) and for the \qulsgrb{} it is $1013$ ($134$).} We compare against the LS based methods as they directly optimize on the same objective as the DA.  
+to the other LS-based methods, the DA speed-up is even larger than when compared to \qclsgrb{}.  At $1$ ($10$) seconds, the mean speedup factor compared to \ilplsgrb{} 
+is $1430$ ($197$) and for the \qulsgrb{} it is $1013$ ($134$). We compare against the LS based methods as they directly optimize on the same objective as the DA.  
 For full speedup results see Tables~\ref{tab:results3} and~\ref{tab:results4} in~\ref{app1}.
@@
 	\includegraphics[width=\textwidth]{Figures/MRE_100s_DA3_NEW.png}  
-	\caption{\textcolor{purple}{Mean Relative Energy (wrt best feasible solution) vs. Runtime (after 100 seconds).}}%
+	\caption{Mean Relative Error (wrt best known feasible solution) vs. Runtime (after 100 seconds).}
 
@@
 In addition, the runtime performance of \qulsda{} is second
-only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes (a speed-up factor of 3 orders of magnitude when the DA runs for $10$ seconds).
+only to the \qclsgrb{} and outperforms all methods in terms of performance in short runtimes (a speed-up factor of three orders of magnitude when the DA runs for $10$ seconds).
 
@@
 	\vspace{0.5em}
-	\caption{\textcolor{purple}{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 1 second. The value 3600 implies a lower bound on the speedup factor.}}\label{tab:results3}
+	\caption{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 1 second. The value 3600 implies a lower bound on the speedup factor.}\label{tab:results3}
 \end{table}
@@
 	\vspace{0.5em}
-	\caption{\textcolor{purple}{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 10 second. The value 360 implies a lower bound on the speedup factor.}}\label{tab:results4}
+	\caption{Speedup factor per LS-based method for all instances w.r.t \qulsda{} at 10 second. The value 360 implies a lower bound on the speedup factor.}\label{tab:results4}
 \end{table}
@@
 Table~\ref{tab:results4} shows similar results for the speedup factor
-compared to DA at 10 seconds. \textcolor{purple}{The average lower bound on the speedup at $1$ seconds
+compared to DA at 10 seconds. The average lower bound on the speedup at $1$ seconds
 is 988, 1430, and 1013 for \qclsgrb, \ilplsgrb, and \qulsgrb,  respectively. For the $10$ second
 results, the average
-speedup factor is at least 131, 197, and 134 for the respective methods.}
+speedup factor is at least 131, 197, and 134 for the respective methods.
 For instances where the speedup factor in Table~\ref{tab:results4} is 10 times less than in Table~\ref{tab:results3},

### 2021-11-05 main.tex
Add footnote and acknowledgement.

@@
 
-\item We show through numerical experiments that, when solved with specialized hardware (Fujitsu's Digital Annealer), the QUBO model finds solutions competitive with the state of the art in one second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.
+\item We show through numerical experiments that, when solved with specialized hardware, the QUBO model finds solutions competitive with the state of the art in one second, representing a computational speed-up of two to three orders of magnitude compared to software based methods.
 Such a short runtime enables an interactive wind farm design cycle.
@@
 The third generation DA (DA3), a hybrid system of hardware and software, is capable of representing QUBOs with up to 100000 variables.
-For our DA environment, the coefficients for the quadratic terms range from $-2^{62}$ to $2^{62}$ and those for the linear terms range from $-2^{73}$ to $2^{73}$ \cite{DA3}.
+For our DA environment\footnote{All experiments were conducted on the Digital Annealer environment prepared exclusively for the University of Toronto’s research.}, the coefficients for the quadratic terms range from $-2^{62}$ to $2^{62}$ and those for the linear terms range from $-2^{73}$ to $2^{73}$ \cite{DA3}.
 The algorithm used by the DA is based on simulated annealing \cite{kirkpatrick1983optimization}, however it takes advantage of the massive parallelization provided by the custom CMOS hardware and uses a dynamic offset mechanism to escape local minima \cite{aramon2019physics}.
@@
 Specifically, \qcls{} yields the best performance (after 3600 seconds).
-Second, the QUBO model can be solved with recent specialized hardware, Fujitsu's DA, as demonstrated by the \qulsda{} approach.
+Second, the QUBO model can be solved with recent computer architecture, Fujitsu's DA, as demonstrated by the \qulsda{} approach.
 \qulsda{} yields the best results in 8 out of 12 problem instances (both easy instances with 1 wind direction and hard instances with 36 wind directions).
@@
 coupled with our \qcls{} and \quls{} models they are able to solve small WFLO instances to optimality in less than 3600 seconds.
-We also show that when an advanced hardware is used with the QUBO representation, large WFLO instances can be solved quickly with no significant degradation in the total energy compared to exact methods.
+We also show that when an advanced computer architecture is used with the QUBO representation, large WFLO instances can be solved quickly with no significant degradation in the total energy compared to exact methods.
 This runtime improvement and state-of-the-art performance in terms of the total energy of the resulting solutions makes our quadratic framework useful for designing new wind farms, while varying multiple input parameters.
@@
 %using our quadratic programming approaches.
+
+\section*{Acknowledgement}
+The authors would like to thank Fujitsu Ltd. and Fujitsu Consulting (Canada) Inc. for providing financial support and access to Digital Annealer at the University of Toronto.
 

### 2021-11-07 main.tex
Minor modifications.

@@
 The third generation DA (DA3), a hybrid system of hardware and software, is capable of representing QUBOs with up to 100000 variables.
-For our DA environment\footnote{All experiments were conducted on the Digital Annealer environment prepared exclusively for the University of Toronto’s research.}, the coefficients for the quadratic terms range from $-2^{62}$ to $2^{62}$ and those for the linear terms range from $-2^{73}$ to $2^{73}$ \cite{DA3}.
+For our DA environment\footnote{All experiments were conducted on the Digital Annealer environment prepared exclusively for the research at the University of Toronto.}, the coefficients for the quadratic terms range from $-2^{62}$ to $2^{62}$ and those for the linear terms range from $-2^{73}$ to $2^{73}$ \cite{DA3}.
 The algorithm used by the DA is based on simulated annealing \cite{kirkpatrick1983optimization}, however it takes advantage of the massive parallelization provided by the custom CMOS hardware and uses a dynamic offset mechanism to escape local minima \cite{aramon2019physics}.
@@
 \section*{Acknowledgement}
-The authors would like to thank Fujitsu Ltd. and Fujitsu Consulting (Canada) Inc. for providing financial support and access to Digital Annealer at the University of Toronto.
+The authors would like to thank Fujitsu Ltd. and Fujitsu Consulting (Canada) Inc. for providing financial support and access to the Digital Annealer at the University of Toronto.
 
